### 2008-09-10 developments/PDB/BIT_VECTOR_DEFINITION.tex

\documentstyle[11pt,supertabular]{article}
\setlength{\topmargin}{13mm}
\setlength{\headheight}{0mm}
\setlength{\headsep}{0mm}
\setlength{\textheight}{225mm}
\setlength{\oddsidemargin}{0mm}
\setlength{\textwidth}{160mm}
\input{psfig}

\begin{document}
\begin{sloppypar}

% DEFINITION DES CARACTERES MATHEMATIQUES B
%------------------------------------------
\def\@setmcodes#1#2#3{{\count0=#1 \count1=#3
	\loop \global\mathcode\count0=\count1 \ifnum \count0<#2
	\advance\count0 by1 \advance\count1 by1 \repeat}}

\@setmcodes{`A}{`Z}{"7441}
\@setmcodes{`a}{`z}{"7461}

\mathcode`\;="8000 % Makes ; active in math mode
{\catcode`\;=\active \gdef;{\semicolon\;}}
\mathchardef\semicolon="003B
%    Nominal distance from top of paper to top of page
\topmargin 0 pt
\textheight 53\baselineskip

%   Left margin on odd-numbered pages
\oddsidemargin  0.15 in
%   Left margin on even-numbered pages
\evensidemargin 0.35 in
%   Width of marginal notes.
\marginparwidth 1 in
%   Note that \oddsidemargin = \evensidemargin
\oddsidemargin 0.25 in
\evensidemargin 0.25 in
\marginparwidth 0.75 in
\textwidth 5.875 in % Width of text line.

\setlength{\parindent}{0pt}
\setlength{\parskip}{0ex}

% DEFINITION DES FONTS
%---------------------
% The AMS extra symbol fonts are loaded.
% Note: sometimes called euxm10
\font\msx=msam10
% Note: sometimes called euym10
\font\msy=msbm10

\newfam\msxfam \textfont\msxfam=\msx
\newfam\msyfam \textfont\msyfam=\msy

\def\famletter#1{\ifcase #1 0\or 1\or 2\or 3\or 4\or 5\or 6\or 7\or
	8\or 9\or A\or B\or C\or D\or E\or F\fi}

\edef\fx{\famletter\msxfam}
\edef\fy{\famletter\msyfam}

\def\bbold{\fam\msyfam \msy}

% SYMBOLES B
%-----------
% makes a quoted expression in mathematical text
\def\token#1{\hbox{`$#1$'}}
% used for error messages in Z specs
\def\report#1{\hbox{`{\tt #1}'}}

% \@myop makes an operator, with a strut to defeat TeX's vertical adjustment.
\def\@myop#1{\mathop{\mathstrut{#1}}\nolimits}

% This underscore doesn't have the little kern --- you get an italic
% correction anyway in math mode.
\def\_{\leavevmode \vbox{\hrule width0.5em}}

% Save \q as \xq for quantifiers q.
\let\xforall=\forall
\let\xexists=\exists
\let\xlambda=\lambda
\let\xmu=\mu

% \p and \f make arrows with 1 and 2 crossings resp.
\def\p#1{\mathrel{\ooalign{\hfil$\mapstochar\mkern 5mu$\hfil\cr$#1$}}}
\def\f#1{\mathrel{\ooalign{\hfil
	$\mapstochar\mkern 3mu\mapstochar\mkern 5mu$\hfil\cr$#1$}}}

\let\mc=\mathchardef

\def	\pow		{\mbox{${\cal P}$}}
\def	\po1		{\mbox{${\cal P}_1$}}
\let	\cross		\times
\def	\lambda		{\@myop{\xlambda}}
\def	\lnot		{\neg\;}
\def	\land		{\mathrel{\wedge}}
\def	\lor		{\mathrel{\vee}}
\let	\implies	\Rightarrow
\let	\iff		\Leftrightarrow
\def	\forall		{\@myop{\xforall}}
\def	\exists		{\@myop{\xexists}}
\def	\semi		{\mathrel{\comp}}
\def	\ssemi		{\mathbin{\rm ;}}
\let	\ensembleVide	\emptyset
\let	\rel		\leftrightarrow
\def	\dom		{\@myop{\sf dom}}
\def	\ran		{\@myop{\sf ran}}
\def	\id		{\@myop{\sf id}}
\def	\comp		{\mathbin{\raise
			0.6ex\hbox{\oalign{\hfil$\scriptscriptstyle
			\rm o$\hfil\cr\hfil$\scriptscriptstyle\rm 9$\hfil}}}}
\def	\para		{\mbox{$\mid\mid$}}
\mc	\dres		"2\fx43
\mc	\rres		"2\fx42
\def	\ndres		{\mathbin{{\dres} \llap{$-$}}}
\def	\nrres		{\mathbin{{\rres}\llap{$-$}}}
\def	\lover		{\mathbin{{\dres} \llap{$-\!\!\!\!-\!$}}}
\def	\rover		{\mathbin{{\rres}\llap{$\!-\!\!\!-$}}}
\let	\fun		\rightarrow
\def	\pfun		{\p\fun}
\def	\pinj		{\p\inj}
\mc	\inj		"3\fx1A
\def	\psurj		{\p\surj}
\mc	\surj		"3\fx10
\def	\bij		{\surj\!\!\!\!\!\!\!\inj}
\def	\nat		{\mbox{${\cal N}$}}
\def	\na1		{\mbox{${\cal N}_1$}}
\def	\num		{\mbox{${\cal Z}$}}
\def	\int		{\mbox{${\cal Z}$}}
\def	\rat		{\mbox{${\cal Q}$}}
\def	\div		{\mathbin{\rm /}}
\def	\mod		{\mathbin{\bf mod}}
\def	\upto		{\mathbin{\ldotp\ldotp}}
\def	\finset		{\mbox{${\cal F}$}}
\def	\finse1		{\mbox{${\cal F}_1$}}
\def	\ffun		{\f\fun}
\def	\finj		{\f\inj}
\def	\seq		{\@myop{\rm seq}}
\def	\cat		{\mathbin{\raise 0.8ex\hbox{$\mathchar"2\fx61$}}}
\def	\sep		{\hspace*{.05in}}

\setcounter{secnumdepth}{0}
\setcounter{tocdepth}{0}

%-------------------%
% Debut du document %
%-------------------%

\bf MACHINE

\hspace*{0.20in}\it BIT\_VECTOR\_DEFINITION

\vspace*{4mm}
\bf SEES

\hspace*{0.20in}\it BIT\_DEFINITION

\vspace*{4mm}
\bf CONSTANTS

\hspace*{0.20in}\it BIT\_VECTOR\rm ,

\hspace*{0.20in}\it bv\_size\rm , \it bv\_catenate\rm , \it bv\_sub\rm , \it bv\_zero\rm , \it bv\_one\rm ,

\hspace*{0.20in}\it bv\_not\rm ,\hspace*{0.20in}\it bv\_and\rm ,\hspace*{0.20in}\it bv\_or\rm , \it bv\_xor\rm ,

\hspace*{0.20in}\it bv\_at\rm , \it bv\_set\rm , \it bv\_clear

\vspace*{4mm}
\bf PROPERTIES

\hspace*{0.20in}\it BIT\_VECTOR \rm = \bf seq1 \rm (\it BIT\rm )  $\land$

\vspace*{4mm}
\hspace*{0.20in}\it bv\_size  $\in$  \it BIT\_VECTOR  $\fun$   $ \na1 $   $\land$

\hspace*{0.20in}\it bv\_size \rm =  $\lambda$ \rm (\it bv\rm )\rm .\rm (\it bv  $\in$  \it BIT\_VECTOR  $\mid$  \bf size\rm (\it bv\rm )\rm )  $\land$

\vspace*{4mm}
\hspace*{0.20in}\it bv\_catenate  $\in$  \it BIT\_VECTOR  $\times$  \it BIT\_VECTOR  $\fun$  \it BIT\_VECTOR  $\land$

\hspace*{0.20in}\it bv\_catenate \rm =  $\lambda$ \rm (\it v1\rm ,\it v2\rm ) \rm . \rm (\it v1  $\in$  \it BIT\_VECTOR  $\land$  \it v2 $\in$  \it BIT\_VECTOR  $\mid$  \it v2  $\cat$  \it v1\rm )  $\land$

\vspace*{4mm}
\hspace*{0.20in}\it bv\_sub  $\in$  \it BIT\_VECTOR  $\times$   $\nat$   $\times$   $\nat$   $\fun$  \it BIT\_VECTOR  $\land$

\hspace*{0.20in}\it bv\_sub \rm =  $\lambda$ \rm (\it bv\rm , \it low\rm , \it high\rm ) \rm . \rm (\it bv  $\in$  \it BIT\_VECTOR  $\land$  \it low $\in$   $\nat$   $\land$  \it high $\in$   $\nat$   $\land$  \it low  $\leq$  \it high  $\mid$  \rm 0 $\upto$ \rm (\it high\rm -\it low\rm )  $\dres$  \it bv\rm )  $\land$

\vspace*{4mm}
\hspace*{0.20in}\it bv\_zero  $\in$   $ \na1 $   $\fun$  \it BIT\_VECTOR  $\land$

\hspace*{0.20in}\it bv\_zero \rm =  $\lambda$ \rm (\it sz\rm ) \rm . \rm (\it sz  $\in$   $ \na1 $   $\mid$  \rm (\rm 1 $\upto$ \it sz\rm )  $\times$  \rm \{\rm 0\rm \}\rm )  $\land$

\vspace*{4mm}
\hspace*{0.20in}\it bv\_one  $\in$   $ \na1 $   $\fun$  \it BIT\_VECTOR  $\land$

\hspace*{0.20in}\it bv\_one \rm =  $\lambda$ \rm (\it sz\rm ) \rm . \rm (\it sz  $\in$   $ \na1 $   $\mid$  \rm (\rm 1 $\upto$ \it sz\rm )  $\times$  \rm \{\rm 1\rm \}\rm )  $\land$

\vspace*{4mm}
\hspace*{0.20in}\it bv\_not  $\in$  \it BIT\_VECTOR  $\fun$  \it BIT\_VECTOR  $\land$

\hspace*{0.20in}\it bv\_not \rm =  $\lambda$ \rm (\it v1\rm )\rm .\rm (\it v1 $\in$  \it BIT\_VECTOR  $\mid$   $\lambda$ \rm (\it idx\rm )\rm .\rm (\it idx  $\in$  \rm 1 $\upto$ \bf size\rm (\it v1\rm )  $\mid$  \it bit\_not\rm (\it v1\rm (\it idx\rm )\rm )\rm )\rm )  $\land$

\vspace*{4mm}
\hspace*{0.20in}\it bv\_and  $\in$  \it BIT\_VECTOR  $\times$  \it BIT\_VECTOR  $\fun$  \it BIT\_VECTOR  $\land$

\hspace*{0.20in}\it bv\_and \rm =  $\lambda$ \rm (\it v1\rm ,\it v2\rm )\rm .\rm (\it v1 $\in$  \it BIT\_VECTOR  $\land$  \it v2  $\in$  \it BIT\_VECTOR  $\land$  \bf size\rm (\it v1\rm ) \rm = \bf size\rm (\it v2\rm )  $\mid$

\hspace*{0.40in} $\lambda$ \rm (\it idx\rm )\rm .\rm (\it idx  $\in$  \rm 1 $\upto$ \bf size\rm (\it v1\rm )  $\mid$  \it bit\_and\rm (\it v1\rm (\it idx\rm )\rm ,\it v2\rm (\it idx\rm )\rm )\rm )\rm )  $\land$

\vspace*{4mm}
\hspace*{0.20in}\it bv\_or  $\in$  \it BIT\_VECTOR  $\times$  \it BIT\_VECTOR  $\fun$  \it BIT\_VECTOR  $\land$

\hspace*{0.20in}\it bv\_or \rm =  $\lambda$ \rm (\it v1\rm ,\it v2\rm )\rm .\rm (\it v1 $\in$  \it BIT\_VECTOR  $\land$  \it v2  $\in$  \it BIT\_VECTOR  $\land$  \bf size\rm (\it v1\rm ) \rm = \bf size\rm (\it v2\rm )  $\mid$

\hspace*{0.40in} $\lambda$ \rm (\it idx\rm )\rm .\rm (\it idx  $\in$  \rm 1 $\upto$ \bf size\rm (\it v1\rm )  $\mid$  \it bit\_or\rm (\it v1\rm (\it idx\rm )\rm ,\it v2\rm (\it idx\rm )\rm )\rm )\rm )  $\land$

\vspace*{4mm}
\hspace*{0.20in}\it bv\_xor  $\in$  \it BIT\_VECTOR  $\times$  \it BIT\_VECTOR  $\fun$  \it BIT\_VECTOR  $\land$

\hspace*{0.20in}\it bv\_xor \rm =  $\lambda$ \rm (\it v1\rm ,\it v2\rm )\rm .\rm (\it v1 $\in$  \it BIT\_VECTOR  $\land$  \it v2  $\in$  \it BIT\_VECTOR  $\land$  \bf size\rm (\it v1\rm ) \rm = \bf size\rm (\it v2\rm )  $\mid$

\hspace*{0.40in} $\lambda$ \rm (\it idx\rm )\rm .\rm (\it idx  $\in$  \rm 1 $\upto$ \bf size\rm (\it v1\rm )  $\mid$  \it bit\_xor\rm (\it v1\rm (\it idx\rm )\rm ,\it v2\rm (\it idx\rm )\rm )\rm )\rm )  $\land$

\vspace*{4mm}
\hspace*{0.20in}\it bv\_at  $\in$  \it BIT\_VECTOR  $\times$   $\nat$   $\fun$  \it BIT  $\land$

\hspace*{0.20in}\it bv\_at \rm =  $\lambda$ \rm (\it v1\rm ,\it idx\rm )\rm .\rm (\it v1 $\in$  \it BIT\_VECTOR  $\land$  \it idx  $\in$   $\nat$   $\land$  \it idx $<$ \bf size\rm (\it v1\rm )  $\mid$  \it v1\rm (\it idx\rm +\rm 1\rm )\rm )  $\land$

\vspace*{4mm}
\hspace*{0.20in}\it bv\_set  $\in$  \it BIT\_VECTOR  $\times$   $\nat$   $\fun$  \it BIT\_VECTOR  $\land$

\hspace*{0.20in}\it bv\_set \rm =  $\lambda$ \rm (\it v1\rm ,\it idx\rm )\rm .\rm (\it v1 $\in$  \it BIT\_VECTOR  $\land$  \it idx  $\in$   $\nat$   $\land$  \it idx $<$ \bf size\rm (\it v1\rm )  $\mid$  \it v1  $\lover$  \rm \{ \it idx\rm +\rm 1  $\mapsto$  \rm 1 \rm \}\rm )  $\land$

\hspace*{0.20in}\it bv\_clear  $\in$  \it BIT\_VECTOR  $\times$   $\nat$   $\fun$  \it BIT\_VECTOR  $\land$

\hspace*{0.20in}\it bv\_clear \rm =  $\lambda$ \rm (\it v1\rm ,\it idx\rm )\rm .\rm (\it v1 $\in$  \it BIT\_VECTOR  $\land$  \it idx  $\in$   $\nat$   $\land$  \it idx $<$ \bf size\rm (\it v1\rm )  $\mid$  \it v1  $\lover$  \rm \{ \it idx\rm +\rm 1  $\mapsto$  \rm 0 \rm \}\rm )

\vspace*{4mm}
\bf ASSERTIONS

\hspace*{0.20in} $\forall$  \it bv\rm . \rm (\it bv  $\in$  \it BIT\_VECTOR  $\implies$  \it bv\_size\rm (\it bv\_not\rm (\it bv\rm )\rm ) \rm = \it bv\_size\rm (\it bv\rm )\rm )\rm ;

\hspace*{0.20in} $\forall$  \it bv\rm . \rm (\it bv  $\in$  \it BIT\_VECTOR  $\implies$  \rm (\it bv\_not\rm (\it bv\_not\rm (\it bv\rm )\rm ) \rm = \it bv\rm )\rm )\rm ;

\vspace*{4mm}
\hspace*{0.20in} $\forall$  \rm (\it v1\rm ,\it v2\rm )\rm . \rm (\it v1  $\in$  \it BIT\_VECTOR  $\land$  \it v2  $\in$  \it BIT\_VECTOR  $\implies$

\hspace*{0.60in}\it bv\_size\rm (\it bv\_catenate\rm (\it v1\rm , \it v2\rm )\rm ) \rm = \it bv\_size\rm (\it v1\rm ) \rm + \it bv\_size\rm (\it v2\rm )\rm )\rm ;

\vspace*{4mm}
\hspace*{0.20in} $\forall$  \rm (\it bv\rm ,\it low\rm ,\it high\rm )\rm . \rm (\it bv  $\in$  \it BIT\_VECTOR  $\land$  \it low  $\in$   $\nat$   $\land$  \it high  $\in$   $\nat$   $\land$  \it low  $\leq$  \it high  $\implies$

\hspace*{0.60in}\it bv\_size\rm (\it bv\_sub\rm (\it bv\rm , \it low\rm , \it high\rm )\rm ) \rm = \rm 1\rm +\it high\rm -\it low\rm )\rm ;

\vspace*{4mm}
\hspace*{0.20in} $\forall$  \rm (\it v1\rm ,\it v2\rm )\rm . \rm (\it v1  $\in$  \it BIT\_VECTOR  $\land$  \it v2  $\in$  \it BIT\_VECTOR  $\land$  \it bv\_size\rm (\it v1\rm ) \rm = \it bv\_size\rm (\it v2\rm )  $\implies$

\hspace*{0.60in}\it bv\_size\rm (\it bv\_and\rm (\it v1\rm , \it v2\rm )\rm ) \rm = \it bv\_size\rm (\it v2\rm )\rm )\rm ;

\hspace*{0.20in} $\forall$  \rm (\it v1\rm ,\it v2\rm )\rm . \rm (\it v1  $\in$  \it BIT\_VECTOR  $\land$  \it v2  $\in$  \it BIT\_VECTOR\hspace*{0.10in} $\land$  \it bv\_size\rm (\it v1\rm ) \rm = \it bv\_size\rm (\it v2\rm )  $\implies$

\hspace*{0.60in}\rm (\it bv\_and\rm (\it v1\rm , \it v2\rm ) \rm = \it bv\_and\rm (\it v2\rm , \it v1\rm )\rm )\rm )\rm ;

\hspace*{0.20in} $\forall$  \rm (\it v1\rm ,\it v2\rm ,\it v3\rm )\rm . \rm (\it v1  $\in$  \it BIT\_VECTOR  $\land$  \it v2  $\in$  \it BIT\_VECTOR $\land$  \it v3  $\in$  \it BIT\_VECTOR\hspace*{0.10in} $\land$  \it bv\_size\rm (\it v1\rm ) \rm = \it bv\_size\rm (\it v2\rm )  $\land$  \it bv\_size\rm (\it v1\rm ) \rm = \it bv\_size\rm (\it v3\rm )  $\implies$

\hspace*{0.60in}\rm (\it bv\_and\rm (\it v1\rm , \it bv\_and\rm (\it v2\rm ,\it v3\rm )\rm ) \rm = \it bv\_and\rm (\it bv\_and\rm (\it v1\rm , \it v2\rm )\rm , \it v3\rm )\rm )\rm )\rm ;

\hspace*{0.20in} $\forall$  \it bv\rm . \rm (\it bv  $\in$  \it BIT\_VECTOR  $\implies$  \rm (\it bv\_and\rm (\it bv\rm , \it bv\_zero\rm (\it bv\_size\rm (\it bv\rm )\rm )\rm ) \rm = \it bv\_zero\rm (\it bv\_size\rm (\it bv\rm )\rm )\rm )\rm )\rm ;

\hspace*{0.20in} $\forall$  \it bv\rm . \rm (\it bv  $\in$  \it BIT\_VECTOR  $\implies$  \rm (\it bv\_and\rm (\it bv\rm , \it bv\_one\rm (\it bv\_size\rm (\it bv\rm )\rm )\rm ) \rm = \it bv\rm )\rm )\rm ;

\vspace*{4mm}
\hspace*{0.20in} $\forall$  \rm (\it v1\rm ,\it v2\rm )\rm . \rm (\it v1  $\in$  \it BIT\_VECTOR  $\land$  \it v2  $\in$  \it BIT\_VECTOR  $\land$  \it bv\_size\rm (\it v1\rm ) \rm = \it bv\_size\rm (\it v2\rm )  $\implies$

\hspace*{0.60in}\it bv\_size\rm (\it bv\_or\rm (\it v1\rm , \it v2\rm )\rm ) \rm = \it bv\_size\rm (\it v1\rm )\rm )\rm ;

\hspace*{0.20in} $\forall$  \rm (\it v1\rm ,\it v2\rm )\rm . \rm (\it v1  $\in$  \it BIT\_VECTOR  $\land$  \it v2  $\in$  \it BIT\_VECTOR  $\land$  \it bv\_size\rm (\it v1\rm ) \rm = \it bv\_size\rm (\it v2\rm )  $\implies$

\hspace*{0.60in}\it bv\_size\rm (\it bv\_or\rm (\it v1\rm , \it v2\rm )\rm ) \rm = \it bv\_size\rm (\it v2\rm )\rm )\rm ;

\hspace*{0.20in} $\forall$  \rm (\it v1\rm ,\it v2\rm )\rm . \rm (\it v1  $\in$  \it BIT\_VECTOR  $\land$  \it v2  $\in$  \it BIT\_VECTOR  $\land$  \it bv\_size\rm (\it v1\rm ) \rm = \it bv\_size\rm (\it v2\rm )  $\implies$

\hspace*{0.60in}\rm (\it bv\_or\rm (\it v1\rm , \it v2\rm ) \rm = \it bv\_or\rm (\it v2\rm , \it v1\rm )\rm )\rm )\rm ;

\hspace*{0.20in} $\forall$  \rm (\it v1\rm ,\it v2\rm ,\it v3\rm )\rm . \rm (\it v1  $\in$  \it BIT\_VECTOR  $\land$  \it v2  $\in$  \it BIT\_VECTOR $\land$  \it v3  $\in$  \it BIT\_VECTOR  $\land$  \it bv\_size\rm (\it v1\rm ) \rm = \it bv\_size\rm (\it v2\rm )  $\land$  \it bv\_size\rm (\it v1\rm ) \rm = \it bv\_size\rm (\it v3\rm )  $\implies$

\hspace*{0.60in}\rm (\it bv\_or\rm (\it v1\rm , \it bv\_or\rm (\it v2\rm ,\it v3\rm )\rm ) \rm = \it bv\_or\rm (\it bv\_or\rm (\it v1\rm , \it v2\rm )\rm , \it v3\rm )\rm )\rm )\rm ;

\hspace*{0.20in} $\forall$  \it bv\rm . \rm (\it bv  $\in$  \it BIT\_VECTOR  $\implies$  \rm (\it bv\_or\rm (\it bv\rm , \it bv\_one\rm (\it bv\_size\rm (\it bv\rm )\rm )\rm ) \rm = \it bv\_one\rm (\it bv\_size\rm (\it bv\rm )\rm )\rm )\rm )\rm ;

\hspace*{0.20in} $\forall$  \it bv\rm . \rm (\it bv  $\in$  \it BIT\_VECTOR  $\implies$  \rm (\it bv\_or\rm (\it bv\rm , \it bv\_zero\rm (\it bv\_size\rm (\it bv\rm )\rm )\rm ) \rm = \it bv\rm )\rm )\rm ;

\vspace*{4mm}
\hspace*{0.20in} $\forall$  \rm (\it v1\rm ,\it v2\rm )\rm . \rm (\it v1  $\in$  \it BIT\_VECTOR  $\land$  \it v2  $\in$  \it BIT\_VECTOR  $\land$  \it bv\_size\rm (\it v1\rm ) \rm = \it bv\_size\rm (\it v2\rm )  $\implies$

\hspace*{0.60in}\it bv\_size\rm (\it bv\_xor\rm (\it v1\rm , \it v2\rm )\rm ) \rm = \it bv\_size\rm (\it v1\rm )\rm )\rm ;

\hspace*{0.20in} $\forall$  \rm (\it v1\rm ,\it v2\rm )\rm . \rm (\it v1  $\in$  \it BIT\_VECTOR  $\land$  \it v2  $\in$  \it BIT\_VECTOR  $\land$  \it bv\_size\rm (\it v1\rm ) \rm = \it bv\_size\rm (\it v2\rm )  $\implies$

\hspace*{0.60in}\it bv\_size\rm (\it bv\_xor\rm (\it v1\rm , \it v2\rm )\rm ) \rm = \it bv\_size\rm (\it v2\rm )\rm )\rm ;

\hspace*{0.20in} $\forall$  \rm (\it v1\rm ,\it v2\rm )\rm . \rm (\it v1  $\in$  \it BIT\_VECTOR  $\land$  \it v2  $\in$  \it BIT\_VECTOR  $\land$  \it bv\_size\rm (\it v1\rm ) \rm = \it bv\_size\rm (\it v2\rm )  $\implies$

\hspace*{0.60in}\rm (\it bv\_xor\rm (\it v1\rm , \it v2\rm ) \rm = \it bv\_xor\rm (\it v2\rm , \it v1\rm )\rm )\rm )\rm ;

\hspace*{0.20in} $\forall$  \rm (\it bv\rm )\rm . \rm (\it bv  $\in$  \it BIT\_VECTOR  $\implies$  \it bv\_xor\rm (\it bv\rm ,\it bv\rm ) \rm = \it bv\_zero\rm (\it bv\_size\rm (\it bv\rm )\rm )\rm )

\bf END\newpage
\end{sloppypar}\end{document}