### 2008-11-12 trunk/developments/platforms/PIC16C432/PDB/DATA_MEMORY.tex

\documentstyle[11pt,supertabular]{article}
\setlength{\topmargin}{13mm}
\setlength{\headheight}{0mm}
\setlength{\headsep}{0mm}
\setlength{\textheight}{225mm}
\setlength{\oddsidemargin}{0mm}
\setlength{\textwidth}{160mm}
\input{psfig}

\begin{document}
\begin{sloppypar}

% DEFINITION DES CARACTERES MATHEMATIQUES B
%------------------------------------------
\def\@setmcodes#1#2#3{{\count0=#1 \count1=#3
	\loop \global\mathcode\count0=\count1 \ifnum \count0<#2
	\advance\count0 by1 \advance\count1 by1 \repeat}}

\@setmcodes{`A}{`Z}{"7441}
\@setmcodes{`a}{`z}{"7461}

\mathcode`\;="8000 % Makes ; active in math mode
{\catcode`\;=\active \gdef;{\semicolon\;}}
\mathchardef\semicolon="003B
%    Nominal distance from top of paper to top of page
\topmargin 0 pt
\textheight 53\baselineskip

%   Left margin on odd-numbered pages
\oddsidemargin  0.15 in
%   Left margin on even-numbered pages
\evensidemargin 0.35 in
%   Width of marginal notes.
\marginparwidth 1 in
%   Note that \oddsidemargin = \evensidemargin
\oddsidemargin 0.25 in
\evensidemargin 0.25 in
\marginparwidth 0.75 in
\textwidth 5.875 in % Width of text line.

\setlength{\parindent}{0pt}
\setlength{\parskip}{0ex}

% DEFINITION DES FONTS
%---------------------
% The AMS extra symbol fonts are loaded.
% Note: sometimes called euxm10
\font\msx=msam10
% Note: sometimes called euym10
\font\msy=msbm10

\newfam\msxfam \textfont\msxfam=\msx
\newfam\msyfam \textfont\msyfam=\msy

\def\famletter#1{\ifcase #1 0\or 1\or 2\or 3\or 4\or 5\or 6\or 7\or
	8\or 9\or A\or B\or C\or D\or E\or F\fi}

\edef\fx{\famletter\msxfam}
\edef\fy{\famletter\msyfam}

\def\bbold{\fam\msyfam \msy}

% SYMBOLES B
%-----------
% makes a quoted expression in mathematical text
\def\token#1{\hbox{`$#1$'}}
% used for error messages in Z specs
\def\report#1{\hbox{`{\tt #1}'}}

% \@myop makes an operator, with a strut to defeat TeX's vertical adjustment.
\def\@myop#1{\mathop{\mathstrut{#1}}\nolimits}

% This underscore doesn't have the little kern --- you get an italic
% correction anyway in math mode.
\def\_{\leavevmode \vbox{\hrule width0.5em}}

% Save \q as \xq for quantifiers q.
\let\xforall=\forall
\let\xexists=\exists
\let\xlambda=\lambda
\let\xmu=\mu

% \p and \f make arrows with 1 and 2 crossings resp.
\def\p#1{\mathrel{\ooalign{\hfil$\mapstochar\mkern 5mu$\hfil\cr$#1$}}}
\def\f#1{\mathrel{\ooalign{\hfil
	$\mapstochar\mkern 3mu\mapstochar\mkern 5mu$\hfil\cr$#1$}}}

\let\mc=\mathchardef

\def	\pow		{\mbox{${\cal P}$}}
\def	\po1		{\mbox{${\cal P}_1$}}
\let	\cross		\times
\def	\lambda		{\@myop{\xlambda}}
\def	\lnot		{\neg\;}
\def	\land		{\mathrel{\wedge}}
\def	\lor		{\mathrel{\vee}}
\let	\implies	\Rightarrow
\let	\iff		\Leftrightarrow
\def	\forall		{\@myop{\xforall}}
\def	\exists		{\@myop{\xexists}}
\def	\semi		{\mathrel{\comp}}
\def	\ssemi		{\mathbin{\rm ;}}
\let	\ensembleVide	\emptyset
\let	\rel		\leftrightarrow
\def	\dom		{\@myop{\sf dom}}
\def	\ran		{\@myop{\sf ran}}
\def	\id		{\@myop{\sf id}}
\def	\comp		{\mathbin{\raise
			0.6ex\hbox{\oalign{\hfil$\scriptscriptstyle
			\rm o$\hfil\cr\hfil$\scriptscriptstyle\rm 9$\hfil}}}}
\def	\para		{\mbox{$\mid\mid$}}
\mc	\dres		"2\fx43
\mc	\rres		"2\fx42
\def	\ndres		{\mathbin{{\dres} \llap{$-$}}}
\def	\nrres		{\mathbin{{\rres}\llap{$-$}}}
\def	\lover		{\mathbin{{\dres} \llap{$-\!\!\!\!-\!$}}}
\def	\rover		{\mathbin{{\rres}\llap{$\!-\!\!\!-$}}}
\let	\fun		\rightarrow
\def	\pfun		{\p\fun}
\def	\pinj		{\p\inj}
\mc	\inj		"3\fx1A
\def	\psurj		{\p\surj}
\mc	\surj		"3\fx10
\def	\bij		{\surj\!\!\!\!\!\!\!\inj}
\def	\nat		{\mbox{${\cal N}$}}
\def	\na1		{\mbox{${\cal N}_1$}}
\def	\num		{\mbox{${\cal Z}$}}
\def	\int		{\mbox{${\cal Z}$}}
\def	\rat		{\mbox{${\cal Q}$}}
\def	\div		{\mathbin{\rm /}}
\def	\mod		{\mathbin{\bf mod}}
\def	\upto		{\mathbin{\ldotp\ldotp}}
\def	\finset		{\mbox{${\cal F}$}}
\def	\finse1		{\mbox{${\cal F}_1$}}
\def	\ffun		{\f\fun}
\def	\finj		{\f\inj}
\def	\seq		{\@myop{\rm seq}}
\def	\cat		{\mathbin{\raise 0.8ex\hbox{$\mathchar"2\fx61$}}}
\def	\sep		{\hspace*{.05in}}

\setcounter{secnumdepth}{0}
\setcounter{tocdepth}{0}

%-------------------%
% Debut du document %
%-------------------%

\vspace*{4mm}
\bf MACHINE

\hspace*{0.20in}\it DATA\_MEMORY

\vspace*{4mm}
\bf SEES

\hspace*{0.20in}\it BIT\_VECTOR\_DEFINITION\rm ,

\hspace*{0.20in}\it BIT\_VECTOR\_ARITHMETICS\rm ,

\hspace*{0.20in}\it BYTE\_DEFINITION\rm ,

\vspace*{4mm}
\hspace*{0.20in}\it POWER2

\vspace*{4mm}
\bf CONSTANTS

\hspace*{0.20in}\it BANK\_WIDTH\rm , \it BANK\rm , \it BANK\_SIZE\rm ,

\hspace*{0.20in}\it LOCATION\_WIDTH\rm , \it LOCATION\rm ,

\hspace*{0.20in}\it ADDRESS\_WIDTH\rm , \it ADDRESS\rm ,

\hspace*{0.20in}\it FILE\rm ,

\hspace*{0.20in}\it make\_address\rm ,

\vspace*{4mm}
\hspace*{0.20in}

\hspace*{0.20in}\it INDF\_ADDR\rm , \it TMR0\_ADDR\rm , \it PCL\_ADDR\rm ,

\hspace*{0.20in}\it STATUS\_ADDR\rm , \it FSR\_ADDR\rm ,

\hspace*{0.20in}\it PORTA\_ADDR\rm , \it PORTB\_ADDR\rm ,

\hspace*{0.20in}\it PCLATH\_ADDR\rm , \it INTCON\_ADDR\rm , \it PIR1\_ADDR\rm , \it CMCON\_ADDR\rm ,

\hspace*{0.20in}\it TRISA\_ADDR\rm , \it TRISB\_ADDR\rm ,

\hspace*{0.20in}\it PIE1\_ADDR\rm , \it PCON\_ADDR\rm , \it LININTF\_ADDR\rm , \it VRCON\_ADDR\rm ,

\vspace*{4mm}
\hspace*{0.20in}

\hspace*{0.20in}\it UNIMPLEMENTED\_LOCATIONS\rm ,

\vspace*{4mm}
\hspace*{0.20in}\it MAPPED\_REGISTERS\rm ,

\hspace*{0.20in}\it register\_addresses\rm ,

\vspace*{4mm}
\hspace*{0.20in}\it IRP\_POS\rm , \it RP1\_POS\rm , \it RP0\_POS\rm ,

\vspace*{4mm}
\hspace*{0.20in}\it actual\_address

\vspace*{4mm}
\bf PROPERTIES

\hspace*{0.20in}\it BANK\_WIDTH \rm = \rm 2  $\land$

\hspace*{0.20in}\it BANK  $\subseteq$  \it BIT\_VECTOR  $\land$

\hspace*{0.20in}\it BANK \rm = \rm \{ \it vv  $\mid$  \it vv  $\in$  \it BIT\_VECTOR  $\land$  \it bv\_size\rm (\it vv\rm ) \rm = \it BANK\_WIDTH \rm \}  $\land$

\hspace*{0.20in}\it LOCATION\_WIDTH \rm = \rm 7  $\land$

\hspace*{0.20in}\it LOCATION  $\subseteq$  \it BIT\_VECTOR  $\land$

\hspace*{0.20in}\it LOCATION \rm = \rm \{ \it vv  $\mid$  \it vv  $\in$  \it BIT\_VECTOR  $\land$  \it bv\_size\rm (\it vv\rm ) \rm = \it LOCATION\_WIDTH \rm \}  $\land$

\hspace*{0.20in}\it BANK\_SIZE \rm = \rm 2$^{LOCATION\_WIDTH}$  $\land$

\hspace*{0.20in}\it ADDRESS\_WIDTH \rm = \it BANK\_WIDTH \rm + \it LOCATION\_WIDTH  $\land$

\hspace*{0.20in}\it ADDRESS \rm = \rm 0 $\upto$ \rm (\rm (\rm 2$^{ADDRESS\_WIDTH}$\rm )\rm -\rm 1\rm )  $\land$

\hspace*{0.20in}\it make\_address  $\in$  \it BANK  $\times$  \it LOCATION  $\fun$  \it ADDRESS  $\land$

\hspace*{0.20in}\it make\_address \rm =  $\lambda$ \rm (\it vb\rm , \it vl\rm )\rm . \rm (\it vb  $\in$  \it BANK  $\land$  \it vl  $\in$  \it LOCATION  $\mid$  \it bv\_to\_nat \rm (\it bv\_catenate\rm (\it vb\rm , \it vl\rm )\rm )\rm )  $\land$

\vspace*{4mm}
\hspace*{0.20in}\it FILE \rm = \it ADDRESS  $\fun$  \it BYTE  $\land$

\vspace*{4mm}
\hspace*{0.20in}

\hspace*{0.20in}\it INDF\_ADDR  $\in$  \it ADDRESS  $\land$ \hspace*{0.25in}\it INDF\_ADDR \rm = \rm 0  $\land$

\hspace*{0.20in}\it TMR0\_ADDR  $\in$  \it ADDRESS  $\land$ \hspace*{0.25in}\it TMR0\_ADDR \rm = \rm 1  $\land$

\hspace*{0.20in}\it PCL\_ADDR  $\in$  \it ADDRESS  $\land$ \hspace*{0.25in}\it PCL\_ADDR \rm = \rm 2  $\land$

\hspace*{0.20in}\it STATUS\_ADDR  $\in$  \it ADDRESS  $\land$  \it STATUS\_ADDR \rm = \rm 3  $\land$

\hspace*{0.20in}\it FSR\_ADDR  $\in$  \it ADDRESS  $\land$ \hspace*{0.25in}\it FSR\_ADDR \rm = \rm 4  $\land$

\hspace*{0.20in}\it PORTA\_ADDR  $\in$  \it ADDRESS  $\land$ \hspace*{0.25in}\it PORTA\_ADDR \rm = \rm 5  $\land$

\hspace*{0.20in}\it PORTB\_ADDR  $\in$  \it ADDRESS  $\land$ \hspace*{0.25in}\it PORTB\_ADDR \rm = \rm 6  $\land$

\hspace*{0.20in}\it PCLATH\_ADDR  $\in$  \it ADDRESS  $\land$  \it PCLATH\_ADDR \rm = \rm 1\rm 0  $\land$

\hspace*{0.20in}\it INTCON\_ADDR  $\in$  \it ADDRESS  $\land$  \it INTCON\_ADDR \rm = \rm 1\rm 1  $\land$

\hspace*{0.20in}\it PIR1\_ADDR  $\in$  \it ADDRESS  $\land$ \hspace*{0.25in}\it PIR1\_ADDR \rm = \rm 1\rm 2  $\land$

\hspace*{0.20in}\it CMCON\_ADDR  $\in$  \it ADDRESS  $\land$ \hspace*{0.25in}\it CMCON\_ADDR \rm = \rm 1\rm 2\rm 7  $\land$

\hspace*{0.20in}\it TRISA\_ADDR  $\in$  \it ADDRESS  $\land$ \hspace*{0.25in}\it TRISA\_ADDR \rm = \rm 1\rm 3\rm 3  $\land$

\hspace*{0.20in}\it TRISB\_ADDR  $\in$  \it ADDRESS  $\land$ \hspace*{0.25in}\it TRISB\_ADDR \rm = \rm 1\rm 3\rm 4  $\land$

\hspace*{0.20in}\it PIE1\_ADDR  $\in$  \it ADDRESS  $\land$ \hspace*{0.25in}\it PIE1\_ADDR \rm = \rm 1\rm 4\rm 0  $\land$

\hspace*{0.20in}\it PCON\_ADDR  $\in$  \it ADDRESS  $\land$ \hspace*{0.25in}\it PCON\_ADDR \rm = \rm 1\rm 4\rm 2  $\land$

\hspace*{0.20in}\it LININTF\_ADDR  $\in$  \it ADDRESS  $\land$ \it LININTF\_ADDR \rm = \rm 1\rm 4\rm 4  $\land$

\hspace*{0.20in}\it VRCON\_ADDR  $\in$  \it ADDRESS  $\land$ \hspace*{0.25in}\it VRCON\_ADDR \rm = \rm 1\rm 5\rm 9  $\land$

\vspace*{4mm}
\hspace*{0.20in}

\hspace*{0.20in}\it UNIMPLEMENTED\_LOCATIONS \rm =

\hspace*{0.40in}\rm 7 $\upto$ \rm 9  $\cup$ \hspace*{0.20in}\rm 1\rm 3 $\upto$ \rm 3\rm 0  $\cup$  \rm 1\rm 3\rm 5 $\upto$ \rm 1\rm 3\rm 7  $\cup$  \rm \{\rm 1\rm 4\rm 1\rm \}  $\cup$  \rm \{\rm 1\rm 4\rm 3\rm \}  $\cup$  \rm 1\rm 4\rm 5 $\upto$ \rm 1\rm 5\rm 8  $\cup$  \rm 1\rm 9\rm 2 $\upto$ \rm 2\rm 4\rm 0  $\land$

\vspace*{4mm}
\hspace*{0.20in}\it MAPPED\_REGISTERS  $\subseteq$  \it ADDRESS  $\land$

\hspace*{0.20in}\it MAPPED\_REGISTERS \rm = \rm \{ \it INDF\_ADDR\rm , \it STATUS\_ADDR\rm , \it FSR\_ADDR\rm , \it PCLATH\_ADDR\rm , \it INTCON\_ADDR \rm \}  $\land$

\vspace*{4mm}
\hspace*{0.20in}\it register\_addresses  $\in$  \it ADDRESS  $\fun$   $ \po1 $  \rm (\it ADDRESS\rm )  $\land$

\hspace*{0.20in} $\forall$ \rm (\it ad\rm )\rm . \rm (\it ad  $\in$  \it ADDRESS  $\implies$

\hspace*{0.40in}\rm (\rm (\it ad  $\not\in$  \it MAPPED\_REGISTERS  $\implies$  \it register\_addresses\rm (\it ad\rm ) \rm = \rm \{\it ad\rm \}\rm )  $\land$

\hspace*{0.45in}\rm (\it ad  $\in$  \it MAPPED\_REGISTERS  $\implies$  \it register\_addresses\rm (\it ad\rm ) \rm = \rm \{\it ad\rm , \it ad \rm + \it BANK\_SIZE\rm \}\rm )\rm )\rm )  $\land$

\vspace*{4mm}
\hspace*{0.20in}\it IRP\_POS  $\in$  \it BYTE\_INDEX  $\land$  \it IRP\_POS \rm = \rm 7  $\land$

\hspace*{0.20in}\it RP1\_POS  $\in$  \it BYTE\_INDEX  $\land$  \it RP1\_POS \rm = \rm 6  $\land$

\hspace*{0.20in}\it RP0\_POS  $\in$  \it BYTE\_INDEX  $\land$  \it RP0\_POS \rm = \rm 5  $\land$

\vspace*{4mm}
\hspace*{0.20in}\it actual\_address  $\in$  \it ADDRESS  $\times$  \it FILE  $\fun$  \it ADDRESS  $\land$

\hspace*{0.20in}\it actual\_address \rm =

\hspace*{0.40in} $\lambda$  \rm (\it ad\rm , \it fi\rm ) \rm . \rm (\it ad  $\in$  \it ADDRESS  $\land$  \it fi  $\in$  \it FILE  $\land$  \it ad  $\in$  \it register\_addresses \rm (\it INDF\_ADDR\rm )  $\mid$

\hspace*{0.60in}\rm (\it fi\rm (\it STATUS\_ADDR\rm )\rm (\it RP0\_POS\rm ) $\times$ \rm 2$^{7}$\rm ) \rm + \rm (\it fi\rm (\it STATUS\_ADDR\rm )\rm (\it RP1\_POS\rm ) $\times$ \rm 2$^{8}$\rm ) \rm + \it bv\_to\_nat\rm (\it fi\rm (\it FSR\_ADDR\rm )\rm )\rm )  $\cup$

\hspace*{0.40in} $\lambda$  \rm (\it ad\rm , \it fi\rm ) \rm . \rm (\it ad  $\in$  \it ADDRESS  $\land$  \it fi  $\in$  \it FILE  $\land$  \it ad  $\not\in$  \it register\_addresses \rm (\it INDF\_ADDR\rm )  $\mid$

\hspace*{0.60in}\rm (\it fi\rm (\it STATUS\_ADDR\rm )\rm (\it IRP\_POS\rm ) $\times$ \rm 2$^{8}$\rm ) \rm + \it bv\_to\_nat \rm (\it fi\rm (\it FSR\_ADDR\rm )\rm )\rm )

\vspace*{4mm}
\bf VARIABLES

\hspace*{0.20in}\it file

\vspace*{4mm}
\bf INVARIANT

\hspace*{0.20in}\it file  $\in$  \it FILE  $\land$

\hspace*{0.20in} $\forall$ \rm (\it ad\rm )\rm .\rm (\it ad  $\in$  \it UNIMPLEMENTED\_LOCATIONS  $\implies$  \it file\rm (\it ad\rm ) \rm = \it BYTE\_ZERO\rm )  $\land$

\hspace*{0.20in} $\forall$ \rm (\it reg\rm )\rm .\rm (\it reg  $\in$  \it MAPPED\_REGISTERS  $\implies$  \it file\rm (\it reg\rm ) \rm = \it file\rm (\it reg\rm +\it BANK\_SIZE\rm )\rm )  $\land$

\hspace*{0.20in}\it file\rm (\it STATUS\_ADDR\rm )\rm (\it IRP\_POS\rm ) \rm = \rm 0  $\land$

\hspace*{0.20in}\it file\rm (\it STATUS\_ADDR\rm )\rm (\it RP1\_POS\rm ) \rm = \rm 0

\vspace*{4mm}
\bf ASSERTIONS

\hspace*{0.20in}\it BANK\_SIZE \rm = \rm 1\rm 2\rm 8\rm ;

\hspace*{0.20in}\it register\_addresses \rm (\it INDF\_ADDR\rm ) \rm = \rm \{\rm 0\rm , \rm 1\rm 2\rm 8\rm \}\rm ;

\hspace*{0.20in}\it register\_addresses \rm (\it STATUS\_ADDR\rm ) \rm = \rm \{\rm 3\rm , \rm 1\rm 3\rm 1\rm \}\rm ;

\hspace*{0.20in}\it register\_addresses \rm (\it FSR\_ADDR\rm ) \rm = \rm \{\rm 4\rm , \rm 1\rm 3\rm 2\rm \}\rm ;

\hspace*{0.20in}\it register\_addresses \rm (\it PCLATH\_ADDR\rm ) \rm = \rm \{\rm 1\rm 0\rm , \rm 1\rm 3\rm 8\rm \}\rm ;

\hspace*{0.20in}\it register\_addresses \rm (\it INTCON\_ADDR\rm ) \rm = \rm \{\rm 1\rm 1\rm , \rm 1\rm 3\rm 9\rm \}\rm ;

\vspace*{4mm}
\hspace*{0.20in}\bf dom\rm (\it file\rm ) \rm = \it ADDRESS\rm ;

\hspace*{0.20in}\bf ran\rm (\it file\rm )  $\subseteq$  \it BYTE

\vspace*{4mm}
\bf INITIALISATION

\hspace*{0.20in}\bf ANY \it random\rm , \it init\_status \bf WHERE

\hspace*{0.40in}\it random  $\in$  \it FILE  $\land$

\hspace*{0.40in}\it init\_status  $\in$  \it BYTE  $\land$

\hspace*{0.40in}\it init\_status \rm (\it RP1\_POS\rm ) \rm = \rm 0

\hspace*{0.20in}\bf THEN

\hspace*{0.40in}\it file \rm := \it random  $\lover$  \rm \{ \it INDF\_ADDR  $\mapsto$  \it BYTE\_ZERO\rm ,

\hspace*{1.00in}\it STATUS\_ADDR  $\mapsto$  \it init\_status\rm ,

\hspace*{1.00in}\it INDF\_ADDR \rm + \rm 1\rm 2\rm 8  $\mapsto$  \it BYTE\_ZERO\rm ,

\hspace*{1.00in}\it STATUS\_ADDR \rm + \rm 1\rm 2\rm 8  $\mapsto$  \it init\_status \rm \}

\hspace*{0.80in} $\lover$  \rm (\it UNIMPLEMENTED\_LOCATIONS  $\times$  \rm \{ \it BYTE\_ZERO \rm \}\rm )

\hspace*{0.20in}\bf END

\bf END

\newpage
\end{sloppypar}\end{document}