### 2009-04-20 developments/platforms/Z80/tex/BV16.tex

\bf MACHINE

\hspace*{0.20in}\it BV16\_DEFINITION\hspace*{0.20in}

\bf SEES

\hspace*{0.20in}\it BIT\_DEFINITION\rm ,

\hspace*{0.20in}\it BIT\_VECTOR\_DEFINITION\rm ,

\hspace*{0.20in}\it BYTE\_DEFINITION

\hspace*{0.20in}

\bf DEFINITIONS\hspace*{0.20in}

\hspace*{0.20in}\it BV16\_WIDTH \rm == \rm 1\rm 6

\hspace*{0.20in}

\bf CONSTANTS

\hspace*{0.20in}\it BV16\_INDX\rm ,

\hspace*{0.20in}\it PHYS\_BV16\_INDEX\rm ,

\hspace*{0.20in}\it BV16\_ZERO\rm ,

\hspace*{0.20in}\it BV16\rm ,

\hspace*{0.20in}\it bv16\_byte\rm ,

\hspace*{0.20in}\it byte\_bv16\rm ,

\hspace*{0.20in}\it bv16\_bit\_get

\hspace*{0.20in}

\bf PROPERTIES

\vspace*{4mm}
\hspace*{0.20in}\it BV16\_INDX \rm = \rm 1 $\upto$ \rm (\it BV16\_WIDTH\rm )  $\land$

\hspace*{0.20in}\it PHYS\_BV16\_INDEX \rm = \rm 0 $\upto$ \rm (\it BV16\_WIDTH\rm -\rm 1\rm )  $\land$

\hspace*{0.20in}\it BV16 \rm = \rm \{ \it bt  $\mid$  \it bt $\in$  \it BIT\_VECTOR  $\land$  \it bv\_size\rm (\it bt\rm )\rm =\it BV16\_WIDTH\rm \}  $\land$

\hspace*{0.20in}\it BV16\_ZERO \rm = \it BV16\_INDX $\times$ \rm \{\rm 0\rm \}  $\land$

\hspace*{0.20in}

\hspace*{0.20in}\it bv16\_byte $\in$  \it BV16  $\fun$  \it BYTE  $\times$  \it BYTE  $\land$

\hspace*{0.20in}\it bv16\_byte \rm =  $\lambda$ \rm (\it bv\rm )\rm .\rm (\it bv $\in$ \it BV16  $\mid$  \rm (\rm \{ \rm 1 $\mapsto$ \it bv\rm (\rm 9\rm )\rm ,\rm 2 $\mapsto$ \it bv\rm (\rm 1\rm 0\rm )\rm ,\rm 3 $\mapsto$ \it bv\rm (\rm 1\rm 1\rm )\rm ,\rm 4 $\mapsto$ \it bv\rm (\rm 1\rm 2\rm )\rm ,\rm 5 $\mapsto$ \it bv\rm (\rm 1\rm 3\rm )\rm ,\rm 6 $\mapsto$ \it bv\rm (\rm 1\rm 4\rm )\rm ,\rm 7 $\mapsto$ \it bv\rm (\rm 1\rm 5\rm )\rm ,\rm 8 $\mapsto$ \it bv\rm (\rm 1\rm 6\rm )\rm \}

\hspace*{1.60in}\rm ,\rm \{ \rm 1 $\mapsto$ \it bv\rm (\rm 1\rm )\rm ,\rm 2 $\mapsto$ \it bv\rm (\rm 2\rm )\rm ,\rm 3 $\mapsto$ \it bv\rm (\rm 3\rm )\rm ,\rm 4 $\mapsto$ \it bv\rm (\rm 4\rm )\rm ,\rm 5 $\mapsto$ \it bv\rm (\rm 5\rm )\rm ,\rm 6 $\mapsto$ \it bv\rm (\rm 6\rm )\rm ,\rm 7 $\mapsto$ \it bv\rm (\rm 7\rm )\rm ,\rm 8 $\mapsto$ \it bv\rm (\rm 8\rm )\rm \}\rm ) \rm )  $\land$

\hspace*{0.20in}\it byte\_bv16  $\in$ \hspace*{0.10in}\it BYTE  $\times$  \it BYTE  $\fun$  \it BV16  $\land$ \hspace*{0.80in}

\hspace*{0.25in}\it byte\_bv16 \rm = \it bv16\_byte $^{-1}$   $\land$

\vspace*{4mm}
\hspace*{0.20in}\it bv16\_bit\_get\rm : \rm (\it BV16  $\times$  \it PHYS\_BV16\_INDEX \rm ) $\fun$  \it BIT  $\land$

\hspace*{0.20in}\it bv16\_bit\_get \rm =  $\lambda$ \rm (\it bt\rm ,\it ind\rm )\rm .\rm (\it bt $\in$ \it BV16  $\land$  \it ind $\in$ \it PHYS\_BV16\_INDEX  $\mid$  \it bt\rm (\it ind\rm +\rm 1\rm ) \rm )\hspace*{0.20in}

\vspace*{4mm}
\bf ASSERTIONS

\hspace*{0.20in}

\hspace*{0.20in}\it BV16\_ZERO $\in$  \it BV16  $\land$

\hspace*{0.20in}\it BV16 $\subseteq$ \it BIT\_VECTOR  $\land$

\hspace*{0.20in}

\hspace*{0.20in} $\forall$ \rm (\it bt\rm )\rm .\rm (\it bt  $\in$  \it BV16  $\implies$  \it bv\_size\rm (\it bt\rm )\rm =\bf size\rm (\it bt\rm )  $\land$  \it bv\_size\rm (\it bt\rm )\rm =\rm 1\rm 6  $\land$  \it bt  $\in$  \bf seq1\rm (\it BIT\rm )\hspace*{0.10in}\rm )  $\land$

\vspace*{4mm}
\hspace*{0.20in}\bf dom\rm (\it bv16\_byte\rm ) \rm = \it BV16  $\land$

\hspace*{0.20in}\bf ran\rm (\it bv16\_byte\rm ) \rm = \rm (\it BYTE  $\times$  \it BYTE\rm )  $\land$

\vspace*{4mm}
\hspace*{0.20in}\bf dom\rm (\it byte\_bv16\rm ) \rm = \rm (\it BYTE  $\times$  \it BYTE\rm )\hspace*{0.10in} $\land$

\hspace*{0.20in}\bf ran\rm (\it byte\_bv16\rm ) \rm = \it BV16  $\land$

\hspace*{0.20in}

\hspace*{0.20in}\bf dom\rm (\it bv16\_bit\_get\rm ) \rm = \rm (\it BV16  $\times$  \it PHYS\_BV16\_INDEX \rm )  $\land$

\hspace*{0.20in}\bf ran\rm (\it bv16\_bit\_get\rm ) \rm = \it BIT  $\land$

\hspace*{0.20in}

\hspace*{0.20in}\rm \rm [\rm 1\rm ,\rm 1\rm ,\rm 1\rm ,\rm 1\rm ,\rm 1\rm ,\rm 1\rm ,\rm 1\rm ,\rm 1\rm ,\rm 1\rm ,\rm 1\rm ,\rm 1\rm ,\rm 1\rm ,\rm 1\rm ,\rm 1\rm ,\rm 1\rm ,\rm 1\rm \rm ]  $\in$  \it BV16  $\land$

\hspace*{0.20in}\rm \rm [\rm 0\rm ,\rm 0\rm ,\rm 0\rm ,\rm 0\rm ,\rm 0\rm ,\rm 0\rm ,\rm 0\rm ,\rm 0\rm ,\rm 0\rm ,\rm 0\rm ,\rm 0\rm ,\rm 0\rm ,\rm 0\rm ,\rm 0\rm ,\rm 0\rm ,\rm 0\rm \rm ]  $\in$  \it BV16

\hspace*{0.20in}

\bf END

\newpage
\end{sloppypar}\end{document}