### 2009-04-20 developments/platforms/Z80/tex/BYTE.tex

\bf MACHINE

\hspace*{0.20in}\it BYTE\_DEFINITION

\hspace*{0.20in}

\bf SEES

\hspace*{0.20in}\it BIT\_DEFINITION\rm ,

\hspace*{0.20in}\it BIT\_VECTOR\_DEFINITION\rm ,

\hspace*{0.20in}\it POWER2

\hspace*{0.20in}

\bf DEFINITIONS

\hspace*{0.20in}\it BYTE\_WIDTH \rm == \rm 8 \rm ;

\hspace*{0.20in}

\hspace*{0.20in}\it pow2\_0 \rm == \rm 1\rm ;

\hspace*{0.20in}\it pow2\_1 \rm == \rm 2\rm ;

\hspace*{0.20in}\it pow2\_2 \rm == \rm 4\rm ;

\hspace*{0.20in}\it pow2\_3 \rm == \rm 8\rm ;

\hspace*{0.20in}\it pow2\_4 \rm == \rm 1\rm 6\rm ;

\hspace*{0.20in}\it pow2\_5 \rm == \rm 3\rm 2\rm ;

\hspace*{0.20in}\it pow2\_6 \rm == \rm 6\rm 4\rm ;

\hspace*{0.20in}\it pow2\_7 \rm == \rm 1\rm 2\rm 8\rm ;

\hspace*{0.20in}\it pow2\_8 \rm == \rm 2\rm 5\rm 6\rm ;

\hspace*{0.20in}\it pow2\_9 \rm == \rm 5\rm 1\rm 2\rm ;

\hspace*{0.20in}\it pow2\_10 \rm == \rm 1\rm 0\rm 2\rm 4\rm ;\hspace*{0.20in}

\hspace*{0.20in}\it pow2\_11 \rm == \rm 2\rm 0\rm 4\rm 8\rm ;\hspace*{0.10in}

\hspace*{0.20in}\it pow2\_12 \rm == \rm 4\rm 0\rm 9\rm 6\rm ;\hspace*{0.15in}

\hspace*{0.20in}\it pow2\_13 \rm == \rm 8\rm 1\rm 9\rm 2\rm ;

\hspace*{0.20in}\it pow2\_14 \rm == \rm 1\rm 6\rm 3\rm 8\rm 4\rm ;\hspace*{0.15in}

\hspace*{0.20in}\it pow2\_15 \rm == \rm 3\rm 2\rm 7\rm 6\rm 8\rm ;

\hspace*{0.20in}\it pow2\_16 \rm == \rm 6\rm 5\rm 5\rm 3\rm 6\hspace*{0.35in}

\hspace*{0.20in}

\bf CONSTANTS

\hspace*{0.20in}

\hspace*{0.20in}\it BYTE\rm ,

\hspace*{0.20in}\it BYTE\_INDEX\rm ,

\hspace*{0.20in}\it BYTE\_ZERO\rm ,

\hspace*{0.20in}\it PHYS\_BYTE\_INDEX\rm ,

\hspace*{0.20in}

\hspace*{0.20in}

\hspace*{0.20in}\it is\_zero\rm ,

\hspace*{0.20in}\it parity\_even\rm ,

\hspace*{0.20in}\it and \rm , \it ior \rm , \it xor \rm ,

\hspace*{0.20in}\it bitclear \rm , \it bitset \rm , \it bitget \rm ,

\hspace*{0.20in}\it complement\rm , \it swap \rm , \it rotateleft \rm , \it rotateright\rm ,\hspace*{0.15in}

\hspace*{0.20in}\it get\_upper\_digit \rm ,

\hspace*{0.20in}\it get\_lower\_digit

\hspace*{0.20in}

\bf PROPERTIES

\hspace*{0.20in}

\hspace*{0.20in}\it BYTE\_INDEX \rm = \rm 1 $\upto$  \rm (\it BYTE\_WIDTH\rm )  $\land$

\hspace*{0.20in}\it PHYS\_BYTE\_INDEX \rm = \rm 0 $\upto$ \rm (\it BYTE\_WIDTH\rm -\rm 1\rm )\hspace*{0.10in} $\land$

\hspace*{0.20in}\it BYTE \rm = \rm \{ \it bt  $\mid$  \it bt $\in$ \it BIT\_VECTOR  $\land$  \it bv\_size\rm (\it bt\rm )\rm =\it BYTE\_WIDTH\rm \}\hspace*{0.10in} $\land$

\hspace*{0.20in}\it BYTE\_ZERO  $\in$  \it BYTE  $\land$

\hspace*{0.20in}\it BYTE\_ZERO \rm = \it BYTE\_INDEX  $\times$  \rm \{\rm 0\rm \}  $\land$

\hspace*{0.20in}

\hspace*{0.20in}

\hspace*{0.20in}\it is\_zero  $\in$  \it BYTE  $\fun$  \it BIT  $\land$

\hspace*{0.20in}\it is\_zero \rm =  $\lambda$ \rm (\it w1\rm )\rm .\rm (\it w1  $\in$  \it BYTE  $\mid$  \it bool\_bit\rm ( \bf bool\rm (\it w1\rm (\rm 1\rm ) \rm + \it w1\rm (\rm 2\rm )\rm + \it w1\rm (\rm 3\rm ) \rm + \it w1\rm (\rm 4\rm ) \rm + \it w1\rm (\rm 5\rm ) \rm + \it w1\rm (\rm 6\rm ) \rm + \it w1\rm (\rm 7\rm ) \rm + \it w1\rm (\rm 8\rm ) \rm = \rm 0\rm ) \rm )\rm )  $\land$

\vspace*{4mm}
\hspace*{0.20in}

\hspace*{0.20in}\it parity\_even  $\in$  \it BYTE  $\fun$  \it BIT\hspace*{0.20in} $\land$

\hspace*{0.20in}\it parity\_even \rm =  $\lambda$  \rm (\it bv\rm )\rm . \rm (\it bv  $\in$  \it BYTE  $\mid$  \rm 1\rm -\rm ( \it bv\rm (\rm 1\rm )\rm + \it bv\rm (\rm 2\rm )\rm + \it bv\rm (\rm 3\rm )\rm +\it bv\rm (\rm 4\rm )\rm +\it bv\rm (\rm 5\rm )\rm + \it bv\rm (\rm 6\rm )\rm +\it bv\rm (\rm 7\rm )\rm +\it bv\rm (\rm 8\rm )  $\mod$  \rm 2\rm )\rm )  $\land$

\hspace*{0.20in}

\hspace*{0.20in}\it and \rm : \rm (\it BYTE  $\times$  \it BYTE\rm )  $\fun$  \it BYTE  $\land$

\hspace*{0.20in}\it and \rm =  $\lambda$ \rm (\it bt1\rm , \it bt2\rm )\rm .\rm (\it bt1 $\in$ \it BYTE  $\land$  \it bt2 $\in$ \it BYTE  $\mid$  \it bv\_and\rm (\it bt1\rm ,\it bt2\rm )\rm )  $\land$

\hspace*{0.20in}

\hspace*{0.20in}\it ior \rm : \rm (\it BYTE  $\times$  \it BYTE\rm )  $\fun$  \it BYTE  $\land$

\hspace*{0.20in}\it ior \rm =  $\lambda$ \rm (\it bt1\rm , \it bt2\rm )\rm .\rm (\it bt1 $\in$ \it BYTE  $\land$  \it bt2 $\in$ \it BYTE  $\mid$  \it bv\_or\rm (\it bt1\rm ,\it bt2\rm )\rm )\hspace*{0.15in} $\land$

\hspace*{0.20in}

\hspace*{0.20in}\it xor \rm : \rm (\it BYTE  $\times$  \it BYTE\rm )  $\fun$  \it BYTE  $\land$

\hspace*{0.20in}\it xor \rm =  $\lambda$ \rm (\it bt1\rm , \it bt2\rm )\rm .\rm (\it bt1  $\in$  \it BYTE  $\land$  \it bt2  $\in$  \it BYTE  $\mid$  \it bv\_xor\rm (\it bt1\rm ,\it bt2\rm )\rm )  $\land$

\hspace*{0.20in}

\hspace*{0.20in}

\hspace*{0.20in}\it bitget \rm : \rm (\it BYTE  $\times$  \it PHYS\_BYTE\_INDEX\rm )  $\fun$  \it BIT  $\land$

\hspace*{0.20in}\it bitget \rm =  $\lambda$ \rm (\it bt1\rm , \it ii\rm )\rm .\rm (\it bt1  $\in$  \it BYTE  $\land$  \it ii  $\in$  \it PHYS\_BYTE\_INDEX  $\mid$  \it bt1\rm (\it ii\rm +\rm 1\rm )\rm )  $\land$

\hspace*{0.20in}

\hspace*{0.20in}\it bitset \rm : \rm (\it BYTE  $\times$  \it PHYS\_BYTE\_INDEX\rm )  $\fun$  \it BYTE  $\land$

\hspace*{0.20in}\it bitset \rm =  $\lambda$ \rm (\it bt1\rm , \it ii\rm )\rm .\rm (\it bt1  $\in$  \it BYTE  $\land$  \it ii  $\in$  \it PHYS\_BYTE\_INDEX  $\mid$  \it bv\_set\rm (\it bt1\rm , \it ii\rm )\rm )  $\land$

\hspace*{0.20in}

\hspace*{0.20in}\it bitclear \rm : \rm (\it BYTE  $\times$  \it PHYS\_BYTE\_INDEX\rm )  $\fun$  \it BYTE  $\land$

\hspace*{0.20in} $\forall$  \rm (\it ww\rm , \it ii\rm , \it bb\rm )\rm .\rm (\it ww  $\in$  \it BYTE  $\land$  \it ii  $\in$  \it PHYS\_BYTE\_INDEX  $\land$  \it bb  $\in$  \it BIT  $\implies$  \it bitclear\rm (\it ww\rm ,\it ii\rm ) \rm = \it bv\_clear\rm (\it ww\rm , \it ii\rm )\rm )  $\land$

\hspace*{0.20in}

\hspace*{0.20in}\it complement  $\in$  \it BYTE  $\fun$  \it BYTE  $\land$

\hspace*{0.20in}\it complement \rm =  $\lambda$ \rm (\it bt\rm )\rm .\rm (\it bt  $\in$  \it BYTE  $\mid$  \it bv\_not\rm (\it bt\rm )\rm )  $\land$

\hspace*{0.20in}

\hspace*{0.20in}

\hspace*{0.20in}\it swap  $\in$  \it BYTE  $\fun$  \it BYTE  $\land$

\hspace*{0.20in}\it swap\rm = $\lambda$ \rm (\it bt\rm )\rm .\rm (\it bt $\in$ \it BYTE  $\mid$ \rm \{\rm 1 $\mapsto$ \it bt\rm (\rm 5\rm )\rm , \rm 2 $\mapsto$ \it bt\rm (\rm 6\rm )\rm , \rm 3 $\mapsto$ \it bt\rm (\rm 7\rm )\rm , \rm 4  $\mapsto$ \it bt\rm (\rm 8\rm )\rm , \rm 5 $\mapsto$ \it bt\rm (\rm 1\rm )\rm , \rm 6 $\mapsto$ \it bt\rm (\rm 2\rm )\rm , \rm 7 $\mapsto$ \it bt\rm (\rm 3\rm )\rm , \rm 8 $\mapsto$ \it bt\rm (\rm 4\rm )\rm \}\rm )  $\land$

\hspace*{0.20in}

\hspace*{0.20in}

\hspace*{0.20in}\it rotateleft  $\in$  \it BYTE  $\fun$  \it BYTE\hspace*{0.10in} $\land$

\hspace*{0.20in}\it rotateleft\hspace*{0.10in}\rm =  $\lambda$ \rm (\it bv\rm )\rm .\rm (\it bv $\in$ \it BYTE $\mid$  \rm \{\rm 1 $\mapsto$ \it bv\rm (\rm 8\rm )\rm , \rm 2 $\mapsto$ \it bv\rm (\rm 1\rm )\rm , \rm 3 $\mapsto$ \it bv\rm (\rm 2\rm )\rm , \rm 4 $\mapsto$ \it bv\rm (\rm 3\rm )\rm , \rm 5 $\mapsto$ \it bv\rm (\rm 4\rm )\rm ,\rm 6 $\mapsto$ \it bv\rm (\rm 5\rm )\rm ,\rm 7 $\mapsto$ \it bv\rm (\rm 6\rm )\rm ,\rm 8 $\mapsto$ \it bv\rm (\rm 7\rm )\rm \} \rm )  $\land$

\hspace*{0.20in}

\hspace*{0.20in}

\hspace*{0.20in}\it rotateright  $\in$  \it BYTE  $\fun$  \it BYTE\hspace*{0.10in} $\land$

\hspace*{0.20in}\it rotateright\rm =  $\lambda$ \rm (\it bv\rm )\rm .\rm (\it bv $\in$ \it BYTE $\mid$ \rm \{\rm 1 $\mapsto$ \it bv\rm (\rm 2\rm )\rm , \rm 2 $\mapsto$ \it bv\rm (\rm 3\rm )\rm , \rm 3 $\mapsto$ \it bv\rm (\rm 4\rm )\rm , \rm 4 $\mapsto$ \it bv\rm (\rm 5\rm )\rm ,\rm 5 $\mapsto$ \it bv\rm (\rm 6\rm )\rm , \rm 6 $\mapsto$ \it bv\rm (\rm 7\rm )\rm , \rm 7 $\mapsto$ \it bv\rm (\rm 8\rm )\rm , \rm 8 $\mapsto$ \it bv\rm (\rm 1\rm )\rm \}\rm )\hspace*{0.10in} $\land$

\hspace*{0.20in}

\hspace*{0.20in}

\hspace*{0.20in}\it get\_upper\_digit \rm = $\lambda$ \rm ( \it by \rm )\rm . \rm ( \it by  $\in$  \it BYTE  $\mid$ \hspace*{0.10in}\it pow2\_3  $\times$  \it by \rm (\rm 8 \rm ) \rm + \it pow2\_2  $\times$  \it by \rm ( \rm 7 \rm ) \rm + \rm 2  $\times$  \it by \rm ( \rm 6 \rm ) \rm +\hspace*{0.10in}\it by \rm ( \rm 5 \rm )\hspace*{0.15in}\rm )  $\land$

\hspace*{0.20in}\it get\_lower\_digit \rm = $\lambda$ \rm ( \it by \rm )\rm . \rm ( \it by  $\in$  \it BYTE  $\mid$ \hspace*{0.10in}\it pow2\_3  $\times$  \it by \rm (\rm 4 \rm ) \rm + \it pow2\_2  $\times$  \it by \rm ( \rm 3 \rm ) \rm + \rm 2  $\times$  \it by \rm ( \rm 2 \rm ) \rm +\hspace*{0.10in}\it by \rm ( \rm 1 \rm )\hspace*{0.15in}\rm )

\hspace*{0.20in}

\hspace*{0.20in}

\bf ASSERTIONS

\vspace*{4mm}
\hspace*{0.20in}\bf size\rm (\it BYTE\_ZERO\rm ) \rm = \rm 8  $\land$

\hspace*{0.20in}\it BYTE  $\subseteq$  \it BIT\_VECTOR  $\land$

\hspace*{0.20in}\it BYTE\_ZERO  $\in$  \it BIT\_VECTOR  $\land$

\hspace*{0.20in}\bf first\rm (\it BYTE\_ZERO\rm )\rm =\rm 0  $\land$

\hspace*{0.20in}

\hspace*{0.20in} $\forall$ \rm (\it bt\rm )\rm .\rm (\it bt  $\in$  \it BYTE  $\implies$  \it bv\_size\rm (\it bt\rm )\rm =\bf size\rm (\it bt\rm )  $\land$  \it bv\_size\rm (\it bt\rm )\rm =\rm 8  $\land$  \it bt  $\in$  \bf seq1\rm (\it BIT\rm )\hspace*{0.10in}\rm )  $\land$

\hspace*{0.20in}

\hspace*{0.20in}\it BYTE\_WIDTH  $\in$   $\nat$   $\land$

\hspace*{0.40in}

\hspace*{0.20in}\bf dom\rm (\it is\_zero\rm )\rm = \it BYTE  $\land$

\hspace*{0.20in}\bf ran\rm (\it is\_zero\rm )\rm = \it BIT\hspace*{0.10in} $\land$

\hspace*{0.20in}

\hspace*{0.20in}\bf dom\rm (\it parity\_even\rm ) \rm = \it BYTE  $\land$

\hspace*{0.20in}\bf ran\rm (\it parity\_even\rm ) \rm = \it BIT  $\land$

\hspace*{0.20in}

\hspace*{0.20in}\bf dom\rm (\it and\rm ) \rm = \it BYTE  $\times$  \it BYTE  $\land$

\hspace*{0.20in}\bf ran\rm (\it and\rm ) \rm = \it BYTE\hspace*{0.10in} $\land$

\hspace*{0.20in}

\hspace*{0.20in}\bf dom\rm (\it ior\rm ) \rm = \it BYTE  $\times$  \it BYTE  $\land$ \hspace*{0.20in}

\hspace*{0.20in}\bf ran\rm (\it ior\rm ) \rm = \it BYTE\hspace*{0.10in} $\land$

\hspace*{0.20in}

\hspace*{0.20in}\bf dom\rm (\it xor\rm ) \rm = \it BYTE  $\times$  \it BYTE  $\land$

\hspace*{0.20in}\bf ran\rm (\it xor\rm ) \rm = \it BYTE\hspace*{0.10in} $\land$

\hspace*{0.20in}

\hspace*{0.20in}\bf dom\rm (\it bitget\rm )\rm =\rm (\it BYTE  $\times$  \it PHYS\_BYTE\_INDEX\rm )  $\land$

\hspace*{0.20in}\bf ran\rm (\it bitget\rm )\rm =\it BIT  $\land$

\hspace*{0.20in}

\hspace*{0.20in}\bf dom\rm (\it bitset\rm )\rm =\rm (\it BYTE  $\times$  \it PHYS\_BYTE\_INDEX\rm )  $\land$

\hspace*{0.20in}\bf ran\rm (\it bitset\rm )\rm = \it BYTE  $\land$ \hspace*{0.15in}

\hspace*{0.20in}

\hspace*{0.20in}\bf dom\rm (\it bitclear\rm )\rm =\rm (\it BYTE  $\times$  \it PHYS\_BYTE\_INDEX\rm ) $\land$

\hspace*{0.20in}\bf ran\rm (\it bitclear\rm )\rm = \it BYTE  $\land$

\hspace*{0.20in}

\hspace*{0.20in}\bf dom\rm (\it complement\rm ) \rm = \it BYTE  $\land$

\hspace*{0.20in}\bf ran\rm (\it complement\rm ) \rm = \it BYTE  $\land$

\hspace*{0.20in}

\hspace*{0.20in}\bf dom\rm (\it swap\rm ) \rm = \it BYTE  $\land$

\hspace*{0.20in}\bf ran\rm (\it swap\rm ) \rm = \it BYTE  $\land$

\hspace*{0.20in}

\hspace*{0.20in}\bf dom\rm (\it rotateleft\rm ) \rm = \it BYTE  $\land$

\hspace*{0.20in}\bf ran\rm (\it rotateleft\rm ) \rm = \it BYTE  $\land$

\hspace*{0.20in}

\hspace*{0.20in}\bf dom\rm (\it rotateright\rm ) \rm = \it BYTE  $\land$

\hspace*{0.20in}\bf ran\rm (\it rotateright\rm ) \rm = \it BYTE  $\land$

\hspace*{0.20in}

\hspace*{0.20in}\bf dom\rm (\it get\_upper\_digit \rm ) \rm = \it BYTE  $\land$

\hspace*{0.20in}\bf ran\rm (\it get\_upper\_digit\rm ) \rm = \rm 0 $\upto$ \rm (\it pow2\_4 \rm - \rm 1\rm )  $\land$

\hspace*{0.20in}

\hspace*{0.20in}\bf dom\rm (\it get\_lower\_digit \rm ) \rm = \it BYTE  $\land$

\hspace*{0.20in}\bf ran\rm (\it get\_lower\_digit\rm ) \rm = \rm 0 $\upto$ \rm (\it pow2\_4 \rm - \rm 1\rm )  $\land$

\hspace*{0.20in}

\hspace*{0.20in}

\hspace*{0.20in}\rm \rm [\rm 1\rm ,\rm 1\rm ,\rm 1\rm ,\rm 1\rm ,\rm 1\rm ,\rm 1\rm ,\rm 1\rm ,\rm 1\rm \rm ]  $\in$  \it BYTE  $\land$ \hspace*{0.10in}

\hspace*{0.20in}\rm \rm [\rm 0\rm ,\rm 0\rm ,\rm 0\rm ,\rm 0\rm ,\rm 0\rm ,\rm 0\rm ,\rm 0\rm ,\rm 0\rm \rm ]  $\in$  \it BYTE  $\land$

\hspace*{0.20in}

\hspace*{0.20in}

\hspace*{0.20in}\it pow2\_0 \rm = \rm 2$^{0}$  $\land$

\hspace*{0.20in}\it pow2\_1 \rm = \rm 2$^{1}$  $\land$

\hspace*{0.20in}\it pow2\_2 \rm = \rm 2$^{2}$  $\land$

\hspace*{0.20in}\it pow2\_3 \rm = \rm 2$^{3}$  $\land$

\hspace*{0.20in}\it pow2\_4 \rm = \rm 2$^{4}$  $\land$

\hspace*{0.20in}\it pow2\_5 \rm = \rm 2$^{5}$  $\land$

\hspace*{0.20in}\it pow2\_6 \rm = \rm 2$^{6}$  $\land$

\hspace*{0.20in}\it pow2\_7 \rm = \rm 2$^{7}$  $\land$

\hspace*{0.20in}\it pow2\_8 \rm = \rm 2$^{8}$  $\land$

\hspace*{0.20in}\it pow2\_9 \rm = \rm 2$^{9}$  $\land$ \hspace*{0.20in}

\hspace*{0.20in}\it pow2\_10 \rm = \rm 2$^{10}$  $\land$

\hspace*{0.20in}\it pow2\_11 \rm = \rm 2$^{11}$  $\land$

\hspace*{0.20in}\it pow2\_12 \rm = \rm 2$^{12}$  $\land$

\hspace*{0.20in}\it pow2\_13 \rm = \rm 2$^{13}$  $\land$

\hspace*{0.20in}\it pow2\_14 \rm = \rm 2$^{14}$  $\land$

\hspace*{0.20in}\it pow2\_15 \rm = \rm 2$^{15}$  $\land$

\hspace*{0.20in}\it pow2\_16 \rm = \rm 2$^{16}$

\hspace*{0.20in}

\hspace*{0.20in}

\hspace*{0.20in}

\hspace*{0.20in}

\bf END

\newpage
\end{sloppypar}\end{document}