### 2009-04-20 developments/platforms/Z80/tex/MEMORY.tex

\bf MACHINE

\hspace*{0.20in}\it MEMORY

\hspace*{0.20in}

\bf SEES

\hspace*{0.20in}\it ALU\rm ,

\hspace*{0.20in}\it BIT\_DEFINITION \rm ,

\hspace*{0.20in}\it BIT\_VECTOR\_DEFINITION\rm ,

\hspace*{0.20in}\it BYTE\_DEFINITION\rm ,

\hspace*{0.20in}\it BV16\_DEFINITION\rm ,

\hspace*{0.20in}

\hspace*{0.20in}\it UCHAR\_DEFINITION\rm ,

\hspace*{0.20in}\it SCHAR\_DEFINITION\rm ,

\hspace*{0.20in}\it SSHORT\_DEFINITION\rm ,

\hspace*{0.20in}\it USHORT\_DEFINITION\rm ,

\hspace*{0.20in}\it POWER2

\bf CONSTANTS

\hspace*{0.20in}\it PROGRAM\_R\_ADR \rm ,

\hspace*{0.20in}\it DATA\_R\_ADR \rm ,

\hspace*{0.20in}\it STACK\_R\_ADR

\hspace*{0.20in}

\bf PROPERTIES

\hspace*{0.20in}

\hspace*{0.20in}

\hspace*{0.20in}\it PROGRAM\_R\_ADR \rm = \it USHORT  $\land$

\hspace*{0.20in}\it DATA\_R\_ADR \rm = \it USHORT  $\land$

\hspace*{0.20in}\it STACK\_R\_ADR \rm = \it USHORT

\hspace*{0.20in}

\hspace*{0.20in}

\hspace*{0.20in}

\hspace*{0.20in}

\bf ABSTRACT\_VARIABLES

\hspace*{0.20in}

\hspace*{0.20in}\it mem

\hspace*{0.20in}

\bf INVARIANT

\hspace*{0.20in}

\hspace*{0.20in}\it mem  $\in$  \it BV16  $\fun$  \it BYTE

\hspace*{0.20in}

\hspace*{0.20in}

\bf ASSERTIONS

\hspace*{0.20in}

\hspace*{0.20in}\bf ran\rm (\it mem\rm )  $\subseteq$  \it BYTE  $\land$

\hspace*{0.20in}\bf dom\rm (\it mem\rm ) \rm = \it BV16  $\land$

\hspace*{0.20in} $\forall$ \rm (\it address\rm ,\it value\rm )\rm .\rm ( \it address  $\in$  \it BV16  $\land$  \it value  $\in$  \it BYTE  $\implies$ \hspace*{0.20in}\it mem  $\lover$  \rm \{ \it address  $\mapsto$  \it value \rm \}  $\in$  \it BV16  $\fun$  \it BYTE \rm )

\hspace*{0.20in}

\hspace*{0.20in}

\hspace*{0.20in}

\hspace*{0.20in}

\bf INITIALISATION

\hspace*{0.20in}\it mem  $:\in$  \it BV16  $\fun$  \it BYTE\hspace*{0.15in}

\hspace*{0.20in}

\bf OPERATIONS

\hspace*{0.20in}

\hspace*{0.20in}\bf updateMem\rm (\it NewValues\rm ) \rm =

\hspace*{0.20in}\bf PRE \it NewValues  $\in$  \it BV16  $\fun$  \it BYTE\hspace*{0.10in} $\land$   $\forall$ \rm (\it elem\rm )\rm .\rm ( \it elem  $\in$  \bf dom\rm (\it NewValues\rm )  $\implies$  \it bv16\_ushort\rm (\it elem\rm )  $\in$  \it PROGRAM\_R\_ADR\rm )

\hspace*{0.20in}\bf THEN

\hspace*{0.40in}\it mem\rm := \it mem  $\lover$  \it NewValues

\hspace*{0.20in}\bf END\rm ;

\hspace*{0.20in}

\hspace*{0.20in}\bf updateAddressMem\rm (\it address\rm , \it value\rm ) \rm =

\hspace*{0.20in}\bf PRE\hspace*{0.15in}\it address  $\in$  \it BV16  $\land$  \it value  $\in$  \it BYTE  $\land$  \it bv16\_ushort\rm (\it address\rm )  $\in$  \it PROGRAM\_R\_ADR

\hspace*{0.20in}\bf THEN

\hspace*{0.40in}\it mem\rm := \it mem  $\lover$  \rm \{ \it address  $\mapsto$  \it value \rm \}

\hspace*{0.20in}\bf END\rm ;

\hspace*{0.20in}

\hspace*{0.20in}\bf updateStack\rm (\it NewValues\rm ) \rm =

\hspace*{0.20in}\bf PRE \it NewValues  $\in$  \it BV16  $\fun$  \it BYTE\hspace*{0.10in} $\land$ \hspace*{0.10in} $\forall$ \rm (\it elem\rm )\rm .\rm ( \it elem  $\in$  \bf dom\rm (\it NewValues\rm )  $\implies$  \it bv16\_ushort\rm (\it elem\rm )  $\in$  \it STACK\_R\_ADR\rm )

\hspace*{0.20in}\bf THEN

\hspace*{0.40in}\it mem\rm := \it mem  $\lover$  \it NewValues

\hspace*{0.20in}\bf END\rm ;

\hspace*{0.20in}

\hspace*{0.20in}\bf updateAddressStack\rm (\it address\rm , \it value\rm ) \rm =

\hspace*{0.20in}\bf PRE \it address  $\in$  \it BV16  $\land$  \it value  $\in$  \it BYTE  $\land$ \hspace*{0.10in}\it bv16\_ushort\rm (\it address\rm )  $\in$  \it STACK\_R\_ADR\hspace*{0.15in}

\hspace*{0.20in}\bf THEN

\hspace*{0.40in}\it mem\rm := \it mem  $\lover$  \rm \{ \it address  $\mapsto$  \it value \rm \}

\hspace*{0.20in}\bf END

\hspace*{0.20in}

\bf END\newpage
\end{sloppypar}\end{document}