### 2009-04-20 trunk/developments/platforms/Z80/tex/BIT_VECTOR.tex

\bf MACHINE

\hspace*{0.20in}\it BIT\_VECTOR\_DEFINITION

\vspace*{4mm}
\bf SEES

\hspace*{0.20in}\it BIT\_DEFINITION

\vspace*{4mm}
\bf CONSTANTS

\hspace*{0.15in}

\hspace*{0.20in}\it BIT\_VECTOR\rm ,

\hspace*{0.20in}\it bv\_catenate\rm ,

\hspace*{0.20in}\it bv\_sub\rm ,

\hspace*{0.20in}\it bv\_zero\rm ,

\hspace*{0.20in}\it bv\_one\rm ,

\hspace*{0.20in}\it bv\_size\rm ,

\hspace*{0.20in}\it bv\_not\rm ,\hspace*{0.20in}

\hspace*{0.20in}\it bv\_and\rm ,\hspace*{0.20in}

\hspace*{0.20in}\it bv\_or\rm ,

\hspace*{0.20in}\it bv\_xor\rm ,

\hspace*{0.20in}\it bv\_get\rm ,

\hspace*{0.20in}\it bv\_set\rm ,

\hspace*{0.20in}\it bv\_clear\rm ,

\hspace*{0.20in}\it bv\_put\rm ,

\hspace*{0.20in}\it phys\_bv\_index

\vspace*{4mm}
\bf PROPERTIES

\hspace*{0.20in}\it BIT\_VECTOR \rm = \bf seq1\rm (\it BIT\rm )

\hspace*{0.15in}

\hspace*{0.20in} $\land$  \it bv\_size  $\in$  \it BIT\_VECTOR  $\fun$   $ \na1 $

\hspace*{0.20in} $\land$  \it bv\_size \rm =  $\lambda$ \rm (\it bv\rm )\rm .\rm (\it bv  $\in$  \it BIT\_VECTOR  $\mid$  \bf size\rm (\it bv\rm )\rm )

\hspace*{0.20in}

\hspace*{0.20in} $\land$  \it bv\_catenate  $\in$  \it BIT\_VECTOR  $\times$  \it BIT\_VECTOR  $\fun$  \it BIT\_VECTOR

\hspace*{0.20in} $\land$  \it bv\_catenate \rm =  $\lambda$ \rm (\it v1\rm ,\it v2\rm ) \rm . \rm (\it v1  $\in$  \it BIT\_VECTOR  $\land$  \it v2 $\in$  \it BIT\_VECTOR  $\mid$  \it v1 $\cat$ \it v2\rm )

\vspace*{4mm}
\hspace*{0.20in} $\land$  \it phys\_bv\_index  $\in$  \it BIT\_VECTOR  $\fun$   $\pow$ \rm ( $\nat$ \rm )

\hspace*{0.20in} $\land$  \it phys\_bv\_index \rm =  $\lambda$ \rm (\it v1\rm ) \rm . \rm (\it v1  $\in$  \it BIT\_VECTOR  $\mid$  \rm 0 $\upto$ \rm (\it bv\_size\rm (\it v1\rm )\rm -\rm 1\rm )\rm )

\hspace*{0.15in}

\hspace*{0.20in} $\land$  \it bv\_sub  $\in$  \it BIT\_VECTOR  $\times$   $\nat$   $\times$   $\nat$   $\fun$  \it BIT\_VECTOR

\hspace*{0.20in} $\land$  \it bv\_sub \rm =  $\lambda$ \rm (\it bv\rm ,\it low\rm ,\it high\rm ) \rm . \rm (\it bv  $\in$  \it BIT\_VECTOR  $\land$  \it low $\in$  \rm 0 $\upto$ \rm (\it bv\_size\rm (\it bv\rm )\rm -\rm 1\rm )  $\land$  \it high $\in$ \it low $\upto$ \rm (\it bv\_size\rm (\it bv\rm )\rm -\rm 1\rm )

\hspace*{1.80in} $\mid$   $\lambda$ \rm (\it i0\rm )\rm .\rm (\it i0 $\in$  \rm 1 $\upto$ \rm (\it high\rm -\it low\rm +\rm 1\rm )  $\mid$  \it bv\rm (\it i0\rm +\it low\rm ) \rm )\rm )

\vspace*{4mm}
\hspace*{0.20in} $\land$  \it bv\_zero  $\in$   $ \na1 $   $\fun$  \it BIT\_VECTOR

\hspace*{0.20in} $\land$  \it bv\_zero \rm =  $\lambda$ \rm (\it sz\rm ) \rm . \rm (\it sz  $\in$   $ \na1 $   $\mid$  \rm (\rm 1 $\upto$ \it sz\rm )  $\times$  \rm \{\rm 0\rm \}\rm )

\vspace*{4mm}
\hspace*{0.20in} $\land$  \it bv\_one  $\in$   $ \na1 $   $\fun$  \it BIT\_VECTOR

\hspace*{0.20in} $\land$  \it bv\_one \rm =  $\lambda$ \rm (\it sz\rm ) \rm . \rm (\it sz  $\in$   $ \na1 $   $\mid$  \rm (\rm 1 $\upto$ \it sz\rm )  $\times$  \rm \{\rm 1\rm \}\rm )

\vspace*{4mm}
\hspace*{0.20in} $\land$  \it bv\_not  $\in$  \it BIT\_VECTOR  $\fun$  \it BIT\_VECTOR

\hspace*{0.20in} $\land$  \it bv\_not \rm =  $\lambda$ \rm (\it v1\rm )\rm .\rm (\it v1 $\in$  \it BIT\_VECTOR  $\mid$   $\lambda$ \rm (\it idx\rm )\rm .\rm (\it idx  $\in$  \rm 1 $\upto$ \it bv\_size\rm (\it v1\rm ) $\mid$ \it bit\_not\rm (\it v1\rm (\it idx\rm ) \rm ) \rm ) \rm )

\vspace*{4mm}
\hspace*{0.20in} $\land$  \it bv\_and  $\in$  \it BIT\_VECTOR  $\times$  \it BIT\_VECTOR  $\fun$  \it BIT\_VECTOR

\hspace*{0.20in} $\land$  \it bv\_and \rm =  $\lambda$ \rm (\it v1\rm ,\it v2\rm )\rm .\rm (\it v1 $\in$  \it BIT\_VECTOR  $\land$  \it v2  $\in$  \it BIT\_VECTOR  $\land$  \it bv\_size\rm (\it v1\rm ) \rm = \it bv\_size\rm (\it v2\rm ) $\mid$

\hspace*{0.80in} $\lambda$ \rm (\it idx\rm )\rm .\rm (\it idx  $\in$  \rm 1 $\upto$ \it bv\_size\rm (\it v1\rm )  $\mid$  \it bit\_and\rm (\it v1\rm (\it idx\rm )\rm ,\it v2\rm (\it idx\rm )\rm )\rm )\rm )

\vspace*{4mm}
\hspace*{0.20in} $\land$  \it bv\_or  $\in$  \it BIT\_VECTOR  $\times$  \it BIT\_VECTOR  $\fun$  \it BIT\_VECTOR

\hspace*{0.20in} $\land$  \it bv\_or \rm =  $\lambda$ \rm (\it v1\rm ,\it v2\rm )\rm .\rm (\it v1 $\in$  \it BIT\_VECTOR  $\land$  \it v2  $\in$  \it BIT\_VECTOR  $\land$  \it bv\_size\rm (\it v1\rm ) \rm = \it bv\_size\rm (\it v2\rm )  $\mid$

\hspace*{0.80in} $\lambda$ \rm (\it idx\rm )\rm .\rm (\it idx  $\in$  \rm 1 $\upto$ \it bv\_size\rm (\it v1\rm )  $\mid$  \it bit\_or\rm (\it v1\rm (\it idx\rm )\rm ,\it v2\rm (\it idx\rm )\rm )\rm )\rm )

\vspace*{4mm}
\hspace*{0.20in} $\land$  \it bv\_xor  $\in$  \it BIT\_VECTOR  $\times$  \it BIT\_VECTOR  $\fun$  \it BIT\_VECTOR

\hspace*{0.20in} $\land$  \it bv\_xor \rm =  $\lambda$ \rm (\it v1\rm ,\it v2\rm )\rm .\rm (\it v1 $\in$  \it BIT\_VECTOR  $\land$  \it v2  $\in$  \it BIT\_VECTOR  $\land$  \it bv\_size\rm (\it v1\rm ) \rm = \it bv\_size\rm (\it v2\rm )  $\mid$

\hspace*{0.40in} $\lambda$ \rm (\it idx\rm )\rm .\rm (\it idx  $\in$ \rm 1 $\upto$ \it bv\_size\rm (\it v1\rm )  $\mid$  \it bit\_xor\rm (\it v1\rm (\it idx\rm )\rm ,\it v2\rm (\it idx\rm )\rm )\rm )\rm )

\hspace*{0.20in}

\hspace*{0.20in} $\land$  \it bv\_clear  $\in$  \it BIT\_VECTOR  $\times$   $\nat$   $\fun$  \it BIT\_VECTOR

\hspace*{0.20in} $\land$  \it bv\_clear \rm =  $\lambda$ \rm (\it v1\rm ,\it idx\rm )\rm .\rm (\it v1 $\in$  \it BIT\_VECTOR  $\land$  \it idx  $\in$  \rm 0 $\upto$ \rm (\it bv\_size\rm (\it v1\rm )\rm -\rm 1\rm )  $\mid$  \it v1  $\lover$  \rm \{ \rm (\it idx\rm +\rm 1\rm )  $\mapsto$  \rm 0 \rm \}\rm )

\hspace*{0.30in}

\hspace*{0.20in} $\land$  \it bv\_set  $\in$  \it BIT\_VECTOR  $\times$   $\nat$   $\fun$  \it BIT\_VECTOR

\hspace*{0.20in} $\land$  \it bv\_set \rm =  $\lambda$ \rm (\it v1\rm ,\it idx\rm )\rm .\rm (\it v1 $\in$  \it BIT\_VECTOR  $\land$  \it idx  $\in$  \rm 0 $\upto$ \rm (\it bv\_size\rm (\it v1\rm )\rm -\rm 1\rm )  $\mid$  \it v1  $\lover$  \rm \{ \rm (\it idx\rm +\rm 1\rm )  $\mapsto$  \rm 1 \rm \}\rm )

\vspace*{4mm}
\hspace*{0.20in} $\land$  \it bv\_get  $\in$  \it BIT\_VECTOR  $\times$   $\nat$   $\fun$  \it BIT

\hspace*{0.20in} $\land$  \it bv\_get \rm =  $\lambda$ \rm (\it v1\rm ,\it idx\rm )\rm .\rm (\it v1 $\in$  \it BIT\_VECTOR  $\land$  \it idx  $\in$  \rm 0 $\upto$ \rm (\it bv\_size\rm (\it v1\rm )\rm -\rm 1\rm )  $\mid$  \it v1\rm (\it idx\rm +\rm 1\rm )\rm )

\vspace*{8mm}
\hspace*{0.20in} $\land$  \it bv\_put $\in$  \it BIT\_VECTOR  $\times$   $\nat$   $\times$ \it BIT $\fun$  \it BIT\_VECTOR

\hspace*{0.20in} $\land$  \it bv\_put \rm =  $\lambda$ \rm (\it v1\rm ,\it idx\rm ,\it bit\rm )\rm .\rm (\it v1 $\in$  \it BIT\_VECTOR  $\land$  \it idx  $\in$  \rm 0 $\upto$ \rm (\it bv\_size\rm (\it v1\rm )\rm -\rm 1\rm )  $\land$  \it bit $\in$ \it BIT  $\mid$  \it v1  $\lover$  \rm \{ \rm (\it idx\rm +\rm 1\rm ) $\mapsto$  \it bit \rm \}\rm )

\hspace*{0.15in}

\hspace*{0.15in}

\vspace*{4mm}
\bf ASSERTIONS

\hspace*{0.20in} $\forall$  \it bv\rm .\rm (\it bv  $\in$  \it BIT\_VECTOR  $\implies$  \it bv\_size\rm (\it bv\_not\rm (\it bv\rm )\rm ) \rm = \it bv\_size\rm (\it bv\rm )\rm )\rm ;

\hspace*{0.55in}

\hspace*{0.20in} $\forall$  \rm (\it bv\rm ,\it indx\rm )\rm .\rm (\it bv  $\in$  \it BIT\_VECTOR  $\land$  \it indx  $\in$  \rm 0 $\upto$ \rm (\it bv\_size\rm (\it bv\rm )\rm -\rm 1\rm )  $\implies$  \rm (\it bv\_get\rm (\it bv\_not\rm (\it bv\_not\rm (\it bv\rm )\rm )\rm ,\it indx\rm ) \rm = \it bv\_get\rm (\it bv\rm ,\it indx\rm ) \rm )\rm )\rm ;

\vspace*{4mm}
\hspace*{0.20in} $\forall$  \rm (\it v1\rm ,\it v2\rm )\rm . \rm (\it v1  $\in$  \it BIT\_VECTOR  $\land$  \it v2  $\in$  \it BIT\_VECTOR  $\implies$

\hspace*{1.00in}\it bv\_size\rm (\it bv\_catenate\rm (\it v1\rm , \it v2\rm )\rm ) \rm = \it bv\_size\rm (\it v1\rm ) \rm + \it bv\_size\rm (\it v2\rm )\rm )\rm ;

\vspace*{4mm}
\hspace*{0.20in} $\forall$  \rm (\it bv\rm ,\it low\rm ,\it high\rm )\rm . \rm (\it bv  $\in$  \it BIT\_VECTOR  $\land$  \it low  $\in$  \rm 0 $\upto$ \rm (\it bv\_size\rm (\it bv\rm )\rm -\rm 1\rm )  $\land$  \it high  $\in$  \rm 0 $\upto$ \rm (\it bv\_size\rm (\it bv\rm )\rm -\rm 1\rm )  $\land$  \it low  $\leq$  \it high  $\implies$

\hspace*{1.20in}\it bv\_size\rm (\it bv\_sub\rm (\it bv\rm , \it low\rm , \it high\rm )\rm ) \rm = \it high\rm -\it low\rm )\rm ;

\vspace*{4mm}
\hspace*{0.20in} $\forall$  \rm (\it v1\rm ,\it v2\rm )\rm . \rm (\it v1  $\in$  \it BIT\_VECTOR  $\land$  \it v2  $\in$  \it BIT\_VECTOR  $\land$  \it bv\_size\rm (\it v1\rm ) \rm = \it bv\_size\rm (\it v2\rm )  $\implies$

\hspace*{1.00in}\it bv\_size\rm (\it bv\_and\rm (\it v1\rm , \it v2\rm )\rm ) \rm = \it bv\_size\rm (\it v2\rm )\rm )\rm ;

\hspace*{0.35in}

\hspace*{0.20in} $\forall$  \rm (\it v1\rm ,\it v2\rm ,\it indx\rm )\rm . \rm (\it v1  $\in$  \it BIT\_VECTOR  $\land$  \it v2  $\in$  \it BIT\_VECTOR\hspace*{0.10in} $\land$  \it bv\_size\rm (\it v1\rm ) \rm = \it bv\_size\rm (\it v2\rm )  $\land$  \it indx $\in$  \rm 0 $\upto$ \rm (\it bv\_size\rm (\it v1\rm )\rm -\rm 1\rm ) $\implies$

\hspace*{1.20in}\rm (\it bv\_get\rm (\it bv\_and\rm (\it v1\rm , \it v2\rm )\rm ,\it indx\rm ) \rm = \it bv\_get\rm (\it bv\_and\rm (\it v2\rm , \it v1\rm )\rm ,\it indx\rm )\rm )\rm )\rm ;

\hspace*{0.35in}

\hspace*{0.20in} $\forall$  \rm (\it v1\rm ,\it v2\rm ,\it v3\rm ,\it indx\rm )\rm . \rm (\it v1 $\in$ \it BIT\_VECTOR  $\land$  \it v2 $\in$ \it BIT\_VECTOR  $\land$  \it v3 $\in$ \it BIT\_VECTOR  $\land$  \it bv\_size\rm (\it v1\rm )\rm =\it bv\_size\rm (\it v2\rm )  $\land$  \it bv\_size\rm (\it v1\rm )\rm =\it bv\_size\rm (\it v3\rm )

\hspace*{1.40in} $\land$  \it indx $\in$  \rm 0 $\upto$ \rm (\it bv\_size\rm (\it v1\rm )\rm -\rm 1\rm ) $\implies$  \rm (\it bv\_get\rm (\it bv\_and\rm (\it v1\rm , \it bv\_and\rm (\it v2\rm ,\it v3\rm )\rm )\rm ,\it indx\rm ) \rm = \it bv\_get\rm (\it bv\_and\rm (\it bv\_and\rm (\it v1\rm , \it v2\rm )\rm , \it v3\rm )\rm ,\it indx\rm )\rm ) \rm )\rm ;

\hspace*{0.35in}

\hspace*{0.20in} $\forall$  \rm (\it bv\rm ,\it indx\rm )\rm . \rm (\it bv  $\in$  \it BIT\_VECTOR  $\land$  \it indx $\in$ \rm 0 $\upto$ \rm (\it bv\_size\rm (\it bv\rm )\rm -\rm 1\rm )  $\implies$  \rm (\it bv\_get\rm (\it bv\_and\rm (\it bv\rm , \it bv\_zero\rm (\it bv\_size\rm (\it bv\rm )\rm )\rm )\rm ,\it indx\rm ) \rm =

\hspace*{1.00in}\it bv\_get\rm (\it bv\_zero\rm (\it bv\_size\rm (\it bv\rm )\rm )\rm ,\it indx\rm ) \rm )\rm )\rm ;

\vspace*{4mm}
\hspace*{0.20in} $\forall$  \rm (\it bv\rm ,\it indx\rm )\rm . \rm (\it bv  $\in$  \it BIT\_VECTOR  $\land$  \it indx $\in$ \rm 0 $\upto$ \rm (\it bv\_size\rm (\it bv\rm )\rm -\rm 1\rm )  $\implies$  \rm (\it bv\_get\rm (\it bv\_and\rm (\it bv\rm , \it bv\_one\rm (\it bv\_size\rm (\it bv\rm )\rm )\rm )\rm ,\it indx\rm ) \rm = \it bv\_get\rm (\it bv\rm ,\it indx\rm )\rm )\rm )\rm ;

\vspace*{4mm}
\hspace*{0.20in} $\forall$  \rm (\it v1\rm ,\it v2\rm )\rm . \rm (\it v1  $\in$  \it BIT\_VECTOR  $\land$  \it v2  $\in$  \it BIT\_VECTOR  $\land$  \it bv\_size\rm (\it v1\rm ) \rm = \it bv\_size\rm (\it v2\rm )  $\implies$  \it bv\_size\rm (\it bv\_or\rm (\it v1\rm , \it v2\rm )\rm ) \rm = \it bv\_size\rm (\it v1\rm )\rm )\rm ;

\hspace*{0.35in}

\hspace*{0.20in} $\forall$  \rm (\it v1\rm ,\it v2\rm ,\it indx\rm )\rm . \rm (\it v1  $\in$  \it BIT\_VECTOR  $\land$  \it v2  $\in$  \it BIT\_VECTOR  $\land$  \it bv\_size\rm (\it v1\rm ) \rm = \it bv\_size\rm (\it v2\rm )  $\land$  \it indx $\in$ \rm 0 $\upto$ \rm (\it bv\_size\rm (\it v1\rm )\rm -\rm 1\rm )  $\implies$

\hspace*{1.20in}\rm (\it bv\_get\rm (\it bv\_or\rm (\it v1\rm , \it v2\rm )\rm ,\it indx\rm ) \rm = \it bv\_get\rm (\it bv\_or\rm (\it v2\rm , \it v1\rm )\rm ,\it indx\rm )\rm )\rm )\rm ;

\hspace*{0.35in}

\hspace*{0.20in} $\forall$  \rm (\it v1\rm ,\it v2\rm )\rm . \rm (\it v1  $\in$  \it BIT\_VECTOR  $\land$  \it v2  $\in$  \it BIT\_VECTOR  $\land$  \it bv\_size\rm (\it v1\rm ) \rm = \it bv\_size\rm (\it v2\rm )  $\implies$

\hspace*{1.00in}\it bv\_size\rm (\it bv\_or\rm (\it v1\rm , \it v2\rm )\rm ) \rm = \it bv\_size\rm (\it v2\rm )\rm )\rm ;\hspace*{0.15in}

\hspace*{0.35in}

\hspace*{0.20in} $\forall$  \rm (\it v1\rm ,\it v2\rm ,\it v3\rm ,\it indx\rm )\rm . \rm (\it v1  $\in$  \it BIT\_VECTOR  $\land$  \it v2  $\in$  \it BIT\_VECTOR $\land$  \it v3  $\in$  \it BIT\_VECTOR  $\land$  \it bv\_size\rm (\it v1\rm ) \rm = \it bv\_size\rm (\it v2\rm )  $\land$  \it bv\_size\rm (\it v1\rm ) \rm = \it bv\_size\rm (\it v3\rm )

\hspace*{1.40in} $\land$  \it indx $\in$ \rm 0 $\upto$ \rm (\it bv\_size\rm (\it v1\rm )\rm -\rm 1\rm ) $\implies$  \rm (\it bv\_get\rm (\it bv\_or\rm (\it v1\rm , \it bv\_or\rm (\it v2\rm ,\it v3\rm )\rm )\rm ,\it indx\rm ) \rm = \it bv\_get\rm (\it bv\_or\rm (\it bv\_or\rm (\it v1\rm , \it v2\rm )\rm , \it v3\rm )\rm ,\it indx\rm )\rm )\rm )\rm ;

\hspace*{0.15in}

\hspace*{0.20in} $\forall$  \rm (\it bv\rm ,\it indx\rm )\rm . \rm (\it bv $\in$ \it BIT\_VECTOR  $\land$  \it indx $\in$ \rm 0 $\upto$ \rm (\it bv\_size\rm (\it bv\rm )\rm -\rm 1\rm )  $\implies$  \rm (\it bv\_get\rm (\it bv\_or\rm (\it bv\rm , \it bv\_one\rm (\it bv\_size\rm (\it bv\rm )\rm )\rm )\rm ,\it indx\rm ) \rm = \it bv\_get\rm (\it bv\_one\rm (\it bv\_size\rm (\it bv\rm )\rm )\rm ,\it indx\rm )\rm )\rm )\rm ;

\hspace*{0.20in}

\hspace*{0.20in} $\forall$  \rm (\it bv\rm ,\it indx\rm )\rm . \rm (\it bv $\in$ \it BIT\_VECTOR  $\land$  \it indx $\in$ \rm 0 $\upto$ \rm (\it bv\_size\rm (\it bv\rm )\rm -\rm 1\rm ) $\implies$  \rm (\it bv\_get\rm (\it bv\_or\rm (\it bv\rm , \it bv\_zero\rm (\it bv\_size\rm (\it bv\rm )\rm )\rm )\rm ,\it indx\rm ) \rm = \it bv\_get\rm (\it bv\rm ,\it indx\rm )\rm )\rm )\rm ;

\vspace*{4mm}
\hspace*{0.20in} $\forall$  \rm (\it v1\rm ,\it v2\rm )\rm . \rm (\it v1  $\in$  \it BIT\_VECTOR  $\land$  \it v2  $\in$  \it BIT\_VECTOR  $\land$  \it bv\_size\rm (\it v1\rm ) \rm = \it bv\_size\rm (\it v2\rm )  $\implies$ \it bv\_size\rm (\it bv\_xor\rm (\it v1\rm , \it v2\rm )\rm ) \rm = \it bv\_size\rm (\it v1\rm )\rm )\rm ;

\hspace*{0.20in}

\hspace*{0.20in} $\forall$  \rm (\it v1\rm ,\it v2\rm )\rm . \rm (\it v1  $\in$  \it BIT\_VECTOR  $\land$  \it v2  $\in$  \it BIT\_VECTOR  $\land$  \it bv\_size\rm (\it v1\rm ) \rm = \it bv\_size\rm (\it v2\rm )  $\implies$ \it bv\_size\rm (\it bv\_xor\rm (\it v1\rm , \it v2\rm )\rm ) \rm = \it bv\_size\rm (\it v2\rm )\rm )\rm ;

\hspace*{0.20in}

\hspace*{0.20in} $\forall$  \rm (\it v1\rm ,\it v2\rm ,\it indx\rm )\rm . \rm (\it v1  $\in$  \it BIT\_VECTOR  $\land$  \it v2  $\in$  \it BIT\_VECTOR  $\land$  \it bv\_size\rm (\it v1\rm ) \rm = \it bv\_size\rm (\it v2\rm )  $\land$  \it indx $\in$ \rm 0 $\upto$ \rm (\it bv\_size\rm (\it v1\rm )\rm -\rm 1\rm )  $\implies$

\hspace*{1.20in}\rm (\it bv\_get\rm (\it bv\_xor\rm (\it v1\rm , \it v2\rm )\rm ,\it indx\rm ) \rm = \it bv\_get\rm (\it bv\_xor\rm (\it v2\rm , \it v1\rm )\rm ,\it indx\rm )\rm )\rm )\rm ;

\hspace*{1.00in}

\hspace*{0.20in} $\forall$  \rm (\it bv\rm ,\it indx\rm )\rm . \rm (\it bv  $\in$  \it BIT\_VECTOR  $\land$  \it indx $\in$ \rm 0 $\upto$ \rm (\it bv\_size\rm (\it bv\rm )\rm -\rm 1\rm )  $\implies$  \it bv\_get\rm (\it bv\_xor\rm (\it bv\rm ,\it bv\rm )\rm ,\it indx\rm ) \rm = \it bv\_get\rm (\it bv\_zero\rm (\it bv\_size\rm (\it bv\rm )\rm )\rm ,\it indx\rm )\rm )

\bf END\newpage
\end{sloppypar}\end{document}