### 2009-04-22 papers/FM2009/source/paper.tex

\documentclass{llncs}

\usepackage{graphicx,array}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsfonts}

\include{comandos}
\include{SymbolsB_AtelierB}

%
\begin{document}

\title{Formal Verification of Real Models from Platforms Using The B-method }

\author{Val\'{e}rio Medeiros Jr\inst{1}, David D\'{e}harbe\inst{1} , Stephenson Galv\~{a}o \inst{1}}

\institute{Federal University of Rio Grande do Norte, Natal RN 59078-970, Brazil}

\maketitle

\begin{abstract}
This paper describes an approach to model the microcontroller platforms. More specifically, it shows details about
the Z80 model. The model is created using the B method; it applies math and logic concepts to describe the
characteristics from platforms. Therefore, this model can be used in platforms projects to document, build
simulators, verify properties about the model and verify the software in assembly level. The verification in
assembly is the feature more important, because this allow the developing software faithful to the algorithm
language. Finally, this paper show some relevant techniques used in models and to describe several details.
\end{abstract}
%
\section{Introduction}
% Explica o contexto geral do artigo
%  Porque modelar os microcontroladores. \footnote{The interested reader in details from full specification or in
% other study case is invited to visit the repository from authors at \url{http://b2asm.googlecode.com}.
%    Porque podemos verificar várias propriedades em tempo de projeto, documentar, simulacao através das ferramentas de suporte do método B...
%    Porque permite verificar software em nível de assemblagem.[citar trabalho de BHDL]
%    Porque utilizamos o método B puro  - [citar proposta do ALEMAO do ProB]
%  Contextualizacao com grande desafio da verificao formal em nivel de assemblagem[citar grande challenge] 
%  Conclusão da introducao.
%  Organizacao do paper
%  Basear-se no Relatorio técnico e no ERMAC
%

\section{B Notation}

[One page]

\subsubsection{Desing components B }
[half page]
[Explicar como as coias são modeladas]
% Use the tech report

\section{General definitions for hardware models}

We think that We find it convenient to provide a set of basic definitions to model hardware concepts. These
definitions are grouped in a library that we named $\mathit{HARDWARE\_LIBRARY}$ and arrange through four modules:
$\mathit{BIT\_DEFINITION}$, $\mathit{BIT\_VECTOR\_DEFINITION}$, $\mathit{BYTE\_DEFINITION}$ and
$\mathit{BIT\_VECTOR\_ARITHMETICS}$.

[[Explain better the figure]]

The dependency graph between these modules is depicted in
Figure~\ref{fig:hardware-definition-graph}.

\begin{figure}[h]
\centering
\includegraphics[width=.7\textwidth]{diagramaEstrutural.jpg}
\caption{Dependencies between modules of $\mathit{HARDWARE\_LIBRARY}$ and
platform model.}
\label{fig:hardware-definition-graph}

\end{figure}

Now, We will explain each of these modules and present them in detail.

\subsection{Definitions to represent and manipulate bits}

The entities defined in the module $\mathit{BIT\_DEFINITION}$ are the
type for bits, logical operations on bits (negation, conjunction,
disjunction, exclusive disjunction), as well as a conversion function
from booleans to bits.

First, bits are modeled as the set of integers $\{0, 1\}$:\\

$
\begin{array}{l}
\mathit{BIT} = \mathit{0..1} \\ 
\end{array}
$

The negation is a unary function on bits and it is defined as:

$
\begin{array}{l}
\mathit{bit\_not}  \in  \mathit{BIT}  \fun  \mathit{BIT}  \land \\
\forall ( \mathit{bb}). (\mathit{bb} \in \mathit{BIT} \implies \mathit{bit\_not}(\mathit{bb}) =
1-\mathit{bb})\\
\end{array}
$

The module also provides lemmas on negation that may be useful for the users of the library to develop proofs:

$
\begin{array}{l}
 \mathit{bit\_not}(0) = 1;  \mathit{bit\_not}(1) = 0; \\
\forall (\mathit{bb}).(\mathit{bb} \in \mathit{BIT} \implies \mathit{bit\_not}(\mathit{bit\_not}(\mathit{bb})) = \mathit{bb});
\end{array}
$

Conjunction is a unary function on bits and it is defined as:

$
\begin{array}{l}
\mathit{bit\_and} \in \mathit{BIT} \times \mathit{BIT} \fun \mathit{BIT} \land \\
\forall (\mathit{b1}, \mathit{b2}).(\mathit{b1}  \in \mathit{BIT}  \land \mathit{b2} \in \mathit{BIT} \implies \\
\quad ((\mathit{bit\_and}(\mathit{b1}, \mathit{b2}) = 1) \iff (\mathit{b1} = 1)  \land  (\mathit{b2} = 1)))
\end{array}
$

The module provides the following lemmas for conjunction, either:

$
\begin{array}{l}
 \mathit{bit\_and}(0,0) = 0;  \mathit{bit\_and}(0,1) = 0; \\
 \mathit{bit\_and}(1,0) = 0;  \mathit{bit\_and}(1,1) = 1; \\
\forall (\mathit{b1},\mathit{b2}).(\mathit{b1} \in \mathit{BIT} \land \mathit{b2} \in \mathit{BIT} \implies \\
\quad (\mathit{bit\_and}(\mathit{b1}, \mathit{b2}) = \mathit{bit\_and}(\mathit{b2},\mathit{b1}))); \\
\forall (\mathit{b1},\mathit{b2},\mathit{b3}).(\mathit{b1} \in \mathit{BIT} \land  \mathit{b2} \in \mathit{BIT} \land \mathit{b3} \in \mathit{BIT} \implies \\
\quad (\mathit{bit\_and}(\mathit{b1}, \mathit{bit\_and}(\mathit{b2},\mathit{b3})) = \mathit{bit\_and}(\mathit{bit\_and}(\mathit{b1},\mathit{b2}),\mathit{b3}))); \\
\forall (\mathit{b1}).(\mathit{b1} \in \mathit{BIT} \implies (\mathit{bit\_and}(\mathit{b1}, 1) = \mathit{b1})); \\
\forall (\mathit{b1}).(\mathit{b1} \in \mathit{BIT} \implies (\mathit{bit\_and}(\mathit{b1}, 0) = 0));
\end{array}
$

The module provides definitions of $\mathit{bit\_or}$ (disjunction) and $\mathit{bit\_xor}$ (exclusive disjunction),
as well as lemmas on those operators. These are standard and their expression in B is similar as for
$\mathit{bit\_and}$, they are thus omitted.

Finally, the conversion from booleans to bits is simply defined as:

$
\begin{array}{l}
\mathit{bool\_to\_bit} \in \BOOL \fun \mathit{BIT} \land \mathit{bool\_to\_bit} = \{ \TRUE \mapsto 1, \FALSE \mapsto 0 \} \\
\end{array}
$

Observe that all the lemmas that are provided in this module have been
mechanically proved by the theorem prover included with our B
development environment. None of these proofs requires human insight.

\subsection{Representation and manipulation of bit vectors}

Sequences are pre-defined in B, as functions whose the domain is an integer range with lower bound 1 (one). Indices
in bit vectors usually range from 0 (zero) upwards and the model we propose obeys this convention by making an
one-position shift where necessary.  We thus define bit vectors as non-empty sequences of bits, and
$\mathit{BIT\_VECTOR}$ is the set of all such sequences:

$
\begin{array}{l}
\mathit{BIT\_VECTOR} = \seq (\mathit{BIT})
\end{array}
$

The function $\mathit{bv\_size}$ returns the size of a given bit vector. It is basically a wrapper for the
predefined function $\mathbf{size}$ that applies to sequences.

$
\begin{array}{l}
\mathit{bv\_size} \in \mathit{BIT\_VECTOR} \fun \nat_1 \land \\
\mathit{bv\_size} = \lambda bv \bullet (bv \in \mathit{BIT\_VECTOR} \mid \mathbf{size}(bv))
\end{array}
$

We also define two functions $\mathit{bv\_set}$ and $\mathit{bv\_clear}$ that, given a bit vector, and a position of
the bit vector, return the bit vector resulting from setting the corresponding position to 0 or to 1, and a function
$\mathit{bv\_get}$ that, given a bit vector, and a valid position, each one returns the value of the bit at that
position. Only the first definition is shown here:

$
\begin{array}{l}
\mathit{bv\_set} \in \mathit{BIT\_VECTOR} \times \nat \fun \mathit{BIT\_VECTOR} \land \\
\mathit{bv\_set} = \lambda v, n \bullet (v \in \mathit{BIT\_VECTOR} \land 0 \in \nat \land n < \mathit{bv\_size}(v) \mid v \ovr \{ n+1 \mapsto 1 \})
\end{array}
$

The function $bv\_catenate$ takes as parameters two bit vectors $v$ and $w$, and returns the result of the
concatenation of $v$ and $w$, such that $v$ constitutes the most significant part of the result.

% $
% \begin{array}{l}
% \mathit{bv\_catenate} \in \mathit{BIT\_VECTOR} \times \mathit{BIT\_VECTOR} \fun \mathit{BIT\_VECTOR} \land \\
% \mathit{bv\_catenate} = \lambda v, w \bullet (v \in \mathit{ BIT\_VECTOR} \land w \in \mathit{ BIT\_VECTOR}  \mid v 
% %\conc
%   w)
% \end{array}
% $

\hspace*{0.00in} \it bv\_catenate  $\in$  \it BIT\_VECTOR  $\times$  \it BIT\_VECTOR  $\fun$ \it BIT\_VECTOR $\land$

\hspace*{0.00in} \it bv\_catenate \rm =  $\lambda$  v\rm ,\it w \rm . \rm (\it v  $\in$  \it BIT\_VECTOR $\land$ \it w $\in$  \it BIT\_VECTOR  $\mid$  \it v $\cat$ \it w\rm )

We also define a function $\mathit{bv\_zero}$ that, given a positive
integer $n$, return a bit vector of size $n$, with all bits set to 0.
A similar function, called $\mathit{bv\_one}$, with all bits set to 1
is also defined but not presented here.

$
\begin{array}{l}
\mathit{bv\_zero} \in \nat_1 \fun \mathit{BIT\_VECTOR} \land \\
\mathit{bv\_zero} = \lambda n \bullet (n \in \nat_1 \mid 1..n \times \{0\}) 
\end{array}
$

Additionally, the module provides definitions for the classical
logical combinations of bit vectors: $\mathit{bit\_not}$,
$\mathit{bit\_and}$, $\mathit{bit\_or}$ and $\mathit{bit\_xor}$. Only
the first two are presented here. Observe that the domain of the
binary operators is restricted to pairs of bit vectors of the same
length:

$
\begin{array}{l}
\mathit{bv\_not} \in \mathit{BIT\_VECTOR} \fun \mathit{BIT\_VECTOR} \land \\
\mathit{bv\_not} = \lambda v \bullet (v \in \mathit{BIT\_VECTOR} \mid \\
\quad \lambda i \bullet 0 .. \mathit{bv\_size}(v)-1 \mid \mathit{bit\_not}(v(i))) \land \\
\mathit{bv\_and} \in \mathit{BIT\_VECTOR} \times \mathit{BIT\_VECTOR} \fun \mathit{BIT\_VECTOR} \land \\
\mathit{bv\_and} = \lambda v_1, v_2 \bullet (v_1 \in \mathit{BIT\_VECTOR} \land v_2 \in \mathit{BIT\_VECTOR} \land \\
\quad \mathit{bv\_size}(v_1) = \mathit{bv\_size}(v_2) \mid \lambda i \bullet 0 .. \mathit{bv\_size}(v_1)-1 \mid \mathit{bit\_and}(v_1(i), v_2(i)))
\end{array}
$

We provide several lemmas on bit vector operations. These lemmas
express properties on the size of the result of the operations
as well as classical algebraic properties such as associativity
and commutativity. For example:

$
\begin{array}{l}
\forall v \bullet (v \in \mathit{BIT\_VECTOR} \implies \mathit{bv\_size}(\mathit{bv\_not}(v)) = \mathit{bv\_size}(v)) \\

\forall v \bullet (v \in \mathit{BIT\_VECTOR} \implies \mathit{bv\_not}(\mathit{bv\_not}(v)) = v) \\

\forall v_1, v_2 \bullet (\{v_1, v_2\} \subseteq \mathit{BIT\_VECTOR} \land \mathit{bv\_size}(v_1) = \mathit{bv\_size}(v_2) \implies \\
\quad \mathit{bv\_size}(\mathit{bv\_and}(v_1, v_2)) = \mathit{bv\_size}(v_1)) \\

\forall v_1, v_2 \bullet (\{v_1, v_2\} \subseteq \mathit{BIT\_VECTOR} \land \mathit{bv\_size}(v_1) = \mathit{bv\_size}(v_2) \implies \\
\quad \mathit{bv\_size}(\mathit{bv\_and}(v_1, v_2)) = \mathit{bv\_size}(v_2)) \\

\forall v_1, v_2 \bullet (\{v_1, v_2\} \subseteq \mathit{BIT\_VECTOR} \land \mathit{bv\_size}(v_1) = \mathit{bv\_size}(v_2) \implies \\
\quad \mathit{bv\_size}(\mathit{bv\_and}(v_1, v_2)) = \mathit{bv\_and}(v_2, v_1)) \\

\forall v_1, v_2, v_3 \bullet (\{v_1, v_2, v_3\} \subseteq \mathit{BIT\_VECTOR}
\land \mathit{bv\_size}(v_1) = \mathit{bv\_size}(v_2) \land \\
\quad \mathit{bv\_size}(v_1) = \mathit{bv\_size}(v_3) \implies \\ 
\quad \mathit{bv\_and}(\mathit{bv\_and}(v_1, v_2), v_3) = \mathit{bv\_and}(v_1,
\mathit{bv\_and}(v_2, v_3))) \\

\forall v \bullet (v \in \mathit{BIT\_VECTOR} \implies \\
\quad \mathit{bv\_and}(v, \mathit{bv\_zero}(\mathit{bv\_size}(v))) = \mathit{bv\_zero}(\mathit{bv\_size}(v))) \\

\forall v \bullet (v \in \mathit{BIT\_VECTOR} \implies \\
\quad \mathit{bv\_and}(v, \mathit{bv\_one}(\mathit{bv\_size}(v))) = v) \\

\end{array}
$

\subsection{Modelling bytes and bit vector of length 16}[I need review the english]

Bit vectors of length 8 are bytes. They form a common entity in
hardware design. We provide the following definitions:

% $
% \begin{array}{l}
% \mathit{BYTE\_WIDTH} = 8 \land \mathit{BYTE\_INDEX} = 0..(\mathit{BYTE\_WIDTH}-1) \land \\
% \mathit{BYTE} \subseteq \mathit{BIT\_VECTOR} \land \mathit{BYTE} = \{ v \in \mathit{BIT\_VECTOR} \land \mathit{bv\_size}(v) = \mathit{BYTE\_WIDTH}\} \land \\
% \mathit{BYTE\_ZERO} \in \mathit{BYTE} \land \mathit{BYTE} = \mathit{BYTE\_INDEX} \times \{0\}
% \end{array}
% $
% Version 2

\hspace*{0.0in}

\hspace*{0.0in}\it BYTE\_WIDTH \rm = 8 $\land$ \it BYTE\_INDEX \rm = 1 $\upto$ \rm (\it BYTE\_WIDTH\rm )$\land$

\hspace*{0.0in}\it PHYS\_BYTE\_INDEX \rm = \rm 0 $\upto$ \rm (\it BYTE\_WIDTH\rm -\rm 1\rm )\hspace*{0.10in} $\land$

\hspace*{0.0in}\it BYTE \rm = \rm \{ \it bt  $\mid$  \it bt $\in$ \it BIT\_VECTOR  $\land$  \it bv\_size\rm (\it bt\rm )\rm =\it BYTE\_WIDTH\rm \}\hspace*{0.10in} $\land$

\hspace*{0.0in}\it BYTE\_ZERO  $\in$  \it BYTE  $\land$

\hspace*{0.0in}\it BYTE\_ZERO \rm = \it BYTE\_INDEX  $\times$  \rm \{\rm 0\rm \}

\hspace*{0.0in}

The $\mathit{BYTE\_INDEX}$ is domain of byte modeled. it start at 1 to  obey a definition of sequences from B
method. However, the functions from byte encapsulate the acces and these functions use the
$\mathit{PHYS\_BYTE\_INDEX}$. The $\mathit{BYTE}$ type is a specializated type from $\mathit{BIT\_VECTOR}$, but it
has a size limit. Some other more specific definitions are created to help the developer. For example the constant
$\mathit{BYTE\_ZERO}$ and some below functions.

\begin{itemize}
  \item \it is\_zero $\in$ BYTE  $\fun$   BIT \rm - It returns 1 if the received byte is zero, otherwise
  returns 0.

  \item \it rotateright  $\in$  \it BYTE  $\fun$  \it BYTE \rm - It receives a byte and returns its rotated to
  right. There is a similar function named $\mathit{rotateleft}$.

\end{itemize}

By the very similar way, the type $\mathit{BV16}$ is created.

\subsection{Bit vector arithmetics}

Bit vectors are used to represent and combine numbers: integer ranges
(signed or unsigned) and floating point numbers.

Our library defines a function $\mathit{bv\_to\_nat}$ that maps bit
vectors to natural numbers:

$
\begin{array}{l}
\mathit{bv\_to\_nat} \in \mathit{BIT\_VECTOR} \fun \nat \land \\
\mathit{bv\_to\_nat} = \lambda v \bullet (v \in \mathit{BIT\_VECTOR} \mid \sum i \bullet (i \in \dom(v) \bullet v(i) \times 2^i))
\end{array}
$

We provide the following lemma:

$
\begin{array}{l}
\forall n \bullet (n \in \nat_1 \implies \mathit{bv\_to\_nat}(\mathit{bv\_zero}(n)) = 0)
\end{array}
$

\subsection{Basics data types}

There are data types commonly used in the microcontrollers. They are placed in
the project Types Librarys and define functions to manipulate data on levels of
basic types from hardware. However, in general is needed to detail or redefines
the types and its functions for each platform. Then, the specific concepts are
defined in the project's own platform.

There are six common basics data types, see details in table below.

\begin{tabular}{|p{3.5cm}|c|c|c|}\hline
 $Type\ Name$ & $Range$ & $Physical\ Size $\\ 
\hline
 UCHAR& 0..255 & 1 byte\\ \hline SCHAR& -128..127 & 1 byte\\ \hline BYTE & -- & 1
 byte\\ \hline USHORTINT & 0..65.535 & 2 byte\\ \hline SSHORTINT &
 -32.768..32.767 & 2 byte\\ \hline BV16 & -- & 2 byte\\ \hline
\end{tabular}

There is, for example, the function $\mathit{bv\_to\_nat}$ that is specialized
to $\mathit{byte\_uchar}$. How the set $\mathit{BYTE}$ is a subset of the
$\mathit{BIT\_VECTOR}$, then this function can defined as follows:

$
\begin{array}{l}
\mathit{byte\_uchar} = \lambda (v) \bullet ( v \in BYTE | bv\_nat(v) )
\end{array}
$

Usually, the definitions of the functions of the module $\mathit{TYPES}$ are the
re-used from the basic functions of the hardware library. This provides greater
confidence and facilitates the process of proof, because the prover can use the
defined previous lemma.

A more specific model of $\mathit{byte\_uchar}$ functions is:

$
\begin{array}{l}
\mathit{byte\_uchar} \in \mathit{BYTE}  \fun  \mathit{UCHAR}  \land \\
   \mathit{byte\_uchar} =  \lambda ( v0 )  \bullet  
   ( v0 \in \mathit{BYTE} |  2^{7}* bv\_get( v0 ,7 ) + 2^{6} * bv\_get( v0 ,6 ) + 2^{5}* bv\_get( v0 , 5   )\\ 
   + 2^{4}* bv\_get( v0 , 4 ) + 2^{3}*bv\_get( v0 , 3 ) + 2^{2}* bv\_get( v0 , 2 ) + 2 * bv\_get( v0 , 1 ) +  bv\_get( v0 ,0) )

\end{array}
$

The inverse function is easily defined as $\mathit{uchar\_byte}$.

$
\begin{array}{l}
\mathit{uchar\_byte} \in \mathit{UCHAR}  \fun  \mathit{BYTE}  \land \\
   \mathit{uchar\_byte} = \  ^{-1}(\mathit{byte\_uchar}
   )
\end{array}
$

We also created the following lemmas:

$
\begin{array}{l}
 \forall (val) \bullet (val \in \mathit{UCHAR} |
 \mathit{byte\_uchar}(\mathit{uchar\_byte}(val)) = val) \land\\
 \forall (by) \bullet (by \in \mathit{BYTE} |
 \mathit{uchar\_byte}(\mathit{byte\_uchar}(by)) = by)
\end{array}
$

Similarly, several other functions and lemmas were created for all other
data types.

\section{Description of the B model from Z80}
\label{sec:z80}

The \textit{Z80} is an important microcontroller developed in 1976 by \textit{Zilog}. In this time, It got a great
acceptance because suported all set instruction of \textit{8080}, one of the predecessors of the \textit{Pentium}
from \textit{Intel Corporation}. Then, the Z80 has the set instructions similar to the x86 architecture, it helps 
to build the model of x86 and it provides good features.

The main module includes a instance of memory module and accesses the definitions from: basic data types modules,
\textit{ALU} and \textit{Power2} \footnote{The \textit{Power2} module has the basics definitions to help the
theorems prover about calculus of power.}.

\begin{sloppypar}

\bf MACHINE

\hspace*{0.15in}\it Z80

\bf INCLUDES

\hspace*{0.10in}\it MEMORY

\bf SEES

\hspace*{0.10in}\it ALU, \it BIT\_DEFINITION, \it BIT\_VECTOR\_DEFINITION,

\hspace*{0.10in}\it BYTE\_DEFINITION, \it BV16\_DEFINITION,

\hspace*{0.10in}\it UCHAR\_DEFINITION, \it SCHAR\_DEFINITION,

\hspace*{0.10in}\it SSHORT\_DEFINITION ,\it USHORT\_DEFINITION, \it POWER2
\end{sloppypar}

% Chacteristics of microcontroller
The \textit{Z80} supports 158 differents instructions including all the 78 from
\textit{8080}.The instructions are classified into these categories: load and exchange;
block transfer and search; arithmetic and logical; rotate and shift; bit
manipulation (set, reset, test); jump, call and return; input/output; and basic
cpu control. Each instruction and external action at model is represented by B
operations. By default, all parameters from operations are or pre-defined elements
in the model or integers values in the decimal representation.

The internal registers contain 208 bits of reading/writing memory. It includes two sets of six general purpose
registers which may be used individually as  8-bits registers or as 16-bits register pairs.  The work registers are
represented by variable $\mathit{rgs8}$. The domain of $\mathit{rgs8}$ ($\mathit{id\_regs8}$) is a set formed by
identifiers of registers of 8 bits. These registers can be accessed in pairs, forming 16-bits, resulting in other
set of identifiers of 16-bits registers, named $\mathit{id\_reg16}$. The main work register of Z80 is the
accumulator ($\mathit{rgs8(a0)}$) used for arithmetic, logic, input/output and loading/storing operations.

\begin{sloppypar}
\bf SETS

\hspace*{0.10in}\it id\_reg\_8 \rm = \rm \{ \it a0 \rm , \it f0 \rm , \it f\_0 \rm , \it a\_0 \rm ,

\hspace*{1.0 in}\it b0 \rm , \it c0 \rm , \it b\_0 \rm , \it c\_0 \rm ,

\hspace*{1.00in}\it d0 \rm , \it e0 \rm , \it d\_0 \rm , \it e\_0 \rm ,

\hspace*{1.0in}\it h0 \rm , \it l0 \rm , \it h\_0 \rm , \it l\_0 \} ;

\hspace*{0.10in}\it id\_reg\_16 \rm = \rm \{ \it BC \rm , \it DE \rm , \it HL \rm , \it SP \rm , \it AF \rm \}
\end{sloppypar}

\subsection{Modelling registers and input and output ports} [I need review the english]

The Z80 CPU includes alternative set of accumulator, flag and general registers. The CPU contains a stack
pointer ($\mathit{sp}$), program counter ($\mathit{pc}$), two index registers ($\mathit{ix}$ and $\mathit{iy}$), an
interrupt register ($\mathit{i\_}$), a refresh register ($\mathit{r\_}$), two bits ($\mathit{iff1}$,
$\mathit{iff2}$) used to control the interruptions, a pair of bits to define the interruption mode ($\mathit{im}$)
and the input and output ports ($\mathit{i\_o\_ports}$). Below, its definitions are represented by
\textit{INVARIANT}.

\begin{sloppypar}
\bf INVARIANT

\hspace*{0.10in}\it rgs8  $\in$  \it id\_reg\_8  $\fun$  \it BYTE  $\land$

\hspace*{0.10in}\it pc  $\in$  \it INSTRUCTION  $\land$  \it sp  $\in$  \it BV16  $\land$  \it ix  $\in$  \it BV16  $\land$  \it iy  $\in$  \it BV16  $\land$

\hspace*{0.10in}\it i\_  $\in$  \it BYTE  $\land$  \it r\_\hspace*{0.10in} $\in$  \it BYTE  $\land$

\hspace*{0.10in}\it iff1  $\in$  \it BIT  $\land$ \hspace*{0.10in}\it iff2  $\in$  \it BIT  $\land$

\hspace*{0.10in}\it im \rm : \rm (\it BIT $\times$ \it BIT\rm )  $\land$

\hspace*{0.10in}\it i\_o\_ports  $\in$  \it BYTE  $\fun$  \it BYTE
\end{sloppypar}

\subsection{Flag register} [I need review the english]
Another important element is the ``z'' register ($\mathit{rgs8(z0)}$),  that is used as a flag register. This
register uses only six bits to represent the result status of each instruction. 	
According to the official manual the bits 3 and 5 are not used and the others bits have the follow meaning:
\begin{description}
  \item[$\mathit{bv\_get(rgs8(z0),0)}$] - The Carry bit
  \item[$\mathit{bv\_get(rgs8(z0),1)}$] - The Add/Subtract bit
  \item[$\mathit{bv\_get(rgs8(z0),2)}$] - The Parity or Overflow bit
  \item[$\mathit{bv\_get(rgs8(z0),4)}$] - The Half Carry bit
  \item[$\mathit{bv\_get(rgs8(z0),6)}$] - The Zero bit
  \item[$\mathit{bv\_get(rgs8(z0),7)}$] - The Sign bit 
\end{description}

\begin{property}[Assuring the absence of Overflow]
 To assure that don´t happen overflow the developer can add this expression ($\mathit{bv\_get(rgs8(z0),0)} \neq 1
 \land \mathit{bv\_get(rgs8(z0),2) \neq 1}$)in the invariant. By default, the overflow can happen, but many times it
 can be dangerous. Then, the developer may prohibit its use. This restriction can also become the more dificult to
 specify.
\end{property}

\subsection{Manipulation data functions from Z80} [I need to review the english] There are some specific functions
from Z80 to manipulate the data. The $\mathit{bv\_ireg\_plus\_d}$ is  used to indexed address. It received the
value of register ($\mathit{ix}$ or $\mathit{iy}$) and displacement to return the sum, the result is the memory address dislocated, see
its the definition.

\hspace*{0.0in}\it bv\_ireg\_plus\_d \rm : \rm(\it BV16  $\times$  \it SCHAR  $\fun$  \it BV16\rm )  $\land$

\hspace*{0.0in}\it bv\_ireg\_plus\_d \rm =  $\lambda$  \rm ( \it ix\_iy \rm , \it disloc \rm ) \rm . \rm ( \it ix\_iy  $\in$  \it BV16  $\land$  \it disloc  $\in$  \it SCHAR

\hspace*{0.20in}$\mid$ \it ushort\_bv16 \rm ( \rm (\it bv16\_ushort \rm ( \it ix\_iy \rm ) \rm + \it disloc \rm ) 
$\mod$ \rm 6\rm 5\rm 5\rm 3\rm 6 \rm ) \rm )

Another derived function is  $\mathit{bv\_9ireg\_plus\_d0}$ , this returns the value in the memory  address
returned by $\mathit{bv\_ireg\_plus\_d}$ function ands its definition is similar.

There is a specific function to refresh the flag register, it is named $\mathit{update\_reg\_flag}$. It is typed of
follow mode: \it update\_flag\_reg \rm $\in$ \rm (\it BIT  $\times$  \it BIT  $\times$  \it BIT  $\times$  \it
BIT $\times$  \it BIT  $\times$  \it BIT\rm) $\fun$  \rm (\rm \{\it f0\rm \}  $\times$  \it BYTE\rm ).

\it update\_flag\_reg \rm =  $\lambda$  \rm (\it s7 \rm, \it z6 \rm,\it h4 \rm,\it pv2 \rm ,\it n1
\rm ,\it c0 \rm)$\bullet$

\rm ( \it s7 $\in$ \it BIT $\land$ \it z6 $\in$ \it BIT $\land$ \it h4 $\in$ \it
BIT $\land$ \it pv2 $\in$ \it BIT $\land$ \it n1 $\in$ \it BIT $\land$ \it c0 $\in$ \it BIT

\hspace*{0.20in}\rm $\mid$( \it f0  $\mapsto$  \rm \rm [\it c0\rm , \it n1\rm , \it pv2\rm , \rm 1\rm , \it
h4\rm , \rm 1\rm , \it z6\rm , \it s7\rm \rm ]\rm ) \rm )

\subsection{Program, stack and data memory}[I need review the english]

The Z80 uses a unique memory for storing program instructions, data stack and data work. The memory has 16-bits
addressing and each address represent a byte. Thus, the data from the memory module is very simple, as shown below,
but an additional care must be taken to preserve the consistency of memory.

\begin{sloppypar}
\bf INVARIANT \\
\hspace*{0.10in}\it mem  $\in$  \it BV16  $\fun$  \it BYTE

\end{sloppypar}

In general, the instructions can access all memory address, but it is very dangerous. For add more security, is
importante that the program instructions has the access limited by region. Thus, the designer can specify address
regions to restrict the access from instructions. The address regions can be especified how shown below.

$
\begin{array}{l}
\mathit{PROGRAM\_R\_ADR} = 0..16384 \land \mathit{DATA\_R\_ADR} = 16385..49151 \land \\
\mathit{STACK\_R\_ADR} = 49152..65535
\end{array}
$

\begin{property}[Assuring the absence of overlapping of address regions]
 To assure that address regions are well defined, then the designer must to verify the below expression.
\end{property}

\begin{sloppypar}
\hspace*{0.10in}\it PROGRAM\_R\_ADR $\cap$ DATA\_R\_ADR $\cap$  STACK\_R\_ADR $=$ \{\}
\end{sloppypar}

\begin{property}[Preserving the consistency of the memory]
In general, the access to some address regions is dangerous. Then, in each instruction has a specific
pre-condition that verify if the new address memory, that will be updated, is member of its region. For example,
the $\mathit{PUSH}$ program instruction allow write only in the region of stack ($\mathit{STACK\_R\_ADR}$). 
%Thus,if the user try to write in incorrect region then the called of operation cannot be proved.
\end{property}

\subsection{Arithmetic logic unit}

There are many functions defined in the module \textit{ALU}. In general, these functions take basic definitions to build
new specific functions. The function $\mathit{half8UCHAR}$ is used to get the half part of $\mathit{UCHAR}$ value.
It is importante to know the half carry and it is used in the $\mathit{add8UCHAR}$.

\hspace*{0.0in}

\hspace*{0.0in}\it half8UCHAR  $\in$  \it UCHAR  $\fun$  \it UCHAR  $\land$

\hspace*{0.0in}\it half8UCHAR \rm =  $\lambda$  \rm (\it ww\rm )\rm .\rm (\it ww  $\in$  \it UCHAR  $\mid$  \it ww  $\mod$  \it $2^{4}$\rm )

\hspace*{0.0in}

The function $\mathit{add8UCHAR}$ receive a bit carry and two $\mathit{UCHAR}$ values and return respectively the 
sum, the sign bit, the carry bit, the half carry bit and the zero bit. It is typed how follow: \it add8UCHAR \rm :
\rm (\it BIT $\times$ \it UCHAR $\times$ \it UCHAR\rm ) $\fun$ \rm (\it UCHAR $\times$  \it BIT  $\times$  \it BIT  $\times$  \it BIT  $\times$  \it BIT\rm ) and its definitions is:

\hspace*{0.20in}

\hspace*{0.0in}\it add8UCHAR\rm = $\lambda$ \rm(\it carry\rm,\it w1\rm \it w2\rm)\rm.\rm

\hspace*{0.0in}(\it carry $\in$  \it BIT  $\land$  \it w1  $\in$  \it UCHAR  $\land$  \it w2  $\in$  \it UCHAR
$\mid$

\hspace*{0.40in}\rm(\rm(\rm(\it carry \rm + \it w1 \rm + \it w2 \rm )  $\mod$  \it $2^{8}$ \rm ),

\hspace*{0.40in}\it bool\_bit\rm ( \it carry \rm + \it uchar\_schar\rm (\it w1\rm ) \rm + \it uchar\_schar \rm (\it
w2\rm ) $<$ \rm 0\rm ),

\hspace*{0.40in}\it bool\_bit\rm ( \it carry \rm + \it w1 \rm + \it w2 $>$ \it UCHAR\_MAX\rm )\rm ,

\hspace*{0.40in}\it bool\_bit\rm ( \it carry \rm + \it half8UCHAR\rm (\it w1\rm ) \rm + \it half8UCHAR\rm ( \it
w2\rm )  $\geq$  \it $2^{4}$\rm )\rm,

\hspace*{0.40in}\it bool\_bit\rm ( \rm ( \rm (\it carry \rm + \it w1 \rm + \it w2 \rm )  $\mod$  \it $2^{8}$ \rm
)\rm = \rm 0\rm )\rm )\hspace*{0.10in}\rm )

\hspace*{0.20in}

A similar function to subtract operation is $\mathit{substract8UCHAR}$. There are the same functions for the
$\mathit{SCHAR}$ type, they are respectively $\mathit{add8SCHAR}$ and $\mathit{substract8SCHAR}$, all these
functions are of 8 bits ($\mathit{BYTE}$) and defined similarly. In the same way, the arithmetic functions for 16
bits ($\mathit{BV16}$) are defined.

Other more simply functions also are typed and explained below:

\begin{itemize}
  \item \it inc  $\in$ \it BYTE $\fun$  \it BYTE \rm - It receives a byte and returns its decrement. There is a
  similar function named $\mathit{dec}$.

  \item \it instruction\_next  $\in$  USHORT  $\fun$  USHORT \rm - It receives the pc actual
  value and return its increment.

  \item \it is\_negative  $\in$  \it BYTE  $\fun$  \it BIT \rm - It returns 1 if the received
  byte is zero, otherwise returns 0.

  \item \it update\_refresh\_reg \rm - It receives a byte and returns its increment until the seventh bit.

\end{itemize}

Every logic functions are defined in the \textit{BYTE} and \textit{BV16} module are included in the \textit{ALU}
module. Then, it was not necessary rewrite these functions.

\subsection{Modelling the instructions}[I need review the english]

The main module (\textit{Z80}) has two types of operations, a type represent the external actions and the other
represent the instructions of microcontrollers. The external actions are shown in the section
~\ref{sec:externalactions} and the other type represents the instructions of microcontroller. A simple example from
instruction is a $\mathit{LD\_(nn)\_A}$ shown below. Many times to model a instruction is necessary to use the pre-defined functions,
this facilitates the construction of model. This instruction use the $\mathit{updateAddressMem}$ function from
\textit{Memory} module and it receives a address memory and its new memory value. After, it increments the program
counter ($\mathit{pc}$) and update the refresh register ($\mathit{r\_}$).

\hspace*{0.00in}\bf LD\_9nn0\_A \rm ( \it nn \rm ) \rm =

\hspace*{0.20in}\bf PRE \it nn $\in$ \it USHORT\hspace*{0.15in} $\land$ \hspace*{0.10in}\it nn\hspace*{0.10in} $\in$  \it DATA\_R\_ADR

\hspace*{0.20in}\bf THEN

\hspace*{0.20in}\bf updateAddressMem \rm ( \it ushort\_bv16 \rm ( \it nn \rm ) \rm , \it rgs8 \rm ( \it a0 \rm )
\rm )  $\para$

\hspace*{0.20in}\it pc \rm := \it instruction\_next \rm ( \it pc \rm )  $\para$  \it r\_ \rm := \it update\_refresh\_reg\rm (\it r\_\rm )

\hspace*{0.00in}\bf END\rm ;

The other instructions have a similar structure.

\subsection{Modelling the external action}
\label{sec:externalactions}

The external actions change the state of microcontroller and they aren´t instructions, for example, the refreshing
of I/O port and the interruptions request. The extern action are also model by operations and they are prefixed
with ``$ext\_$'' and after the name of action. There are just four extern actions: $ext\_update\_io\_ports$,
$ext\_NMI$ and $ext\_INT$, $ext\_Reset$. The $ext\_update\_io\_ports$ just change the state of I/O port, see.

\hspace*{0.20in}\bf ext\_update\_io\_ports\rm (\it address\rm ,\it value\rm )\rm =

\hspace*{0.20in}\bf PRE \it address  $\in$  \it UCHAR  $\land$ \hspace*{0.10in}\it value  $\in$  \it SCHAR \bf THEN

\hspace*{0.40in}\it io\_ports \rm ( \it uchar\_byte \rm ( \it address \rm ) \rm ) \rm := \it schar\_byte \rm ( \it
value \rm )

\hspace*{0.20in}\bf END\rm

The others are related to interrupts. The interrupts allows that the devices suspend a routine from CPU and start
another service routine. This service routine can exchange data or signals between CPU and external devices. When a
routine is finished, then the CPU come back to the last routine that was interrupted.

For the interrupts, the following things are important:  the interrupt flip-flops ($\mathit{iff1}$ and
$\mathit{iff2}$), the types of interrupts (maskable and non-maskable), the interrupt mode (set with the $\mathit{IM
0}$, $\mathit{IM 1}$, $\mathit{IM 2}$ instructions) and the $\mathit{i\_}$ register.

The $\mathit{iff1}$ and $\mathit{iff2}$ control the maskable interrupts ($\mathit{INT}$). When the $\mathit{iff1}$
is set, the interrupt is enable, otherwise it is disable. The $\mathit{iff2}$ is used only as a tempory storage
place for $\mathit{iff1}$. The $\mathit{iff1}$ is set or reset by instructions $\mathit{EI}$ and $\mathit{DI}$.

The interruptions and the \textit{reset} action change the state of
microcontroller. Theses actions are modeled by B operations and the its main
effects are shown below\footnote{Some definitions: $\mathit{sp\_minus\_two}$ =
$\mathit{dec\_BV16(dec\_BV16(sp))}$, $\mathit{sp\_minus\_one}$=
$\mathit{dec\_BV16(sp)}$, $\mathit{pc\_high}$ is the most significant 8 bits and
$\mathit{pc\_low}$ is the least significant. }.

 \textbf{NMI} - Non-maskable interrupts - The non-maskable cannot be disable
 by the programmer. Then, when a device makes a request, the $sp$ is pushed, the $pc$ receive
 $66H$ (102 in decimal), the $\mathit{iff1}$ is reset , $\mathit{iff2}$ stores
 $\mathit{iff1}$ and refresh register is updated.

\begin{sloppypar}
\bf updateStack\rm (\rm \{ \rm (\it sp\_minus\_two  $\mapsto$  \it pc\_low\rm )\rm ,\rm (\it sp\_minus\_one  $\mapsto$ \it pc\_high \rm ) \rm \}\rm )$\para$

\it sp \rm := \it sp\_minus\_two  $\para$ \it pc \rm := \rm 1\rm 0\rm 2 $\para$ \it iff1\rm :=\rm 0  $\para$  \it iff2\rm := \it iff1 $\para$

\it r\_ \rm := \it update\_refresh\_reg\rm (\it r\_\rm )\\
\end{sloppypar}

  \textbf{INT} - Maskable Interrupt -  This is usualy reserved to important functions that can be enabled and
  disabled by the programmer. When a maskable interrupt action happens, both $\mathit{iff1}$ and $\mathit{iff2}$ are
  cleared, disabling the interrupts, the $sp$ is pushed, the refresh register is updated  and the other effects
  depend on the interrupt mode.

 \begin{itemize}

  \item The mode 0 is compatible with 8080 and  \it im \rm = \rm ( \rm 0 $\mapsto$  \rm 0 \rm ). When a 
  non-maskable interruption happen , the CPU fetches an instruction of one byte from an external device, usually an
  RST instruction, and the CPU executes it. The instruction code is received from a external device by data bus and
  it is represented by integer parameter called $\mathit{byte\_bus}$ .

  \item The mode 1 is the easiest and \it im \rm = \rm ( \rm 0  $\mapsto$  \rm 1 \rm ). Simply, when a non-maskable
  interruption happens, the program counter receive $38H$ (56 in decimal).

  \item The mode 2 is the most powerfull and  \it im \rm = \rm ( \rm 1  $\mapsto$ \rm 1 \rm ). When a non-maskable
  interruption happens, an indirect call can be made to any address memory. The program counter receives in the
  part most significant the $\mathit{i\_}$ register and, in the  part least siginificant, the $\mathit{byte\_bus}$
  with the last bit reset.
 \end{itemize}

The essential part of maskable interrupt is shown below\footnote{ The
 $\mathit{byte\_bus}$ is parameter from  $\mathit{INT}$ operation }.

\begin{sloppypar}

\hspace*{0.00in}\bf IF \it im \rm = \rm ( \rm 0  $\mapsto$  \rm 0 \rm ) \bf THEN

\hspace*{0.10in}\bf IF\hspace*{0.10in}\it byte\_bus \rm $\in$  \it opcodes\_RST\_instruction

\hspace*{0.10in}\bf THEN\hspace*{1.05in}

\hspace*{0.40in}\it pc \rm := \it byte\_bus \rm - \rm 1\rm 9\rm
9\hspace*{0.10in} $\para$

\hspace*{0.40in}\bf updateStack\rm ( \rm \{ \it stack\rm ( \it
sp\_minus\_one\rm )  $\mapsto$  \it pc\_low\rm ,

\hspace*{0.40in}\it stack\rm (\it sp\_minus\_two\rm )  $\mapsto$  \it pc\_high
\rm \} \rm )  $\para$

\hspace*{0.40in}\it sp \rm := \it sp\_minus\_two  $\para$ \hspace*{0.10in}\it
r\_ \rm := \it update\_refresh\_reg\rm (\it r\_\rm )

\hspace*{0.10in}\bf ELSIF \it byte\_bus \rm = \it opcode\_\ldots\_instruction

\hspace*{0.40in}\bf \ldots

\hspace*{0.10in}\bf END

\hspace*{0.00in}\bf ELSIF\hspace*{0.10in}\it im \rm =\hspace*{0.10in}\rm ( \rm
0  $\mapsto$  \rm 1 \rm ) \bf THEN

\hspace*{0.10in}\it pc \rm :=\hspace*{0.10in}\rm 5\rm 6  \hspace*{0.80in}
$\para$

\hspace*{0.10in}\bf updateStack\rm ( \rm \{ \it stack\rm (\it
sp\_minus\_one\rm )  $\mapsto$  \it pc\_low\rm ,

\hspace*{0.10in}\it stack\rm (\it sp\_minus\_two\rm )  $\mapsto$  \it pc\_high
\rm \} \rm )  $\para$

\hspace*{0.10in}\it sp \rm := \it sp\_minus\_two  $\para$ \hspace*{0.10in}\it
r\_ \rm := \it update\_refresh\_reg\rm (\it r\_\rm )\hspace*{0.85in}

\hspace*{0.00in}\bf ELSIF\hspace*{0.10in}\it im \rm = \rm ( \rm 1  $\mapsto$ 
\rm 1 \rm ) \bf THEN  \hspace*{0.70in}

\hspace*{0.10in}\it pc \rm := \it bv16\_ushort\rm (\it byte\_bv16\rm ( \it i\_
\rm ,\it bv\_clear\rm (\it rotateleft\rm (\it uchar\_byte\rm (\it byte\_bus\rm )\rm )\rm ,\rm 0\rm )\rm )\rm )$\para$

\hspace*{0.10in}\bf updateStack\rm ( \rm \{ \it stack\rm ( \it
sp\_minus\_one\rm )  $\mapsto$  \it pc\_low\rm ,

\hspace*{0.10in}\it stack\rm (\it sp\_minus\_two\rm )  $\mapsto$  \it pc\_high
\rm \} \rm )  $\para$

\hspace*{0.10in}\it sp \rm := \it sp\_minus\_two  $\para$ \hspace*{0.10in}\it
r\_ \rm := \it update\_refresh\_reg\rm (\it r\_\rm )

\hspace*{0.00in}\bf END\hspace*{0.10in}\\
\end{sloppypar}

\textbf{RESET}  - This just reset the registers related to the interruptions.

\begin{sloppypar}
\it iff1 \rm :=\rm 0 $\para$ \it iff2\rm :=\rm 0 $\para$  \it  im\rm := \rm (\rm 0 $\mapsto$ \rm 0\rm )  $\para$ 
\it pc\rm :=\rm 0 $\para$ \it i\_ \rm := [0,0,0,0,0,0,0,0] $\para$

\it rgs8 \rm := \it rgs8  $\lover$  \rm \{ \rm (\it a0  $\mapsto$ \rm [0,0,0,0,0,0,0,0] , \rm (\it
f0 $\mapsto$ \rm [0,0,0,0,0,0,0,0] \rm \} $\para$

\it r\_  \rm := [0,0,0,0,0,0,0,0] $\para$ \it sp \rm := [0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0]
\end{sloppypar}

% 
%     Para controlar o registrador I, existem duas intrucoes:
% 
%     - LD A,I (carrega o acumulador com o valor contido em I);
%     - LD I,A (carrega I com o valor contido no acumulador).
% 
%     A instrucao LD A,I possui uma funcao extra, ja' discutida
% em outra aula: ela carrega na `flag' de paridade/sobrecarga o
% valor do IFF ("Interruption Flip-Flop"). Se IFF for 0, entao a
% interrupcao mascaravel esta' desativada. Se IFF for 1, entao
% ela esta' ativada.
% %%%% FIM - http://download.unesp.br/msx/asm/course_4.txt
% 
% 
% 
% 
% Instruções relacionadas com o sistema de interrupção:
%  HALT,EI,DI,M0,M1,M2 
%  
%  LD_A_R, LD_A_I\ldots,  
%  CALL, RET,  RETI,RETN

\section{Input and output ports}

The Z80 has an extensive set of input and output (I/O) instruction and 256 ports for
devices. This model can transfer data blocks and between the I/O devices and any
of the internal registers or memory address.

The $\mathit{IN\_r(C)}$ \footnote{The tools B don't allow use parenteses in identifiers, then the characteres
\{``('',``)''\} are replaced respectively by \{``9'',``0''\} in the real specification.} instruction is represented
by the follow B operation. It is from I/O group, then it reveives a identifier of register
$\mathit{r}$ and, in this place, it stores the value of $\mathit{C}$ port address. Besides, it increments the
program counter and updates the flag registers.

\hspace*{0.0in}\bf IN\_r\_9C0 \rm ( \it rr \rm ) \rm =

\hspace*{0.0in}\bf PRE \it rr  $\in$  \it id\_reg\_8  $\land$  \it rr  $\not =$  \it f0\hspace*{0.15in}\bf THEN

\hspace*{0.20in}\bf ANY

\hspace*{0.40in}\it negative \rm , \it zero \rm , \it half\_carry \rm , \it pv \rm , \it add\_sub \rm , \it carry

\hspace*{0.20in}\bf WHERE

\hspace*{0.40in}\it negative $\in$ \it BIT $\land$ \it zero $\in$ \it BIT $\land$ \it half\_carry $\in$ \it BIT 
$\land$ \it pv $\in$ \it BIT $\land$

 \hspace*{0.40in}\it add\_sub $\in$ \it BIT $\land$ \it carry $\in$ \it BIT  $\land$

\hspace*{0.40in}\it negative \rm = \it is\_negative \rm ( \it io\_ports \rm ( \it rgs8 \rm ( \it c0 \rm ) \rm ) \rm )  $\land$

\hspace*{0.40in}\it zero \rm = \it is\_zero \rm ( \it io\_ports \rm ( \it rgs8 \rm ( \it c0 \rm ) \rm ) \rm )  $\land$

\hspace*{0.40in}\it half\_carry \rm = \rm 0  $\land$

\hspace*{0.40in}\it pv \rm = \it parity\_even \rm ( \it io\_ports \rm ( \it rgs8 \rm ( \it c0 \rm ) \rm ) \rm ) $\land$

\hspace*{0.40in}\it add\_sub \rm =\hspace*{0.10in}\rm 0  $\land$

\hspace*{0.40in}\it carry \rm = \it z\_c

\hspace*{0.20in}\bf THEN

\hspace*{0.40in}\it rgs8 \rm := \it rgs8  $\lover$  \rm \{ \rm ( \it rr  $\mapsto$  \it io\_ports \rm ( \it rgs8 \rm ( \it c0 \rm ) \rm ) \rm ) \rm ,

\hspace*{0.40in}\it update\_flag\_reg\rm (\it negative\rm,\it zero\rm,\it half\_carry\rm,\it pv\rm,\it add\_sub\rm,\it carry)\rm\}$\para$

\hspace*{0.40in}\it pc \rm := \it instruction\_next \rm ( \it pc \rm )  $\para$  \it r\_ \rm := \it update\_refresh\_reg\rm (\it r\_\rm )

\hspace*{0.0in}\bf END \rm

% $
% \begin{array}{l} 
% \mathit{IN\_r\_9C0} ( rr ) = \\
% \quad   \PRE rr \in id\_reg\_8    \THEN\\
% \quad\quad \ANY data\_in, negative , zero , half\_carry , pv , add\_sub , carry\\ 
% \quad\quad \WHERE data\_in \in \mathit{BYTE} \land negative \in \mathit{BIT}\land\\
% \quad\quad\quad carry \in \mathit{BIT} \land half\_carry \in \mathit{BIT} \land zero \in \mathit{BIT} \land \\
% \quad\quad\quad negative = is\_negative (data\_in) \land zero = is\_zero(data\_in ) \land\\
% \quad\quad\quad half\_carry = 0 \land pv =parity\_even\_BYTE ( data\_in )    \land\\
% \quad\quad\quad add\_sub =  0 \land carry = z\_c \\
% \quad\quad  \THEN \\
% \quad\quad\quad i\_o\_ports ( rgs8 ( c0 ) ) := data\_in ||\\
% \quad\quad\quad rgs8 := rgs8 <+ \{ ( rr |-> data\_in ) ,\\
% \quad\quad\quad get\_new\_flag\_register\_SZ\_H\_PvNC ( rgs8 , negative , zero, half\_carry , pv , add\_sub , carry ) \} ||\\
% \quad\quad\quad pc := instruction\_next( pc )\\
% \quad\quad \END\\
% \quad \END\\
% \end{array}
% $

\section{Proofs}

[O que as provas garantem?]
As obrigações de provas geradas pela ferramentas garante a verficação dos tipos de dados e que as
expressoes são bem formadas \footnote{An expression is called "well-defined" (or unambiguous) if its definition
assigns it a unique interpretation or value.}. Deste modo, as propriedades citadas anteriomente também são
verificadas. As propriedades citadas anteriormente oferecem garantias que

[Enfatizar a vantagem]
The traditional don´t allow verify this things and the properties is flexible. The flexbility allow add
restrictions about tempo de execuçao and until specifics restrictions about the application of microcontroller.

This flexibility of create propertis allow create conditions to verify if the software to finish in cert instant.

[Estatisticas]
Types Library

Hardware Library

[Groups]
\begin{center}

\begin{tabular}{llll}
\textbf{Group} &  \textbf{Proofs No Obvios} & \textbf{Proofs WD} & × \\ 
Input and Output & 54 & 296  \\ 
Logic and Ari & 134 & 413  \\ 
Manipulation of Bits & 160 & 742  \\ 
Extern actions & 51 & 84  \\
General & 140 & 615  \\
Header & 68 & 169  \\
 & \textbf{Subtotal} 607 & \textbf{Subtotal} 2319  &  \textbf{Total} 2926 \\
\end{tabular}
\end{center}

[Conclusao]
A primeira vez que realiza as provas é um pouco trabalhoso, porque as definições de tipos e de hardware não
estavam completas. Então, um trabalho maior foi demandado, por outro lado, foi nece

The first time
Paralelismo vantagem

 \section{Conlusitions}

[]

%  Para a modelagen dessa plataforma, essa modelagen contém linhas e $np $
%  provas não obvias. Uma forma de acelerar o processo de prova, foi paralelizar
%  as provas. Isto foi feito em 4 IMacs e consumiu $ tp $ tempo para verificar
%  automaticamente $np - p $. As demais foi necessário uma certa interação com
%  o provador para sua verificação.

%  Para oferecer uma maior garantia do modelo com a implementação do processador
%  algumas atividades foram realizadas. Primeiramente, tentantamos animar e gerar
%  código respectivamente nas ferramentas ProB e AtelierB. Entratanto, o modelo
%  cotém algumas construções complexas que ainda não são bem suportada por elas.
%  Então, foi realizada uma apresentação sobre a descrição do modelo e foi
%  realizada um nova revisão com 3 pessoas.

\section{Acknowledges}

This work received supports from CNPq and ANP (National Agency of Oil). Besides, the research Bruno Gomes help in
the revision.

\begin{thebibliography}{5}

% Paper about BHDL

% Paper of ALEMÃO PROB ,java virtual machine

% O manual do Z80

\bibitem {Abrial}
Abrial, J. R. The B Book: Assigning Programs to Meanings. Cambridge University Press, United States of
America, 1 edition, 1996.

\bibitem {CLEARSY}
CLEARSY. Atelier B. http://www.atelierb.eu.

\bibitem {Dantas_Semish08}
 Dantas, B; D\'{e}harbe, D; Galv\~{a}o, S; ET AL. Proposta e avalia\c{c}\~{a}o de uma abordagem de desenvolvimento de
 software fidedigno por constru\c{c}\~{a}o com o m\'{e}todo B. In: SEMISH, Bel\'{e}m, 2008. SBC.

\bibitem {Dantas_Report08}
Dantas, B; D\'{e}harbe, D; Galv\~{a}o, S; ET AL.Verified compilation based on the B method: initial
appraisal.Technical Report, UFRN - DIMAP, 2008.

\bibitem {Bell} 
 D. Bell, I. M. E. J. P. Software Engineering: A Programming Approach. Prentice Hall, New York, 2 edition, 1992.

\bibitem {HOARE}
HOARE, C. A. R. The verifying compiler, a grand challenge for computing research. In: VMCAI, p. 78-78, 2005.

\bibitem{Medeiros}
 Medeiros Jr, Val\'{e}rio; Galv\~{a}o, Stephenson; D\'{e}harbe, David; Modelagem de Microcontroladores em B. In: ERMAC, Natal, 2008. SBMAC.

\bibitem {WORDSWORTH}
WORDSWORTH, J. B. Software Engineering with B. Addison Wesley, Boston, 1996.

\end{thebibliography}

\end{document}
%
% \bibitem {clar:eke}
% Clarke, F., Ekeland, I.:
% Nonlinear oscillations and
% boundary-value problems for Hamiltonian systems.
% Arch. Rat. Mech. Anal. {\bf 78} (1982) 315--333
% %
% \bibitem {clar:eke:2}
% Clarke, F., Ekeland, I.:
% Solutions p\'{e}riodiques, du
% p\'{e}riode donn\'{e}e, des \'{e}quations hamiltoniennes.
% Note CRAS Paris {\bf 287} (1978) 1013--1015
% %
% \bibitem {mich:tar}
% Michalek, R., Tarantello, G.:
% Subharmonic solutions with prescribed minimal
% period for nonautonomous Hamiltonian systems.
% J. Diff. Eq. {\bf 72} (1988) 28--55
% %
% \bibitem {tar}
% Tarantello, G.:
% Subharmonic solutions for Hamiltonian
% systems via a $\bbbz_{p}$ pseudoindex theory.
% Annali di Matematica Pura (to appear)
% %
% \bibitem {rab}
% Rabinowitz, P.:
% On subharmonic solutions of a Hamiltonian system.
% Comm. Pure Appl. Math. {\bf 33} (1980) 609--633
% \end{thebibliography}
%