### 2009-08-06 trunk/papers/SBMF2009/Presentation/content/5_Proofs.tex

\section{Prove Process}

\begin{frame}
\frametitle{Prove Process of Projects}

\begin{itemize}
   \item Statistics:\\
    \begin{center}{\footnotesize
  	 \begin{tabular}{|c|c|c|c|}
		\hline
		 \textsl{Projects} &  \textsl{POs}	& \textsl{POs WD} &	\textsl{Total}\\
		\hline
		\textit{Power} & 	21 &	4 &	25\\
		\hline
		\textit{Types Library} &	22 &	68 &	90\\
		\hline
		\textit{Hardware Library} &	108 &	324 &	432\\
		\hline
		\textit{Z80} &	631	 & 2470 & 	3101\\
		\hline
		   &    &  & 		\textbf{3648}\\
		\hline
	 \end{tabular}
   }
	\end{center}

   \item  Approximately 25\% are not solved automatically
	\item  Approximately 18 days are needed to make almost all the proofs using a monolithic computer
		\begin{itemize}
			\item But You can reduce this time if you use several computer to prove
		\end{itemize}
	\item  Sometimes was needed  replay the proofs when were needed news fixes
\end{itemize}

\end{frame}

\begin{frame}
\frametitle{A example of user pass}

\begin{itemize}
   \item Specialized:
   \item \ldots
   \item Generally:
\end{itemize}

\end{frame}