### 2009-08-07 papers/SBMF2009/Presentation/content/7_Conclusions.tex

\section{Conclusions}

 \begin{frame}
 \frametitle{Considerations about the Work}

 \begin{itemize}
   \item It be able to: \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ 
    \includegraphics[width=.2\textwidth]{figures/processador_generico.jpg}
 	\begin{itemize} 
 	  \item Identify  errors on the official reference
 	  \item Help assembly programmers
 	  \item Formal verification until assembly language 
     \end{itemize} 
     \item Future work is the verified development of a petroleum production test system\\ 
     \begin{center}   
 	   \includegraphics[height=.25\textheight]{figures/dutos.png}
	 \end{center}%\ \ \ \ \ \ \ \ \ \ 
     %   \includegraphics[height=.25\textheight]{figures/trem_bala2.png}

 \end{itemize}

 \end{frame}

\begin{frame}
  \frametitle{Sugestions/Doubts/Critics }
  	\begin{center}
    {\huge ?}\\
    \includegraphics[width=.5\textwidth]{figures/rodin.png}\\
    \large{ If you prefer, you can send an e-mail to 
    %\textcolor{blue!70!black}{\textit{valerio}}@ppgsc.ufrn.br or
    \textcolor{blue!70!black}{\textit{valerio.jr}}@gmail.com \\
    Project available at http://code.google.com/p/b2asm
    }
	\end{center}

\end{frame}