### 2012-01-17 trunk/papers/ArticleZ80/source/paper.tex
Several updates.

\documentclass[a4paper]{llncs}

%\documentclass[11pt]{report} % use larger type;

\usepackage[utf8]{inputenc} % set input encoding (not needed with XeLaTeX)

%%% PAGE DIMENSIONS
\usepackage{geometry} % to change the page dimensions
\usepackage{graphicx,array}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsfonts}
\usepackage{epsfig}
\usepackage{float}
%\usepackage{supertabular} % used in B Symbols
\geometry{a4paper} % or letterpaper (US) or a5paper or....

\usepackage[english]{babel} %\usepackage{supertabular} % used in B Symbols
\usepackage{hyperref} % add URL in bibliography
\usepackage{url} % allow add URL in document

\include{comandos}
\include{SymbolsB_AtelierB}

% Allow use multiples footnote references 
\newcommand{\footnoteremember}[2]{
  \footnote{#2}
  \newcounter{#1}
  \setcounter{#1}{\value{footnote}}
}
\newcommand{\footnoterecall}[1]{
  \footnotemark[\value{#1}]
}

%
\begin{document}

\title{Experiences in Modelling a Microcontroller Instruction Set Using B}
%
\hyphenation{se-pa-ra-te}
\hyphenation{in-struc-tion}
\hyphenation{re-gis-ter}
\hyphenation{re-gis-ters}
%
%
%
\author{Val\'{e}rio Medeiros Jr\inst{1}, David D\'{e}harbe\inst{2}}

\institute{ Federal Institute of Education, Science and Technology of Rio Grande do Norte, Natal RN 59015-000, Brazil
           \and 
		    Federal University of Rio Grande do Norte, Natal RN 59078-970, Brazil}

\maketitle

\begin{abstract}

  This paper describes an approach to model the functional aspects of
  the instruction set of microcontrollers and several details about
  the representation of microcontrollers.  Different models were
  developed using the notation of the B method. This paper presents
  specifically a B-based formal model of the Z80 instruction set.

%  \paragraph[Keywords]{
%   \textbf{Keywords}: Embedded Software, Simulation, B method and Verification%}
\end{abstract}

%\keywords{ Validation, Verification, Microcontroller, Hardware}
%
%\setcounter{secnumdepth}{1}
%\setcounter{chapter}{1}
%\setcounter{section}{1}

\section{Introduction}

The B method~\cite{Abrial} supports the formal development of software
from a specification of functional requirements down to an imperative
implementation of those requirements, amenable to synthesis to a
programming language such as C or Ada. However, the synthesis step is
not amenable to proof of correctness, and~\cite{Dantas_SBMF08} proposed
an approach to extend the scope of the B method up to the assembly
level language. One key component of this approach is to build, within
the framework of the B method, formal models of the instruction set of
such assembly languages.

This work presents a formal modell of the instruction set of the Z80 
microcontroller~\cite{Z80_manual}. This microcontroller
was chosen because of several factors: it contains essential and common
concepts in microcontrollers and until microprocessors; it has extensive
documentation available; and it was widely used and is still in use. 
A first Z80 modell presented in \cite{Valerio_SBMF09} was verified completely,
but it uses constructs that are not well supported to animate. 
This modelling was changed to support animation in ProB\cite{proB}.
Basically, the new modelling\footnote{The interested reader in more
details is invited to visit our repository at: http://code.google.com/p/b2asm/.}
changed elements represented by infinite sets and adjusted the implications
of this change. Using the responsibility division mechanisms provided by B, auxiliary
libraries of basic modules were developed as part of the construction of microcontroller
model. Such library has many definitions about common concepts used in
the microcontrollers;besides the new Z80 model, a theoretic case study in 
petroleum production test system was developed using Z80 assembly language
to analyse the code simulation and verification process.

The formal model of instruction set provides many benefits. The animation of 
Z80 model allows to simulate the execution assembly programs including
support for interrupts and input and output instructions. Other possible uses
of a formal model of a microcontroller instruction set include documentation,
the construction of simulators, and can be possibly the starting point of a 
verification effort for the actual implementation of a Z80 design. Moreover, 
the model of the instruction set could be instrumented with non-functional
aspects, such as the number of cycles it takes to execute an instruction, to
prove lower and upper bounds on the execution time of a routine.
Two aspects of these formalization are of particular importance to us:
provide a more solid documentation artifact for microcontrollers, 
build a reference model for a formal compilation approach in B.

\subsection{Problems in formalizing instruction set}
% Compact version 
In general, the manuals of microcontrollers can be more suited for
developers in assembly language. Many times, each instruction of microcontrollers
is shown in its official manual using textual description, math description, examples,
encoding, number of cycles and other information. Some times, the instruction descriptions
have an own notation, some informations are organized in different sections of document 
and the textual description are not standardized. This does not allow the interpretation by parser.
The manuals also have semi-formal and informal descriptions, these descriptions may make mistakes.
The official manual of the microcontroller Z80 has also several problems that were identified
over time~\cite{UndocumentedZ80}.

	A solution to avoid ambiguity and inconsistencies is to specify formally the
assembly instruction set, this creates a pattern of representation and validates
properties of instructions. The formal model also restricts the definitions to
correct typing and uses only well-defined expressions. Furthermore, the developer
can use with B method different levels of abstraction and add more specific details
in refined models.

%In general, the manuals of microcontrollers can be more suited for
%developers in assembly language. Many times, each instruction of microcontrollers
%is shown in its official manual using textual description, math description, examples,
%encoding, number of cycles and other information.Some times, the instruction descriptions
%have an own notation, some informations are organized in different sections of document 
%and the textual description are not standardized. This does not allow the interpretation by parser.
%% b) The case of Z80
%For example in official manual of the microcontroller Z80 \cite{Z80_manual}, if the users
%want to know the action of an instruction and its effects on the flag register then he needs
%to read an informal textual description. So, this description can add
%some ambiguities or even mistakes. The official manual has also several problems that were 
%identified over time~\cite{UndocumentedZ80}. The Z80 is a microcontroller that  has been
%tested and used for many years. This intensive use and test facilitated to find several
%errors in the description of the instructions in its manual. Several of these problems
%have been reported in a technical report \cite{UndocumentedZ80}. Some problems are inaccurate
%information, for example the description of the action of an instruction on the flag
%register, partial information and distributed on different pages of manual.
%
%
%% in the Z80 manual happens was that,
%
%%A simple example of error is in some official documentation saying that a group of instructions leaves the \textit{CF} flag unaffected and the \textit{NF}
%%flag is not always set, but may also be reset \cite{UndocumentedZ80}.
%%% \begin{itemize} \item Some input and output instructionsThe
%%% INI/INIR/IND/INDR/OUTI/OUTD/OTIR/OTDR instructions do affect the CF flag (some
%%% official documentation says they leave it unaffected, important!) and the NF
%%% flag is not always set but may also be reset (see 4.3 for exact operation)
%%% \cite{UndocumentedZ80}. Este erro he descrito na phegina (274 do documento e
%%% 294 do pdf) do manual oficial do Z80 (Z80_CPU_adotado.pdf) Citar apenas esse
%%% exemplo abaixo! \item
%%%when a determined interruption (NMI) is accepted, the bit IFF1 is not   
%%%copied to bit IFF2~\cite{UndocumentedZ80}. 
%%%
%%This example can cause serious problems for programmers.
%%% Este erro he descrito na pagina (22 do documento e 26do pdf) do manual
%%% official do Z80 (Z80_CPU_adotado.pdf) Only IFF1 is reset.
%%This error and others 7 different errors were previously identified by~\cite{UndocumentedZ80}. So to help the 
%%construction of Z80 B model, other references~\cite{UndocumentedZ80,Simulator_z80}
%%were also consulted.
%
%A solution to avoid ambiguity and inconsistencies is to specify formally the
%assembly instruction set, this creates a pattern of representation and validates
%properties of instructions. The formal model also restricts the definitions to
%correct typing and uses only well-defined expressions. Furthermore, the developer
%can use with B method different levels of abstraction and add more specific details
%in refined models.
% 
%Finally, the main goal of this project, though, is to provide a basis for the
%generation of software artefacts at the assembly level, those are amenable to 
%refinement verification within the B method.

%\subparagraph{Report structure}
%
%This work is focused on the presentation of modelling of assembly instruction
%set, including elementary libraries to describe hardware aspects; and the
%modelling and simulation of a case study of an embedded verified software that
%can be used in tanks of the petroleum industry.

This paper is structured as it follows. %Section~\ref{sec:B_method} provides a short introduction to the B method. 
Section~\ref{sec:models} presents the elementary libraries and
the modelling of some elements common to microcontrollers.  Section~\ref{sec:z80}
presents the B model of the Z80 instruction set. %Section ~\ref{sec:Verification} provides some information on the proof effort needed to analyse the presented models.
%Section ~\ref{sec:studycase} explains the building and verifying assembly-level refinements. 
Related works are discussed in section~\ref{sec:relatedworks}. Finally, the last section is devoted to the conclusions.
%
%\section{Introduction to the B Method}
%\label{sec:B_method}
%
%The B method for software development~\cite{Abrial} is based on the B Abstract
%Machine Notation (AMN) for the formal specification of functional requirements and the use of formally proved refinements up to a
%specification sufficiently concrete that programming code can be automatically
%generated from it. Its mathematical basis consists of first order logic, integer
%arithmetic and set theory, and its corresponding constructs are similar to those
%of the Z notation~\cite{Spivey96}.
%
%A B specification is structured in modules. A module defines a set of valid
%states, including a set of initial states, and operations that may provoke a
%transition among states. The design process starts with a module with a
%so-called functional model of the system under development. In this initial
%modelling stage, the B method requires that the user proves that, in a machine,
%all its initial states are valid, and operations do not define
%transitions from valid states to invalid states.
%
%
%Essentially, a B module contains two main parts: a header and the available
%operations. Figure~\ref{fig:maqB} has a very basic example. The clause
%\MACHINE  has the name of module.  The next two clauses, respectively,
%reference external modules and create an instance of an external module.  The
%\VARIABLES clause declares the name of the variables that compose the
%state of the machine. Next, the \INVARIANT clause defines the valid states,
%specifying  the type and other restrictions on the variables. The \INITIALISATION specifies
%the initial states. Finally, operations correspond to the transitions among
%states of the machine.
%
%
%\begin{figure}[h]
%  % \begin{small}
%\begin{center}
%\begin{sloppypar}
%
%  $\begin{array}{lcl}
%    \begin{array}[t]{l}
%      \MACHINE \quad  \textit{micro} \\
%      \SEES \quad \textit{TYPES}, \textit{ALU} \\
%      \INCLUDES  \quad \textit{MEMORY} \\
%      \VARIABLES \quad    \textit{pc}  \\
%      \INVARIANT\\ 
%      \quad \textit{pc} \in \textit{INSTRUCTION} 
%    \end{array}
%    & \hspace*{0cm} &
%    \begin{array}[t]{l}
%      \INITIALISATION  \textit{pc} := 0 \\
%      \OPERATIONS\\
%      \textit{JMP} \textit{( jump )} = \\
%      \quad \PRE \textit{jump} \in \textit{INSTRUCTION}\\ 
%      \quad \THEN \textit{pc} := \textit{jump}\\  
%      \quad \END \\
%      \END
%    \end{array}
%  \end{array}
%  $
% \end{sloppypar} 
%\end{center}
%\caption{A very basic B machine.}
%\label{fig:maqB}
%\end{figure}

\section{Model structure and basic components}
\label{sec:models}
%Em cada subsecaoo descrever os principais lemas que foram provados.
%Sugestheo de David:
% a) structure
% b) formalization of bit-level concepts
% c) formalization of bit vectors
% d) formalization of byte and word-level concepts
% e) formalization of integer ranges

We have developed a reusable set of basic definitions to model hardware concepts and data types
concepts. The models contains definitions to represent the: registers, interruptions, input and output
ports, memory and instructions. These definitions are grouped into separated development projects
and are available as libraries.

Thus the workspace is composed of modules group and libraries: 
\begin{itemize}

\item \textbf{Power}: it has the basics definitions to help the theorems
prover about simple calculus of power, especially power of two. It contains the definition of power function
and it has constants, that represent all required results of power function.

\item \textbf{Hardware Library}: it has the definition of bit vectors with size 8 and 16, basic functions to manipulate bit vectors and important assertions. This is presented in sections \ref{subsec:HardwareLibrary1}, \ref{subsec:HardwareLibrary2} and \ref{subsec:HardwareLibrary3}.

 \item \textbf{Types Library}: it has the definition of naturals and integers represented with  8 bits and 16 bits, basic conversion functions between the types and bit vectors and important assertions. This is presented in section \ref{subsec:Types}.

 \item \textbf{Platform}: it has the definition functions of arithmetic logic unit,  memory unit, registers and assembly instructions. This is presented in section \ref{sec:z80}.

\end{itemize}

 The corresponding dependency
diagram is depicted in Figure~\ref{fig:hardware-definition-graph}; information specific to each project is
presented in the following.

\begin{figure}[h]
\centering
\includegraphics[width=.77\textwidth]{diagramaEstrutural_vertical_ProB.png}
 \caption{Dependency diagram of the Z80 model.}
\label{fig:hardware-definition-graph}
\end{figure}

\subsection{Bit representation and manipulation}
\label{subsec:HardwareLibrary1}

The entities defined in the module $\textit{BIT\_DEFINITION}$ are the
type for bits, logical operations on bits (negation, conjunction,
disjunction, exclusive disjunction), as well as a conversion function
from Boolean to bit.

First, bits are modelled as a set of integers: $\textit{BIT} =
\textit{0..1}$. The negation is an unary function on bits and it is
defined as:

$
\begin{array}{l}
\textit{bit\_not}  \in  \textit{BIT}  \fun  \textit{BIT}  \land 
\textit{bit\_not} \rm = \rm \{ \rm 0  \mapsto  \rm 1 , \rm 1  \mapsto  \rm 0 \rm \}
\end{array}
$
%\forall ( \textit{bb}). (\textit{bb} \in \textit{BIT} \implies {bit\_not}(\textit{bb}) =1-\textit{bb})\\

The module also provides lemmas on negation that may be useful for the
users of the library to develop proofs:

$
\begin{array}{l}
%  \textit{bit\_not}(0) = 1;  \textit{bit\_not}(1) = 0; \\
\forall (\textit{bb}).(\textit{bb} \in \textit{BIT} \implies \textit{bit\_not}(\textit{bit\_not}(\textit{bb})) = \textit{bb})
\end{array}
$

Conjunction is a binary function on bits and it is defined as:

$
\begin{array}{l}
\textit{bit\_and} \in \textit{BIT} \times \textit{BIT} \fun \textit{BIT} \land \\
\forall (\textit{b1}, \textit{b2}).(\textit{b1}  \in \textit{BIT}  \land \textit{b2} \in \textit{BIT} \implies \\
\quad ((\textit{bit\_and}(\textit{b1}, \textit{b2}) = 1) \iff (\textit{b1} = 1)  \land  (\textit{b2} = 1)))
\end{array}
$

The module provides the following lemmas for commutativity and associativity:

$
\begin{array}{l}
%  \textit{bit\_and}(0,0) = 0;  \textit{bit\_and}(0,1) = 0; \\
%  \textit{bit\_and}(1,0) = 0;  \textit{bit\_and}(1,1) = 1; \\
\forall (\textit{b1},\textit{b2}).(\textit{b1} \in \textit{BIT} \land \textit{b2} \in \textit{BIT} \implies \\
\quad (\textit{bit\_and}(\textit{b1}, \textit{b2}) = \textit{bit\_and}(\textit{b2},\textit{b1})))\land \\
\forall (\textit{b1},\textit{b2},\textit{b3}).(\textit{b1} \in \textit{BIT} \land  \textit{b2} \in \textit{BIT} \land \textit{b3} \in \textit{BIT} \implies \\
\quad (\textit{bit\_and}(\textit{b1}, \textit{bit\_and}(\textit{b2},\textit{b3})) = \textit{bit\_and}(\textit{bit\_and}(\textit{b1},\textit{b2}),\textit{b3})))\\
% \forall (\textit{b1}).(\textit{b1} \in \textit{BIT} \implies (\textit{bit\_and}(\textit{b1}, 1) = \textit{b1})); \\
% \forall (\textit{b1}).(\textit{b1} \in \textit{BIT} \implies (\textit{bit\_and}(\textit{b1}, 0) = 0));
\end{array}
$

The module provides definitions of $\textit{bit\_or}$ (disjunction)
and $\textit{bit\_xor}$ (exclusive disjunction), as well as lemmas on
those operators. These are standard and their expression in B is
similar as for $\textit{bit\_and}$, they are thus omitted from this paper.

Finally, the conversion from Boolean to bit is simply defined as:

$
\begin{array}{l}
\textit{bool\_to\_bit} \in \BOOL \fun \textit{BIT} \land \textit{bool\_to\_bit} = \{ \TRUE \mapsto 1, \FALSE \mapsto 0 \} \\
\end{array}
$

Observe that all the lemmas that are provided in this module have been
mechanically proved by the theorem prover included with our B
development environment. None of these proofs requires human insight.

\subsection{Representation and manipulation of bit vectors}
\label{subsec:HardwareLibrary2}

Sequences are pre-defined in B, as functions whose the domain is an
integer range with lower bound 1 (one). Indices in bit vectors usually
range from 0 (zero) upwards and the model we propose obeys this
convention by making an one-position shift where necessary. This shift
is important to use the predefined functions of sequences. We thus
define bit vectors as non-empty sequences of bits, and
$\textit{BIT\_VECTOR}$ is the set of all such sequences:
$\textit{BIT\_VECTOR} = \seq1(\textit{BIT})$. The $\textit{BIT\_VECTOR}$ is an infinite set and it hinders the animation in ProB. 
To avoid this obstruction, the specialized types (\textit{BYTE} and \textit{BV16}) cannot reference directly \textit{BIT\_VECTOR}.

% CAN COMMENT TO REDUCE SPACE
%The function $\textit{bv\_size}$ returns the size of a given bit vector. It is basically a wrapper for the
%predefined function $\textbf{size}$ that applies to sequences.
%
%$
%\begin{array}{l}
%\textit{bv\_size} \in \textit{BIT\_VECTOR} \fun \nat_1 \land \\
%\textit{bv\_size} = \lambda bv . (bv \in \textit{BIT\_VECTOR} \mid \textbf{size}(bv))
%\end{array}
%$

We also define two functions $\textit{bv\_set}$ and $\textit{bv\_clear}$ that, given a bit vector, and a
position of the bit vector, return the bit vector resulting from setting the corresponding position to 0
or to 1, and a function $\textit{bv\_get}$ that, given a bit vector, and a valid position, each one
returns the value of the bit at that position. Only the first definition is shown here:

$
\begin{array}{l}
\textit{bv\_set} \in \textit{BIT\_VECTOR} \times \nat \fun \textit{BIT\_VECTOR} \land \textit{bv\_set} =\\
\lambda v, n . (v \in \textit{BIT\_VECTOR} \land n \in \nat \land n <\textit{size}(v)
\mid v \lover \{ n+1 \mapsto 1 \})
\end{array}
$

The function $bv\_catenate$ takes as parameters two bit vectors $v$ and $w$, and returns the result of the
concatenation of $v$ and $w$, such that $v$ constitutes the most significant part of the result.

% $
% \begin{array}{l}
% \textit{bv\_catenate} \in \textit{BIT\_VECTOR} \times \textit{BIT\_VECTOR} \fun \textit{BIT\_VECTOR} \land \\
% \textit{bv\_catenate} = \lambda v, w \bullet (v \in \textit{ BIT\_VECTOR} \land w \in \textit{ BIT\_VECTOR}  \mid v 
% %\conc
%   w)
% \end{array}
% $

\hspace*{0.00in} \it bv\_catenate  $\in$  \it BIT\_VECTOR  $\times$  \it BIT\_VECTOR  $\fun$ \it
BIT\_VECTOR $\land$\\
\hspace*{0.21in} \it bv\_catenate \rm =  $\lambda$  v\rm ,\it w \rm . \rm (\it v 
$\in$  \it BIT\_VECTOR $\land$ \it w $\in$  \it BIT\_VECTOR  $\mid$  \it v\^ \it w\rm ) %$\cat$

%COMMENT TO REDUCE SPACE
%We also define a function $\textit{bv\_zero}$ that, given a positive
%integer $n$, return a bit vector of size $n$, with all bits set to 0.
%A similar function, called $\textit{bv\_one}$, with all bits set to 1
%is also defined but not presented here.
%
%$
%\begin{array}{l}
%\textit{bv\_zero} \in \nat_1 \fun \textit{BIT\_VECTOR} \land \\
%\textit{bv\_zero} = \lambda n . (n \in \nat_1 \mid 1..n \times \{0\}) 
%\end{array}
%$
%

Additionally, the module provides definitions for the classical
logical combinations of bit vectors: $\textit{bit\_not}$,
$\textit{bit\_and}$, $\textit{bit\_or}$ and $\textit{bit\_xor}$. Only
the first two are presented here. Observe that the domain of the
binary operators is restricted to pairs of bit vectors of the same
length:

$
\begin{array}{l}
\textit{bv\_not} \in \textit{BIT\_VECTOR} \fun \textit{BIT\_VECTOR} \land \\
\textit{bv\_not} = \lambda v . (v \in \textit{BIT\_VECTOR} \mid \quad \lambda i . (1 .. \textit{size}(v)) \mid \textit{bit\_not}(v(i))) \land \\
\textit{bv\_and} \in \textit{BIT\_VECTOR} \times \textit{BIT\_VECTOR} \fun \textit{BIT\_VECTOR} \land \\
\textit{bv\_and} = \lambda v_1, v_2 . (v_1 \in \textit{BIT\_VECTOR} \land v_2 \in \textit{BIT\_VECTOR} \land \\
\textit{size}(v_1) = \textit{size}(v_2) \mid \lambda i . (1 .. \textit{size}(v_1)) \mid
\textit{bit\_and}(v_1(i), v_2(i)))
\end{array}
$

We provide several lemmas on bit vector operations. These lemmas
express properties on the size of the result of the operations
as well as classical algebraic properties such as associativity
and commutativity.

\subsection{Modelling bytes and bit vectors of length 16}
\label{subsec:HardwareLibrary3}

Bit vectors of length 8 are bytes. They are a common entity in
hardware design. We provide the following definitions:

\hspace*{0.0in}\it BYTE\_WIDTH \rm = 8 $\land$ \it BYTE\_INDEX \rm = 1 $\upto$ \rm  BYTE\_WIDTH\rm  \hspace*{0.03in} $\land$

\hspace*{0.0in}\it PHYS\_BYTE\_INDEX \rm = \rm 0 $\upto$ \rm (\it BYTE\_WIDTH\rm -\rm 1\rm )\hspace*{0.03in} $\land$ \hspace*{0.0in}\it BYTE \rm =\rm (\it BYTE\_INDEX  $\fun$  \it BIT\rm )\hspace*{0.0in} $\land$

\hspace*{0.0in}$\forall$ \rm (\it b1\rm )\rm .\rm (\it b1  $\in$  \it BYTE  $\implies$  \bf size\rm (\it b1\rm )\rm =\it BYTE\_WIDTH  $\land$  \it b1  $\in$  \bf seq1\rm (\it BIT\rm )\rm )

%\hspace*{0.0in}\it BYTE \rm = \rm \{ \it bt  $\mid$  \it bt $\in$ \it BIT\_VECTOR  $\land$  \it size\rm (\it bt\rm )\rm =\it BYTE\_WIDTH\rm \} $\land$

%\hspace*{0.0in}\it BYTE\_ZERO  $\in$  \it BYTE  $\land$ \it BYTE\_ZERO \rm = \it BYTE\_INDEX  $\times$  \rm \{\rm 0\rm \}

The $\textit{BYTE\_INDEX}$ is the domain of the functions modelling bytes. It starts at 1 to obey a
definition of sequences from B. However, it is common in hardware architectures to start indexing from
zero. The definition $\textit{PHYS\_BYTE\_INDEX}$ is used to provide functionalities obeying this
convention. The $\textit{BYTE}$ type is as a specialized type of $\textit{BIT\_VECTOR}$, but it has a size
limit. Other specific definitions are provided to facilitate further modelling: the type $\textit{BV16}$
is created for bit vector of length 16 in a similar way.

\subsection{Bit vector arithmetic}
\label{subsec:Types}

Bit vectors are used to represent and combine numbers: integer ranges (signed or unsigned). Therefore, our
library includes functions to manipulate such data, for example, the function $\textit{bv\_to\_nat}$ that
maps bit vectors to natural numbers:

$
\begin{array}{l}
\textit{bv\_to\_nat} \in \textit{BIT\_VECTOR} \fun \nat  \land \\
\textit{bv\_to\_nat} = \lambda v .  (v \in \textit{BIT\_VECTOR} \mid  \sum i . (i \in \dom(v) . v(i)
\times 2^{i-1}))
\end{array}
$

An  associated lemma is: $\forall n . (n \in \nat_1 \implies \textit{bv\_to\_nat}(\textit{nat\_to\_bv}(n)) = n)$

\subsection{Basics data types}

The instruction set architecture usually have common bit vector type and integer type.
In table~\ref{tab:types}, the first three are placed in the hardware library and the last four types are placed in the integer types library. 
Each type module has functions to manipulate and convert its data. There are seven common basic data types represented by modules.

\begin{table}
\caption{Details of basic data types}
\label{tab:types}
\begin{center}
\begin{tabular}{|c|c|c|c|}
\hline
 $Type\ Name$ & $Range$ & $Physical\ Size $  & $Associated\ Proof\ Obligations$ \\\hline
 BIT & 0..1 & 1 bit &   118 \\\hline %49 + 69 = 118 \\\hline
 BYTE & -- & 1 bytes &  153\\\hline%17 + 136 = 153\\\hline
 BV16 & -- & 2 bytes &  75 \\ \hline%5 + 70 = 75 \\ \hline
 UCHAR & 0..255 & 1 byte &  30 \\\hline %4 + 26 = 30 \\\hline
 SCHAR & -128..127 & 1 byte & 26 \\\hline %2 + 24 = 26 \\\hline
 USHORT & 0..65.535 & 2 byte & 62 \\\hline % 4 + 58 = 62 \\\hline
 SSHORT & -32.768..32.767 & 2 byte & 58\\\hline %  2 + 56 = 58\\\hline
%  &  & Total & 522\\\hline %  2 + 56 = 58\\\hline
\end{tabular}
\end{center}
\end{table}

Usually, each type module just needs to instantiate concepts that were already defined in the hardware
modelling library.  For example, the function $\textit{bv\_to\_nat}$ from bit vector arithmetic is
specialized to $\textit{byte\_uchar}$. As the set $\textit{BYTE}$ is a subset of the
$\textit{BIT\_VECTOR}$, this function can defined as follows:

$
\begin{array}{l}
\textit{byte\_uchar} \in \textit{BYTE} \fun \nat \land \\
\textit{byte\_uchar} = \lambda (v) . ( v \in BYTE | bv\_to\_nat(v) )
\end{array}
$

The definitions of the library types reuse the basic definitions from the hardware library. This provides
greater confidence and facilitates the proof process, because the prover can reuse the previously defined
lemma.

%The inverse function $\textit{uchar\_byte}$ is easily defined:
%
%$
%\begin{array}{l}
%\textit{uchar\_byte} \in \textit{UCHAR}  \fun  \textit{BYTE}  \land \\
%   \textit{uchar\_byte} = \ (\textit{byte\_uchar}) ^{-1}
%\end{array}
%$

 We also created the following lemmas:

 $
 \begin{array}{l}
  \forall (val) . (val \in \textit{UCHAR} |
  \textit{byte\_uchar}(\textit{uchar\_byte}(val)) = val) \land\\
  \forall (by) . (by \in \textit{BYTE} |
  \textit{uchar\_byte}(\textit{byte\_uchar}(by)) = by)
 \end{array}
 $

Similarly, several other general functions and lemmas were created for all other data types.
%[NEW PARAGRAPH-TO-REVIEW]
%Similarly, several other general functions and lemmas were created for all other data types.
%Some times, we need create arithmetic and logic functions that are more specific for a
%determined computer architecture. These functions are developed in the module ALU 
%(arithmetic  logic unit). The following shows some examples of specific functions used in
%microcontroller model of Z80. For example, the function $\textit{instruction_next}$
%receives a $\textit{BYTE}$ and returns a $\textit{BYTE}$ that contains a sequence of bit:
%a signal bit,

Nearly 50\% of all these proof obligations shown in the table~\ref{tab:types} 
are verified almost automatically. Although many others proof obligations are difficult to verify,
because they are functions of type conversion and contain arithmetic expressions difficult 
to manipulate with the interactive prover of AtelierB.

\subsection{Verification in Hardware Library and Types Library} 
\label{sec:VerificationHardwareLibrary}

\input{VerificationLibraryHardware}

\section{A B model of the Z80 instruction set}
\label{sec:z80}

The \textit{Z80} is a CISC microcontroller developed by
\textit{Zilog}~\cite{Z80_manual}. The Z80 is composed by different elements and a
simplified internal organization is shown in the Figure \ref{fig:DiagramBlock}.
This figure has some important elements of Z80 CPU: ALU, registers of 8 and 16
bits and input/output ports.

\begin{figure}[h] \centering
\includegraphics[width=0.70\textwidth]{images/Architecture.png}
\caption{Simplified internal organization of Z80 CPU.}
\label{fig:DiagramBlock}
\end{figure}

The Z80 has 158 different instructions, including all the 78 from Intel 8080
microprocessor, and all of them were formally specified in B. These instructions are classified
into these categories: load and exchange; block transfer and search; arithmetic
and logical; rotate and shift; bit manipulation; jump, call and return;
input/output; and basic cpu control. Each category of instruction has different
elements of specification.

This section shows the elements that make up the different types of instructions.
The main elements are specified in the microcontroller module Z80 and parts of it
are presented below.  Basically, the state of microcontroller is formed by data 
from program counter, ports, registers, etc. The transitions between the states 
are actioned by execution of instruction or an external action. Groups of registers
are represented by the variables of the specification states that appear in clause
\textit{VARIABLES}. The declaration of valid states of variables is
in the \textit{INVARIANT} clause and the initial state is defined in the
\textit{INITIALISATION} clause. The assembly instructions are defined by the
clause \textit{OPERATIONS} and, in general, each instruction has three actions: update 
the program counter, update the flag register and its main effect.
%Several functions were created to Z80.
General functions that can be used in others microcontroller models are defined in
the modules of data types. Specific functions of the microcontroller are defined in the ALU (arithmetic logic unit) module.

The main module includes an instance of the memory module and accesses the definitions from basic data
types modules and the \textit{ALU} module.

	\begin{center}
	\begin{sloppypar}

	  \begin{tabbing}
	    \begin{array}[t]{l}
		\bf MACHINE\\
		\hspace*{0.15in}\it Z80\\
		\bf INCLUDES\\
		\hspace*{0.10in}\it MEMORY
	    \end{array}
	     \hspace*{1cm} 
	    \begin{array}[t]{l}
	    \bf SEES\\
		\hspace*{0.10in}\it ALU, \it BIT\_DEFINITION, \it BIT\_VECTOR\_DEFINITION,\\
		\hspace*{0.10in}\it BYTE\_DEFINITION, \it BV16\_DEFINITION,\\
		\hspace*{0.10in}\it UCHAR\_DEFINITION, \it SCHAR\_DEFINITION,\\
		\hspace*{0.10in}\it SSHORT\_DEFINITION ,\it USHORT\_DEFINITION\\
	    \end{array}
	  \end{tabbing}

	 \end{sloppypar} 
	\end{center}

\subsection{Modelling registers}

The Z80 CPU includes alternative set of accumulator, flag and general registers. The CPU contains a stack
pointer ($\textit{sp}$), program counter ($\textit{pc}$), two index registers ($\textit{ix}$ and $\textit{iy}$), an
interrupt register ($\textit{i\_}$), a refresh register ($\textit{r\_}$), two bits ($\textit{iff1}$,
$\textit{iff2}$) used to control the interruptions, a pair of bits to define the interruption mode ($\textit{im}$)
and the input and output ports ($\textit{i\_o\_ports}$). These definitions are represented by
\textit{INVARIANT}.

\begin{sloppypar}
\bf INVARIANT

\hspace*{0.10in}\it rgs8  $\in$  \it id\_reg\_8  $\fun$  \it BYTE  $\land$ \it pc  $\in$  \it INSTRUCTION  $\land$  \it sp  $\in$  \it BV16 $\land$  \it ix  $\in$  \it BV16  $\land$

\hspace*{0.10in}\it iy  $\in$  \it BV16  $\land$ \it i\_  $\in$  \it BYTE  $\land$  \it r\_ $\in$  \it BYTE  $\land$ \it iff1  $\in$  \it BIT  $\land$ \it iff2  $\in$  \it BIT  $\land$

\hspace*{0.10in}\it im $\in$ (\it BIT $\times$ \it BIT\rm )  $\land$  \it i\_o\_ports  $\in$  \it BYTE  $\fun$  \it BYTE
\end{sloppypar}

%

% Features of microcontroller
  The internal registers contain 176 bits of read/write memory that are represented by
identifiers used as parameters in the instructions . It includes two sets of six general purpose
registers which may be used individually as 8-bit registers or as 16-bit register pairs.  The working
registers are represented by variable $\textit{rgs8}$. The domain of $\textit{rgs8}$
($\textit{id\_regs8}$) is a set formed by identifiers of registers of 8 bits. These registers are called
main register set (\textit{a0,b0,c0,d0,e0,f0,h0,l0}) and alternate register set
(\textit{a\_0,b\_0,c\_0,d\_0,e\_0,f\_0,h\_0,l\_0}). These registers can be accessed in pairs, forming
16-bits, resulting in another set of identifiers of 16-bits registers, named $\textit{id\_reg16}$. These sets are represented below.

\begin{sloppypar}

	\bf SETS\\
	\hspace*{0.10in}\it id\_reg\_8 \rm = \rm \{ \it a0 \rm , \it f0 \rm ,\it b0 \rm , \it c0 \rm ,\it d0 \rm , \it e0 \rm ,\it h0 \rm , \it l0 \rm ,\it a\_0 \rm , \it f\_0 \rm ,\it b\_0 \rm , \it c\_0 \rm ,\it d\_0 \rm , \it e\_0 \rm ,\it h\_0 \rm , \it l\_0 \rm \};\\
	\hspace*{0.10in}\it id\_reg\_16 \rm = \rm \{ \it BC \rm , \it DE \rm , \it HL \rm , \it SP \rm , \it AF \rm \}

\end{sloppypar}

%\begin{sloppypar}
%
%	\bf SETS\\
%	\hspace*{0.10in}\it id\_reg\_8 \rm = \rm \{ \it a0 \rm , \it f0 \rm , \it f\_0 \rm , \it a\_0 \rm ,\\
%	\hspace*{0.93in}\it b0 \rm , \it c0 \rm , \it b\_0 \rm , \it c\_0\rm,\\
%	\hspace*{0.93in}\it d0 \rm , \it e0 \rm , \it d\_0 \rm , \it e\_0\rm,\\
%	\hspace*{0.93in}\it h0 \rm , \it l0 \rm , \it h\_0 \rm , \it l\_0\};\\
%	\hspace*{0.10in}\it id\_reg\_16 \rm = \rm \{ \it BC \rm , \it DE \rm , \it HL \rm , \it SP \rm , \it AF \rm \}
%\end{sloppypar}

The main working register of Z80 is the accumulator ($\textit{rgs8(a0)}$) used for arithmetic/logic, input/output and loading/storing operations.
Another important element is the ``f'' register ($\textit{rgs8(f0)}$),  that is
used as a \textbf{flag register}. This register uses only six bits to represent the execution result status of each instruction.
According to the official manual the bits 3 and 5 are not used and the others bits have the follow meaning: $\textit{bv\_get(rgs8(f0),0)}$ - the carry bit; $\textit{bv\_get(rgs8(f0),1)}$ - the add/subtract bit; $\textit{bv\_get(rgs8(f0),2)}$ - the parity or overflow bit; $\textit{bv\_get(rgs8(f0),4)}$ - the half carry bit; 
$\textit{bv\_get(rgs8(f0),6)}$ - the zero bit; and $\textit{bv\_get(rgs8(f0),7)}$ - the sign bit.
%\begin{description}
%  \item[$\textit{bv\_get(rgs8(f0),0)}$] - The Carry bit
%  \item[$\textit{bv\_get(rgs8(f0),1)}$] - The Add/Subtract bit
%  \item[$\textit{bv\_get(rgs8(f0),2)}$] - The Parity or Overflow bit
%  \item[$\textit{bv\_get(rgs8(f0),4)}$] - The Half Carry bit
%  \item[$\textit{bv\_get(rgs8(f0),6)}$] - The Zero bit
%  \item[$\textit{bv\_get(rgs8(f0),7)}$] - The Sign bit 
%\end{description}
These bits also can be used to specify security properties for microcontroller programs.

\textbf{Assuring the absence of overflow:}
 \emph{To assure that an overflow does not happen, the developer can add this
 expression (\textit{bv\_get(rgs8(f0),0)} $\neq$ 1 $\land$
 \textit{bv\_get(rgs8(f0),2) $\neq$ 1}) in the invariant.
% When a overflow happen,it can be dangerous. So the developer may restrict its use. However, this restriction can also
% become more difficult to verify the model.
}

\subsection{Manipulation data functions from Z80}

 There are some specific functions from Z80 to manipulate the data. In addressing mode, the function
$\textit{bv\_ireg\_plus\_d}$ receives the value of a register ($\textit{ix}$ or $\textit{iy}$) 
used as a base to point to a region in memory from which data is to be stored or retrieved.
%and the displacement to return the sum, the result is the dislocated address memory, 
 An additional byte is included in indexed instructions to specify an offset from this base. This displacement is specified as a two's complement signed integer. This function is defined as:

\hspace*{0.0in}\it bv\_ireg\_plus\_d \rm : \rm(\it BV16  $\times$  \it SCHAR  $\fun$  \it BV16\rm )  $\land$

\hspace*{0.0in}\it bv\_ireg\_plus\_d \rm =  $\lambda$  \rm ( \it ix\_iy \rm , \it offset \rm ) \rm . \rm ( \it ix\_iy  $\in$  \it BV16  $\land$  \it offset  $\in$  \it SCHAR

\hspace*{0.20in}$\mid$ \it ushort\_bv16 \rm ( \rm (\it bv16\_ushort \rm ( \it ix\_iy \rm ) \rm + \it offset \rm ) 
\textbf{\textit{mod}} $2^{16}$ ) \rm )

Another derived function is  $\textit{bv\_(ireg\_plus\_d)}$, this returns the value in the memory  address
returned by $\textit{bv\_ireg\_plus\_d}$ function and its definition is similar.

There is a specific function to refresh the flag register, it is named $\textit{update\_reg\_flag}$. It is typed and defined as follow:\it update\_flag\_reg \rm $\in$ \rm (\it BIT  $\times$  \it BIT  $\times$  \it BIT  $\times$  \it
BIT $\times$  \it BIT  $\times$  \it BIT\rm) $\fun$  \rm (\rm \{\it f0\rm \}  $\times$  \it BYTE\rm ).  The definition is: \it update\_flag\_reg \rm =  $\lambda$  \rm (\it s7 \rm, \it z6 \rm,\it h4 \rm,\it pv2 \rm ,\it n1
\rm ,\it c0 \rm) . 
\rm ( \it s7 $\in$ \it BIT $\land$ \it z6 $\in$ \it BIT $\land$ \it h4 $\in$ \it BIT $\land$ \it pv2 $\in$ \it BIT $\land$ \it n1 $\in$ \it BIT $\land$ \it c0 $\in$ \it BIT \rm $\mid$( \it f0  $\mapsto$ \rm [\it c0\rm , \it n1\rm , \it pv2\rm , \rm 1\rm , \it h4\rm , \rm 1\rm , \it z6\rm , \it s7\rm \rm ]\rm ) \rm )

\subsection{Program, stack and data memory}

The Z80 uses a unique memory for storing program instructions, data stack and
general-purpose data. The memory has 16-bit addresses and each address holds a byte.
Thus, the memory is very simple: \it mem  $\in$  \it BV16  $\fun$  \it BYTE\rm.

In general, the instructions can access all memory address, but it is dangerous. The user may mistakenly
access and change data in memory. For added  security, it is important that the program instructions has
limited access by region. Thus the designer can specify address regions in a refinement model to restrict the access from
instructions. The address regions can be specified using constants
($\textit{PROGRAM\_R\_ADR,DATA\_R\_ADR,STACK\_R\_ADR}$), these define respectively a range of restricted
address for: programs instructions, general purpose data and data stack.
%
%$
%\begin{array}{l}
%\textit{PROGRAM\_R\_ADR} = 0..16384 \land\\
%\textit{DATA\_R\_ADR} = 16385..49151 \land\\
%\textit{STACK\_R\_ADR} = 49152..65535
%\end{array}
%$

\textbf{Assuring the absence of overlapping of address regions:}
 \emph{To assure that address regions are well defined, then the designer must to verify the expression:}
%\begin{sloppypar}
%\hspace*{0.10in}
\it PROGRAM\_R\_ADR $\cap$ DATA\_R\_ADR $\cap$ $=$ \{\} $\land$ STACK\_R\_ADR $\cap$ DATA\_R\_ADR $\cap$ $=$ \{\} $\land$ PROGRAM\_R\_ADR $\cap$  STACK\_R\_ADR  $=$ \{\} 
%\end{sloppypar}

\textbf{Preserving the consistency of the memory:} \emph{In general, the access to
some address regions is dangerous. Then, each instruction has a specific pre-condition that verify
if the new address memory, that will be updated, is member of its region. For example, the $\textit{PUSH}$
program instruction allows write only in the region of stack ($\textit{STACK\_R\_ADR}$).}

\subsection{Arithmetic logic unit}

There are many functions in the module \textit{ALU}. In general, the definition of these 
functions use basic definitions or previously defined functions. For example, the function
$\textit{half8UCHAR}$ is used to get the half part of $\textit{UCHAR}$ value.
It is important to know the half carry and it is used in the function $\textit{add8UCHAR}$.

\hspace*{0.0in}\it half8UCHAR  $\in$  \it UCHAR  $\fun$  \it UCHAR  $\land$

\hspace*{0.0in}\it half8UCHAR \rm =  $\lambda$  \rm (\it ww\rm )\rm .\rm (\it ww  $\in$  \it UCHAR  $\mid$  \it ww  $\mod$  \it $2^{4}$\rm )

The function $\textit{add8UCHAR}$ receives a bit carry and two $\textit{UCHAR}$ values and returns respectively the 
sum, the sign bit, the carry bit, the half carry bit and the zero bit. It is typed as follows: \it add8UCHAR \rm :
\rm (\it BIT $\times$ \it UCHAR $\times$ \it UCHAR\rm ) $\fun$ \rm (\it UCHAR $\times$  \it BIT  $\times$  \it BIT  $\times$  \it BIT  $\times$  \it BIT\rm ) and its definition is:

\hspace*{0.0in}\it add8UCHAR\rm = $\lambda$ \rm(\it carry\rm,\it w1\rm, \it w2\rm)\rm.\rm

\hspace*{0.0in}(\it carry $\in$  \it BIT  $\land$  \it w1  $\in$  \it UCHAR  $\land$  \it w2  $\in$  \it UCHAR
$\mid$

\hspace*{0.40in}\rm(\rm(\rm(\it carry \rm + \it w1 \rm + \it w2 \rm )  $\mod$  \it $2^{8}$ \rm ),

\hspace*{0.40in}\it bool\_bit\rm ( \it carry \rm + \it uchar\_schar\rm (\it w1\rm ) \rm + \it uchar\_schar \rm (\it
w2\rm ) $<$ \rm 0\rm ),

\hspace*{0.40in}\it bool\_bit\rm ( \it carry \rm + \it w1 \rm + \it w2 $>$ \it UCHAR\_MAX\rm )\rm ,

\hspace*{0.40in}\it bool\_bit\rm ( \it carry \rm + \it half8UCHAR\rm (\it w1\rm) \rm + \it half8UCHAR\rm (\it w2\rm)  $\geq$  \it $2^{4}$\rm )\rm,

\hspace*{0.40in}\it bool\_bit\rm ( \rm ( \rm (\it carry \rm + \it w1 \rm + \it w2 \rm )  $\mod$  \it $2^{8}$ \rm
)\rm = \rm 0\rm )\rm )\hspace*{0.10in}\rm )

\hspace*{0.20in}

A related function to subtract operation is $\textit{subtract8UCHAR}$. There are the same functions for the
$\textit{SCHAR}$ type, they are respectively $\textit{add8SCHAR}$ and $\textit{subtract8SCHAR}$, all these
functions are of 8 bits ($\textit{BYTE}$) and defined similarly. In the same way, the arithmetic functions for 16
bits ($\textit{BV16}$) are defined.
The module ALU has several others functions, for example:%but only some simplest functions also are explained below:
%\begin{itemize}
  %\item \it inc  $\in$ \it BYTE $\fun$  \it BYTE \rm - It receives a byte and
  %returns its increment. There is a similar function named $\textit{dec}$.
%  \item 
\it instruction\_next  $\in$  USHORT  $\fun$  USHORT \rm - It receives the  
  actual value from program counter register (\textit{pc}) and returns its increment;
% \item 
and \it is\_negative  $\in$  \it BYTE  $\fun$  \it BIT \rm - It returns the most significant bit,
 in other words, the signal bit.
  % \item \it is\_zero  $\in$  \it BYTE  $\fun$  \it BIT \rm - It returns 1 if the received
 % byte is zero, otherwise returns 0.
  %\item \it update\_refresh\_reg \rm - It receives a byte and returns its increment until the seventh bit.
%\end{itemize}
As the logic functions that are defined in the \textit{BYTE} and \textit{BV16}
module are included in the \textit{ALU} module, they can be seen and used
directly in the \textit{ALU} and \textit{Z80} modules.

\subsection{Modelling the actions and instructions}

Each instruction is represented by a B operation in the module Z80. The main
module (\textit{Z80}) has three categories of operations: a category represents the 
instructions of microcontrollers, a second category represents the input and output of data 
(shown in ~\ref{sec:modellingIO}) and the last category represents the external
actions (shown in~\ref{sec:externalactions}). A simple example of instruction category is
$\textit{LD\_(nn)\_A}$\footnoteremember{myfootnote}{The tools B does not allow
to use parentheses in identifiers, and the characters \{``('',``)''\} are
replaced respectively by \{``9'',``0''\} in the actual specification.} shown below. The
pre-defined functions are necessary many times to model the instructions, these
functions facilitate the construction of instruction set model. By default, all
parameters in operations are either predefined elements in the model or
integers values in decimal representation. This instruction use the
$\textit{updateAddressMem}$ operation from \textit{Memory} module and it receives
a address memory and its new memory value. It also increments the program
counter ($\textit{pc}$) and update the refresh register ($\textit{r\_}$).
The other instructions have a similar structure.

\hspace*{0.00in}\bf LD\_9nn0\_A \rm ( \it nn \rm ) \rm =

\hspace*{0.20in}\bf PRE \it nn $\in$ \it USHORT\hspace*{0.15in} $\land$ \hspace*{0.10in}\it nn\hspace*{0.10in} $\in$  \it DATA\_R\_ADR

\hspace*{0.20in}\bf THEN

\hspace*{0.20in}\bf updateAddressMem \rm ( \it ushort\_bv16 \rm ( \it nn \rm ) \rm , \it rgs8 \rm ( \it a0 \rm )
\rm )  $\para$

\hspace*{0.20in}\it pc \rm := \it instruction\_next \rm ( \it pc \rm )  $\para$  \it r\_ \rm := \it update\_refresh\_reg\rm (\it r\_\rm )

\hspace*{0.00in}\bf END\rm

\subsection{Modelling the input/output instructions}
\label{sec:modellingIO}
The Z80 has an extensive set of input and output (I/O) instructions and 256 ports for
devices. This model can transfer data blocks between the I/O devices and any
of the internal registers or memory address.

The $\textit{IN\_r(C)}$\footnoterecall{myfootnote} instruction is represented
by the following B operation. It receives an register identifier ``$\textit{rr}$'' and, it stores the value from ``rr'' in the $\textit{C}$ port address. Besides, it increments the program counter and updates the flag registers.

\hspace*{0.0in}\bf IN\_r\_9C0 \rm ( \it rr \rm ) \rm =

\hspace*{0.0in}\bf PRE \it rr  $\in$  \it id\_reg\_8  $\land$  \it rr  $\not =$  \it f0\hspace*{0.15in}\bf THEN

\hspace*{0.20in}\bf ANY

\hspace*{0.40in}\it negative \rm , \it zero \rm , \it half\_carry \rm , \it pv \rm , \it add\_sub \rm , \it carry

\hspace*{0.20in}\bf WHERE

\hspace*{0.40in}\it negative $\in$ \it BIT $\land$ \it zero $\in$ \it BIT $\land$ \it half\_carry $\in$ \it BIT 
$\land$ \it pv $\in$ \it BIT $\land$

 \hspace*{0.40in}\it add\_sub $\in$ \it BIT $\land$ \it carry $\in$ \it BIT  $\land$

\hspace*{0.40in}\it negative \rm = \it is\_negative \rm ( \it io\_ports \rm ( \it rgs8 \rm ( \it c0 \rm ) \rm ) \rm )  $\land$

\hspace*{0.40in}\it zero \rm = \it is\_zero \rm ( \it io\_ports \rm ( \it rgs8 \rm ( \it c0 \rm ) \rm ) \rm )  $\land$

\hspace*{0.40in}\it half\_carry \rm = \rm 0  $\land$

\hspace*{0.40in}\it pv \rm = \it parity\_even \rm ( \it io\_ports \rm ( \it rgs8 \rm ( \it c0 \rm ) \rm ) \rm ) $\land$

\hspace*{0.40in}\it add\_sub \rm =\hspace*{0.10in}\rm 0  $\land$

\hspace*{0.40in}\it carry \rm = \it z\_c

\hspace*{0.20in}\bf THEN

\hspace*{0.40in}\it rgs8 \rm := \it rgs8  $\lover$  \rm \{ \rm ( \it rr  $\mapsto$  \it io\_ports \rm ( \it rgs8 \rm ( \it c0 \rm ) \rm ) \rm ) \rm ,

\hspace*{0.40in}\it update\_flag\_reg\rm (\it negative\rm,\it zero\rm,\it half\_carry\rm,\it pv\rm,\it add\_sub\rm,\it carry)\rm\}$\para$

\hspace*{0.40in}\it pc \rm := \it instruction\_next \rm ( \it pc \rm )  $\para$  \it r\_ \rm := \it update\_refresh\_reg\rm (\it r\_\rm )

\hspace*{0.0in}\bf END \rm

%The main I/O instructions work similarly, for example: $\textit{OUT(n),A}$  or $\textit{OUT (C ), r}$.

% $
% \begin{array}{l} 
% \textit{IN\_r\_9C0} ( rr ) = \\
% \quad   \PRE rr \in id\_reg\_8    \THEN\\
% \quad\quad \ANY data\_in, negative , zero , half\_carry , pv , add\_sub , carry\\ 
% \quad\quad \WHERE data\_in \in \textit{BYTE} \land negative \in \textit{BIT}\land\\
% \quad\quad\quad carry \in \textit{BIT} \land half\_carry \in \textit{BIT} \land zero \in \textit{BIT} \land \\
% \quad\quad\quad negative = is\_negative (data\_in) \land zero = is\_zero(data\_in ) \land\\
% \quad\quad\quad half\_carry = 0 \land pv =parity\_even\_BYTE ( data\_in )    \land\\
% \quad\quad\quad add\_sub =  0 \land carry = z\_c \\
% \quad\quad  \THEN \\
% \quad\quad\quad i\_o\_ports ( rgs8 ( c0 ) ) := data\_in ||\\
% \quad\quad\quad rgs8 := rgs8 <+ \{ ( rr |-> data\_in ) ,\\
% \quad\quad\quad get\_new\_flag\_register\_SZ\_H\_PvNC ( rgs8 , negative , zero, half\_carry , pv , add\_sub , carry ) \} ||\\
% \quad\quad\quad pc := instruction\_next( pc )\\
% \quad\quad \END\\
% \quad \END\\
% \end{array}
% $

\subsection{Modelling the external actions}
\label{sec:externalactions}

The external actions change the state of the microcontroller, for example, refreshing the I/O ports
and interruptions request. The external actions are also modelled by operations and they are named with the
prefix ``$ext\_$'' and followed by the name of action. There are four external
actions: $ext\_update\_io\_ports$, $ext\_NMI$ and $ext\_INT$, $ext\_Reset$. The
$ext\_update\_io\_ports$ just change the state of I/O port, see.

\hspace*{0.20in}\bf ext\_update\_io\_ports\rm (\it address\rm ,\it value\rm )\rm =

\hspace*{0.20in}\bf PRE \it address  $\in$  \it UCHAR  $\land$ \hspace*{0.10in}\it value  $\in$  \it SCHAR \bf THEN

\hspace*{0.40in}\it io\_ports \rm ( \it uchar\_byte \rm ( \it address \rm ) \rm ) \rm := \it schar\_byte \rm ( \it
value \rm )

\hspace*{0.20in}\bf END\rm

The other external actions are related to interruptions. The interruptions allow
that the devices suspend a routine from CPU and start another service routine.
This service routine can exchange data or signals between CPU and external
devices. When a routine is finished, then the CPU comes back to the last routine
that was interrupted.

For the interrupts, the following elements are important:  the interrupt flip-flops
($\textit{iff1}$ and $\textit{iff2}$), the types of interrupts (maskable and
non-maskable), the interrupt mode (set with the $\textit{IM 0}$, $\textit{IM 1}$,
$\textit{IM 2}$ instructions) and the $\textit{i\_}$ register.

Flip-flops $\textit{iff1}$ and $\textit{iff2}$ control the maskable interrupts
($\textit{INT}$). When $\textit{iff1}$ is set, the interrupt is enabled,
otherwise it is disabled; $\textit{iff2}$ is used only as backup for $\textit{iff1}$. The
instructions $\textit{EI}$ and $\textit{DI}$ respectively enable and disable the
maskable interruptions, setting  $\textit{iff1}$ to 1 and 0.

The interruptions and the \textit{reset} action can change the state of
program counter. Theses actions are modelled by B operations and its main
effects are shown below\footnote{Some definitions of constants: $\textit{sp\_minus\_two}$ holds the value of stack pointer minus 2,
%=$\textit{dec\_BV16(dec\_BV16(sp))}$
 $\textit{sp\_minus\_one}$ is the value of stack pointer minus 1,
%=$\textit{dec\_BV16(sp)}$
$\textit{pc\_high}$ holds the most significant 8 bits and
$\textit{pc\_low}$ holds the least significant 8 bits.}.

 \textbf{NMI} - Non-maskable interrupts cannot be disabled
 by the programmer. Then, when a device makes a request, $sp$ is pushed,
 $pc$ receives $66H$ (102 in decimal), $\textit{iff1}$ is reset , $\textit{iff2}$ stores
 $\textit{iff1}$ and the refresh register is updated.

\begin{sloppypar}
\bf updateStack\rm (\rm \{ \rm (\it sp\_minus\_two  $\mapsto$  \it pc\_low\rm )\rm ,\rm (\it sp\_minus\_one  $\mapsto$ \it pc\_high \rm ) \rm \}\rm )$\para$

\it sp \rm := \it sp\_minus\_two  $\para$ \it pc \rm := \rm 1\rm 0\rm 2 $\para$ \it iff1\rm :=\rm 0  $\para$  \it iff2\rm := \it iff1 $\para$

\it r\_ \rm := \it update\_refresh\_reg\rm (\it r\_\rm )\\
\end{sloppypar}

  \textbf{INT} - Maskable Interrupt is usually reserved for important functions that can be enabled and
  disabled by the programmer. When a maskable interrupt action happens, both $\textit{iff1}$ and $\textit{iff2}$ are
  cleared, disabling the interrupts, $sp$ is pushed, the refresh register is updated and the other effects
  depend on the interrupt mode register (\it im\rm).

 \begin{itemize}

  \item The mode 0 is compatible with 8080 and this mode is selected when \it im \rm = \rm ( \rm 0 $\mapsto$  \rm 0 \rm ). When a 
  non-maskable interrupt happens, the CPU fetches an instruction of one byte
  from an external device, usually an RST instruction, and the CPU executes it.
  The instruction code is received from an external device by data bus and it is represented by integer parameter called $\textit{byte\_bus}$ .

  \item The mode 1 is the simplest and this mode is selected when \it im \rm = \rm ( \rm 0  $\mapsto$  \rm 1
  \rm ). Simply, when a non-maskable interruption happens, the program counter
  receives $38H$ (56 in decimal).

  \item The mode 2 is the most flexible and this mode is selected when \it im \rm = \rm ( \rm 1  $\mapsto$
  \rm 1 \rm ). When a non-maskable interruption happens, an indirect call can be
  made to any address memory. The program counter receives a bit vector of size
  16 composed with two parts: the most significant part of the $\textit{i\_}$
  register and the least significant part of the $\textit{byte\_bus}$ with the
  last bit cleared.

 \end{itemize}

The essential part of maskable interrupt is shown below, where $\textit{byte\_bus}$ is a parameter of the $\textit{INT}$ operation:

\begin{sloppypar}

\hspace*{0.00in}\bf IF \it im \rm = \rm ( \rm 0  $\mapsto$  \rm 0 \rm ) \bf THEN

\hspace*{0.10in}\bf IF\hspace*{0.10in}\it byte\_bus \rm $\in$  \it opcodes\_RST\_instruction

\hspace*{0.10in}\bf THEN\hspace*{1.05in}

\hspace*{0.40in}\it pc \rm := \it byte\_bus \rm - \rm 1\rm 9\rm
9\hspace*{0.10in} $\para$

\hspace*{0.40in}\bf updateStack\rm ( \rm \{ \it stack\rm ( \it
sp\_minus\_one\rm )  $\mapsto$  \it pc\_low\rm ,

\hspace*{0.40in}\it stack\rm (\it sp\_minus\_two\rm )  $\mapsto$  \it pc\_high
\rm \} \rm )  $\para$

\hspace*{0.40in}\it sp \rm := \it sp\_minus\_two  $\para$ \hspace*{0.10in}\it
r\_ \rm := \it update\_refresh\_reg\rm (\it r\_\rm )

\hspace*{0.10in}\bf ELSIF \it byte\_bus \rm = \it opcode\_\ldots\_instruction

\hspace*{0.40in}\bf \ldots

\hspace*{0.10in}\bf END

\hspace*{0.00in}\bf ELSIF\hspace*{0.10in}\it im \rm =\hspace*{0.10in}\rm ( \rm
0  $\mapsto$  \rm 1 \rm ) \bf THEN

\hspace*{0.10in}\it pc \rm :=\hspace*{0.10in}\rm 5\rm 6  \hspace*{0.80in}
$\para$

\hspace*{0.10in}\bf updateStack\rm ( \rm \{ \it stack\rm (\it
sp\_minus\_one\rm )  $\mapsto$  \it pc\_low\rm ,

\hspace*{0.10in}\it stack\rm (\it sp\_minus\_two\rm )  $\mapsto$  \it pc\_high
\rm \} \rm )  $\para$

\hspace*{0.10in}\it sp \rm := \it sp\_minus\_two  $\para$ \hspace*{0.10in}\it
r\_ \rm := \it update\_refresh\_reg\rm (\it r\_\rm )\hspace*{0.85in}

\hspace*{0.00in}\bf ELSIF\hspace*{0.10in}\it im \rm = \rm ( \rm 1  $\mapsto$ 
\rm 1 \rm ) \bf THEN  \hspace*{0.70in}

\hspace*{0.10in}\it pc \rm := \it bv16\_ushort\rm (\it byte\_bv16\rm ( \it i\_
\rm ,\it bv\_clear\rm (\it rotateleft\rm (\it uchar\_byte\rm (\it byte\_bus\rm )\rm )\rm ,\rm 0\rm )\rm )\rm )$\para$

\hspace*{0.10in}\bf updateStack\rm ( \rm \{ \it stack\rm ( \it
sp\_minus\_one\rm )  $\mapsto$  \it pc\_low\rm ,

\hspace*{0.10in}\it stack\rm (\it sp\_minus\_two\rm )  $\mapsto$  \it pc\_high
\rm \} \rm )  $\para$

\hspace*{0.10in}\it sp \rm := \it sp\_minus\_two  $\para$ \hspace*{0.10in}\it
r\_ \rm := \it update\_refresh\_reg\rm (\it r\_\rm )

\hspace*{0.00in}\bf END\hspace*{0.10in}\\
\end{sloppypar}

\textbf{RESET}  - This just resets the registers related to the interruptions.

\begin{sloppypar}
\it iff1 \rm :=\rm 0 $\para$ \it iff2\rm :=\rm 0 $\para$  \it  im\rm := \rm (\rm 0 $\mapsto$ \rm 0\rm )  $\para$ 
\it pc\rm :=\rm 0 $\para$ \it i\_ \rm := [0,0,0,0,0,0,0,0] $\para$

\it rgs8 \rm := \it rgs8  $\lover$  \rm \{ \rm (\it a0  $\mapsto$ \rm [0,0,0,0,0,0,0,0] , \rm (\it
f0 $\mapsto$ \rm [0,0,0,0,0,0,0,0] \rm \} $\para$

\it r\_  \rm := [0,0,0,0,0,0,0,0] $\para$ \it sp \rm := [0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0]\\
\end{sloppypar}

% The animation

The modelling of instruction set, interruptions and input/output ports become the animation interesting.
The animation in ProB version is not so fast and practical as a simulation, for example in \cite{Simulator_z80},
because ProB works manipulating logic and math concepts that can have infinite sets and non determinism, then ProB requires more processing, specially when it works analysing  complex expressions or very large expressions.
However, a user of ProB with this modelling can: execute manually a sequence of Z80's instructions that represent a program and view the visited states of ports, memory and registers, it is important to debug the effect action on the state space computed; create a tracing of the states; analyse constants, functions and expressions; and search possible violations.

%\subsection{Considerations about the Z80 model}

The Z80 formal model provides many benefits, because of the verification of
generated proof obligations guarantees: the correct use of data types, the developed
security properties  and the well-defined of all the expressions. %\footnote{An expression is called ``well-defined'' (or unambiguous) if its definition assigns it a unique interpretation or value.}.
Furthermore, the designer has a big flexibility to create new and specific security properties, this is very useful
to adjust the verification in accordance with the requirements. Moreover, this
example of model could replace or improve the used documentation for users and
assembly programmers. Z80 model is also useful to develop verified software 
up to the assembly level. There are many benefits when a formal model is
developed, but to verify the model is not an easy task, the first version of Z80 model have already verified completely,but the new Z80 model with support to animation in ProB has not been verified completely,
because the changes in Z80 model are recent and the verification time is long.

A case study using the first version of Z80 model was also created and verified
since the first abstract B model up to B assembly model. This case study is an
object of the pilot project developed to analyse a petroleum production test
system in each oil field. Its modelling was developed in according to \cite{LAUT_SERGIO}
and its B assembly model can be executed manually in ProB. The modelling of
instruction set and the B assembly model allow to simulate the execution of microcontroller
and analyse several important properties.

\section{Related works}
\label{sec:relatedworks}

There are in the literature several similar approaches to model hardware and instruction set using B.
The paper \cite{springerlink:Yuan2011} reports a method to specify, design and construct sound and complete instruction set model by stepwise refinement and formal proof using the formal method Event-B. It also discuss about desirable properties of instruction set model. However, our work \cite{Valerio_SBMF09} and \cite{Subotic2010} use the B method, that seems more appropriated to software development, because it has implementable language defined, 
called B0, and tools to convert the models to a programming language.
In \cite{Marc20113} is present an related experience on using B in the design of secure micro-controllers. It tries to show the feasibility of such a technique for high confidence trustful devices. The work \cite{Subotic2010} describes a similar approach with MIPS CPU architecture and it shows how the CPU is formally specified and implemented using the B method.

% 
%There are in the literature several
%approaches~\cite{BHDL_2003,DBLP:journals/entcs/EvansG08,DBLP:conf/fmco/Leuschel08,DBLP:conf/asm/Wright08}
%to model hardware and the virtual machines using B. In these works, the B
%has been used successfully to model the operational semantic. However the cost of modelling is still expensive. 
%
%These recent works \cite{Marc20113,springerlink:Yuan2011} have a similar approach using Event B. 
%However, our work use the B method, that seems more appropriated to software development, because it has
%implementable language defined, called B0, and tools to convert the models to a programming language.
%
%
%The main motivation of our research is the development of verified
%software up to the assembly level, which requires specifying the
%semantics of the underlying hardware. Thus some aspects were not
%modelled in our work such as the execution time of the instructions.
%Also we did not consider the micro architecture of the hardware as the scope of
%our work does not include hardware verification.

\section{Conclusions}
\label{sec:conclusions}

% Um pequeno resumo do trabalho
This work has shown an approach to the formal modelling of the instruction set of
microcontroller using the B method. During the construction of this model, some
problems were found in the official reference for Z80
microcontroller~\cite{Z80_manual}. The designer fixed the found problems in
documentation and also developed a case study: a verified embedded software. This
case study was interesting to analyse the technique of formal verification up to
assembly level.

% TODO: Adaptar para apenas fundamentar melhor o trabalho já desenvolvido. Citar o artigo SBMF e a dissertação. 
% Os meus trabalhos já desenvolvidos/ resultados atuais.
The next works quoted are directly related to the objective this paper.
The first work \cite{DantasSemish2008} shows the developed metodology to verify
software up to assembly level using B. A second work\cite{Dantas_SBMF08}
presents more details about the verification approach and a small example
software verified up to assembly level in three different platforms. 
The last work \cite{Valerio_SBMF09} shows a general view of the first version of Z80
model and techniques used in verification process.

%Link com o parágrafo anterior 
These works created a important step to software verification up to assembly
language, they show the actuals difficulties and suggest improvements in the
B tools.

% TODO
% [Trabalhos futuros ] LLVM:Existe também a pretensão de especificar as

Future works comprise the development of software with the B method from
functional specification to assembly level, using the Z80 model presented in this
work. The mechanic compilation from B algorithmic constructs to assembly platform
is also envisioned. Finally, another important activity is develop a formal model of a platform used in LLVM Compiler~\cite{DBLP:conf/cgo/LattnerA04}.

\subsubsection{Acknowledgements:} This experiment could not have succeeded without the interesting work of Michael Leuschel who developed several improvements in the ProB and provided important aid to enable the animation of the Z80 modelling.

% This work was partially supported by INES (www.ines.org.br),
%funded by CNPq grant 573964/2008-4 and by CNPq grants 553597/2008-6, 550946/2007-
%1, and 620132/2008-6

%\paragraph{Acknowledges:}
%This work was partially supported by INES (www.ines.org.br), funded by CNPq
%grant \ldots
%573964/2008-4 and by CNPq grants 553597/2008-6, 550946/2007-1, and 620132/2008-6..

%\bibliographystyle{splncs}

\bibliographystyle{plain}
\bibliography{paper}
% [To change the format; to add: thesis suggested by David, Article from ABZ 2010 suggested by David  ]
% \begin{thebibliography}{5}
%

\end{document}

%
% \bibitem {clar:eke}
% Clarke, F., Ekeland, I.:
% Nonlinear oscillations and
% boundary-value problems for Hamiltonian systems.
% Arch. Rat. Mech. Anal. {\bf 78} (1982) 315--333
% %
% \bibitem {clar:eke:2}
% Clarke, F., Ekeland, I.:
% Solutions p\'{e}riodiques, du
% p\'{e}riode donn\'{e}e, des \'{e}quations hamiltoniennes.
% Note CRAS Paris {\bf 287} (1978) 1013--1015
% %
% \bibitem {mich:tar}
% Michalek, R., Tarantello, G.:
% Subharmonic solutions with prescribed minimal
% period for nonautonomous Hamiltonian systems.
% J. Diff. Eq. {\bf 72} (1988) 28--55
% %
% \bibitem {tar}
% Tarantello, G.:
% Subharmonic solutions for Hamiltonian
% systems via a $\bbbz_{p}$ pseudoindex theory.
% Annali di Matematica Pura (to appear)
% %
% \bibitem {rab}
% Rabinowitz, P.:
% On subharmonic solutions of a Hamiltonian system.
% Comm. Pure Appl. Math. {\bf 33} (1980) 609--633
% \end{thebibliography}
%